### 2011-06-07 paper/paper.tex
started on abstract:

% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
% APRIL 2009
%
% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
% LaTeX2e document class file for Conference Proceedings submissions.
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) Page numbering
% ---------------------------------------------------------------------------------------------------------------
% It is an example which *does* use the .bib file (from which the .bbl file
% is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% Questions regarding SIGS should be sent to
% Adrienne Griscti ---> griscti@acm.org
%
% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
% Gerald Murray ---> murray@hq.acm.org
%
% For tracking purposes - this is V3.1SP - APRIL 2009

\documentclass{acm_proc_article-sp}

\begin{document}
%TODO
\title{Compareing Power Optimizations in the Reorder Buffer}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Chen Guo\\
       \affaddr{UCLA Computer Science Department}\\
       \affaddr{4732 Boelter Hall}\\
       \affaddr{Los Angeles, CA 90095}\\
% 2nd. author
\alignauthor
Alex Edwards\\
       \affaddr{UCLA Computer Science Department}\\
       \affaddr{4732 Boelter Hall}\\
       \affaddr{Los Angeles, CA 90095}\\
% 3rd. author
\alignauthor
Henry Hsieh\\
       \affaddr{UCLA Computer Science Department}\\
       \affaddr{4732 Boelter Hall}\\
       \affaddr{Los Angeles, CA 90095}\\
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
In an effort to increase instruction level parallelism techniques such 
as pipelining, out of order execution and operand forwarding.  However 
these efforts come at a large power cost.  Out of order execution requires
a ReOrder Buffer (ROB), to commit the registers in program order.  This 
structure has grown to be monolithic numerous entries and read/write ports.
There is no current methods to efficiently and equally compare various
ROB advancements to improve power efficiency.  
\end{abstract}

\terms{Theory}

\keywords{ACM proceedings, \LaTeX, text tagging} % NOT required for Proceedings

\section{Introduction}
BLAH BLAH BLAH INTRO WEEEE
\section{TITLE1}
hi
\subsection{subsection title}
sub section.. with table!
\begin{table*}
\centering
\caption{Some Typical Commands}
\begin{tabular}{|c|c|l|} \hline
Command&A Number&Comments\\ \hline
\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
\texttt{{\char'134}table}& 300 & For tables\\ \hline
\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
\end{table*}
% end the environment with {table*}, NOTE not {table}!

\section{Conclusions}
ending stuff
%\end{document}  % This is where a 'short' article might terminate

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
if we really need this....

%TODO add in bib

\end{document}

### 2011-06-07 src/rob_latch.cpp

@@ -267,7 +267,8 @@ void ROB_Latch::run (ins_t ins[])
       update_power_totals();
 
       cycles++;
-      print_msgs (cycles);
+      if (m_print & DBG_FLAG)
+        print_msgs (cycles);
 
 
       // Break when we've written all the instructions.
@@ -277,7 +278,7 @@ void ROB_Latch::run (ins_t ins[])
 
   print_power_stats(cycles);
 
-  cout << "Simulation complete." << endl;
+//  cout << "Simulation complete." << endl;
 }
 
 

### 2011-06-07 src/main.cpp
Completed instruction profile generation.
Binary now accepts a mandatory option "-b" to indicate benchmark instruction profile
-b 0     Utilize default profile.  All inst types are equally likely, 10% forwarding rate.
-b 1     Run 173.applu SPEC-FP Parabolic/Elliptic Partial Diff. Equations benchmark. Emphasis on FP ops.
-b 2     Run 191.fma3d SPEC-FP Finite-element Crash Simulation benchmark. Emphasis on INT ops.
-b 3     Run 183.equake SPEC-FP Seismic Wave Propagation benchmark.  Mix of both FP and INT ops.

Check out the main.cpp code if you have questions.  I figure we can just use the thr

@@ -21,7 +21,7 @@ enum {
 };
 
 void help ();
-ins_t *generate_instruction_stream();
+ins_t *generate_instruction_stream(int benchmark);
 int generate_type(ins_profile input, int gen);
 uint32_t generate_register(uint32_t history, float forwardRate);
 
@@ -31,12 +31,13 @@ int main (int argc, char *argv[])
   srand(2);
   char opt;
   int mode = MODE_BASE;
+  int bench = 0;
   ROB *rob;
   ins_t *instructions;
   int rob_size = 128;
   int print_flags = 1;
 
-  while ((opt = getopt (argc, argv, "dhm:")) != -1)
+  while ((opt = getopt (argc, argv, "dhm:b:")) != -1)
     {
       switch (opt)
         {
@@ -66,6 +67,20 @@ int main (int argc, char *argv[])
               return 1;
             }
           break;
+        case 'b':
+          if (strcmp(optarg, "0") == 0)
+            bench = 0;
+          else if (strcmp(optarg, "1") == 0)
+            bench = 1;
+          else if (strcmp(optarg, "2") == 0)
+            bench = 2;
+          else if (strcmp(optarg, "3") == 0)
+            bench = 3;
+          else {
+            cerr << "Unrecognized benchmark." << endl;
+            return 1;
+          }
+          break;
 
         default: break;
 
@@ -74,7 +89,7 @@ int main (int argc, char *argv[])
   srand (10);
 
   // For now, generate arbitrary mix of instructions.
-  instructions = generate_instruction_stream();
+  instructions = generate_instruction_stream(bench);
 
   switch (mode)
     {
@@ -122,14 +137,20 @@ void help ()
   cout << "  -d         Enable debug messages." << endl;
   cout << "  -h         Display this help message." << endl;
   cout << "  -m         Set operating mode. See \"Modes\"" << endl;
+  cout << "  -b         Set benchmark. See \"Benchmarks\"" << endl;
   cout << "Modes:" << endl;
   cout << "  circular   Circular ROB." << endl;
   cout << "  dynamic    Dynamically resized ROB." << endl;
   cout << "  dist       Distributed ROB." << endl;
   cout << "  latch      ROB with retention latches." << endl;
[…]

### 2011-06-07 paper/paper.tex
Adding bits to the paper.  Just pushing for shits.

@@
 %TODO
-\title{Compareing Power Optimizations in the Reorder Buffer}
+\title{Comparing Power Optimizations in the Reorder Buffer}
 %
@@
 \alignauthor
-Alex Edwards\\
+Alexander Edwards\\
        \affaddr{UCLA Computer Science Department}\\
@@
 BLAH BLAH BLAH INTRO WEEEE
-\section{TITLE1}
+\section{Project Implementation}
+Our simulator provides a cycle accurate behavioral model of a standard circular re-order buffer.
+At the outset of the project, the standard re-order buffer we intended to implement was a non-circular RAM-based re-order buffer.
+However, after feedback from the course instructor and students it was decided that a circular buffer is more widely held as being the 'standard' buffer type and would serve as an ideal underlying model for power optimizations.
+Thus, we implement and compare the standard circular re-order buffer to circular re-order buffers with the following optimizations: 1) Distributed (Cluster based) ROB, 2) Dynamically Sized ROB, 3) Reduced Output Port ROB (Retention Latch).
+The following sections outline each of the above model implementations as well as input instruction generation and power results tabulation.
+\subsection{Instruction Generation}
+\subsection{Model Implementation}
+\subsubsection{Circular ROB}
+\subsubsection{Distributed ROB}
+\subsubsection{Dynamic ROB}
+\subsubsection{Reduced Output Port ROB (Retention Latch)}
+\subsection{Power Tabulation Implementation}
+\section{Results}
 hi
-\subsection{subsection title}
+\subsection{subsection titl}
 sub section.. with table!

### 2011-06-07 paper/paper.tex
Adding the instruction generation section.  Planning on including another chart to denote each of the SPEC-FP workloads as well as inserting paragraphs where I wanted them originally ....

@@
 \subsection{Instruction Generation}
+	We model the input to our simulator as a stream of integer and floatpoint instructions.
+The instructions are represented at the bit-level and model MIPS instructions as closely as possible.
+Each instruction consists of 1) Instruction type, 2) Source and destination registers 3) a PC.
+Table XX indicates the types of instructions modeled.
+Each instruction type also has a predetermined cycle execution time designating the number of cycles that must elapse between instruction issue and the completion of the instruction.
+As modeling functional units goes beyond the scope of this simulator, instructions are flagged as complete after the listed number of cycles has elapsed.
+The actual cycle counts are the same as those modeled in MIT's Graphite many-core processor simulator[REFXX], with the exception of LOAD and STORE.
+The cycle counts corresponding to LOAD and STORE instructions are estimated as a predetermined 20 cycles each.
+This is simply an estimation as the actual cycle count is heavily dependent on cache architecture and benchmark specific cache behavior.
+Since both of these topics are beyond the scope of the simulator, we estimate the completion time of LOAD and STORE instructions as a set value substantially longer than the other more accurate instruction types.
+\begin{table*}
+\centering
+\begin{tabular}{|c|c|c|c|} \hline
+Instruction Type&Cycle Count&Instruction Type&Cycle Count\\ \hline
+\texttt{LOAD}&20&\texttt{STORE}&20\\ \hline
+\texttt{ADD}&1&\texttt{FADD}&3\\ \hline
+\texttt{SUB}&1&\texttt{FSUB}&3\\ \hline
+\texttt{MULT}&3&\texttt{FMULT}&5\\ \hline
+\texttt{DIV}&18&\texttt{FDIV}&6\\ \hline\end{tabular}
+\caption{Modeled instruction types and their corresponding completion cycle counts.}
+\end{table*}
+	Registers are generated in bit representation as 4 bits each totalling 12 bits for the s1, s2 and d registers in each instruction.
+Currently, our representation of instructions make no distinction between zero, 1, and 2 operand instructions.
+All instructions regardless of type feature two operands.
+This feature is supported by the fact that a memory hierarchy and functional units are not simulated by our system and as such immediate values have no direct use.
+Further, every instruction features a d register regardless of type.
+Because our simulator features operand forwarding from the ROB, it is necessary for the generation of registers to keep track of past destination registers to use as dependent operands in future instructions.
+This history buffer is implemented as a rotating bit register from which prior destination registers can be indexed.
+	Each instruction also features a program counter associated with the instruction.
+While this is clearly a simplification of a real system, it serves the purposes of our simulation as a uniquely identifiable tag by which to differentiate instructions.
+In both a real system and our simulator, a statically set 32 bit value is used to represent this PC/tag.
+As such, the specific values contained in the PC/tag will yield accurate transition results (nonexistent) and is thus are irrelevant.
+Simply starting the PC of the first instruction in the stream at 0 and incrementing by 4 each instruction guarantees uniqueness among PCs and serves the purposes of an instruction tag.
+	The aforementioned instruction fields have multiple modes of generation.
+Our simulator synthetically generates an instruction input trace based on a prescribed instruction profile.
+The instruction profile features: 1) a percent rate of occurrence for each instruction type, 2) a rate corresponding to the interdependence of registers between instructions, and 3) an overall instruction count.
+	The instruction generator is currently capable of generating four distinct instruction streams corresponding to different instruction workloads.
+The first instruction stream is marked as the default instruction stream.
+It features equal percentages of each type of instruction as well as a 10% register dependence rate.
+This register dependence rate was selected arbitrarily to act as a starting point for modeling data dependencies between instructions.
+	The remaining three instruction streams are modeled after SPEC-FP benchmarks.
+The benchmarks were chosen to represent three major modes of operation:  floating point heavy execution, integer heavy execution, and mixed type execution.
+For a floating point heavy execution, 173.applu was selected and modeled.
+173.applu models parabolic/elliptical partial differential equations and features almost exclusively floating point operations as well as a large number of memory operations making it useful as both a floating point benchmark as well as a long running memory intensive benchmark.
+191.fma3d is used to model integer heavy execution.
+It models finite-element crash simulation and features very few floating point operations and substantially fewer memory operations than 173.applu making it an ideal light-weight integer workload.
+Finally, 183.equake is used to represent the average workload.
+It models seismic wave propagation and features both integer and floating point operations as well as moderate memory usage.
+As such it acts as a more representative general workload than the original default workload.
+Each of the SPEC-FP benchmarks feature a 6% register forwarding frequency as is reported by [XX paper ref].
+	For the purposes of our power comparisons, each of the above instruction workloads are normalized to generate 10000 instructions.
+Such a large instruction window is ideal to allow buffers to fill and behavioral characteristics to develop.
+The specific values used for each workload are listed in Table XX.
+
 \subsection{Model Implementation}

### 2011-06-07 src/common.h
max ins did not properate to latch removed exe in ins struc

@@ -34,7 +34,7 @@ typedef struct instruction {
   int type;
   uint32_t pc;
   uint16_t regs; // 4 bits each reg id, -blank-reg1,-reg2-regdest
-  bool exec; // true is already executed
+//  bool exec; // true is already executed
 } ins_t;
 
 // Enumerated instruction types.

### 2011-06-07 src/main.cpp

@@ -248,13 +248,13 @@ ins_t *instructions = new ins_t [input.instCount+1];
       regD = rand()%0xF;
       reg_history_buffer = (reg_history_buffer << 4) | (regD&0xF);
       instructions[i].regs = ((regS1&0xF) << 8) | ((regS2&0xF) << 4) | (regD&0xF);
-      instructions[i].exec = false;
+//      instructions[i].exec = false;
       pc_start += 4;
-      cout << "Type=" << instructions[i].type << " PC=" << instructions[i].pc << " REGS S1 S2 D= " << ((instructions[i].regs >> 8)&0xF) << " " << ((instructions[i].regs >> 4)&0xF) << " " << (instructions[i].regs&0xF) << " EXEC=" << instructions[i].exec << endl;
+      cout << "Type=" << instructions[i].type << " PC=" << instructions[i].pc << " REGS S1 S2 D= " << ((instructions[i].regs >> 8)&0xF) << " " << ((instructions[i].regs >> 4)&0xF) << " " << (instructions[i].regs&0xF) << endl;
     }
   instructions[input.instCount].type = -1;
   instructions[input.instCount].pc = -1;
-  cout << "Type=" << instructions[input.instCount].type << " PC=" << instructions[input.instCount].pc << " REGS S1 S2 D= " << ((instructions[input.instCount].regs >> 8)&0xF) << " " << ((instructions[input.instCount].regs >> 4)&0xF) << " " << (instructions[input.instCount].regs&0xF) << " EXEC=" << instructions[input.instCount].exec << endl;
+  cout << "Type=" << instructions[input.instCount].type << " PC=" << instructions[input.instCount].pc << " REGS S1 S2 D= " << ((instructions[input.instCount].regs >> 8)&0xF) << " " << ((instructions[input.instCount].regs >> 4)&0xF) << " " << (instructions[input.instCount].regs&0xF) << endl;
 
   return instructions;
 }

### 2011-06-07 src/rob_latch.cpp

@@ -103,23 +103,41 @@ void ROB_Latch::write_to_arf ()
 // Read instructions from issue, up to m_n instructions.
 void ROB_Latch::write_entry (entry *entry, ins_t ins)
 {
-  uint16_t reg_mask = 0xF;
+	uint16_t reg_mask = 0xF;
+	entry->valid = false;
+	entry->cycles = ins_cyc[ins.type];
+	if (ins.type >= FADD)
+		entry->cycles += m_fp_delay;
+		entry->pc = ins.pc;
+		entry->reg_id = ins.regs & reg_mask;
+		entry->isfp = (ins.type >= FADD);
+		
+		m_tail = tail_incr (m_tail);
+		uint32_t operandRegA = ((ins.regs >> 4) & reg_mask);
+		uint32_t operandRegB = ((ins.regs >> 8) & reg_mask);
+		uint32_t extraCyclesA = 0;
+		uint32_t extraCyclesB = 0;
+		if(ReadinROB (operandRegA)) {
+			extraCyclesA = getCyclesToCompletion(operandRegA);
+			//cout << "Forwarding ins.pc=" << ins.pc << " operandA in " << extraCyclesA << " cycles" << endl;
+			//m_nwdu++;
+		}
+	if(ReadinROB (operandRegB)) {
+		extraCyclesB = getCyclesToCompletion(operandRegB);
+		//cout << "Forwarding ins.pc=" << ins.pc << " operandB in " << extraCyclesB << " cycles" << endl;
+		//m_nwdu++;
+	}
+	uint32_t maxExtraCycles = max(extraCyclesA, extraCyclesB);
+	if(maxExtraCycles)
+		//cout << "Increasing ins.pc=" << ins.pc << " cycle count by " << maxExtraCycles << " to allow forwarding" << endl;
+		entry->cycles += maxExtraCycles;
+		}
 
-  entry->valid = false;
-  entry->cycles = ins_cyc[ins.type];
-  if (ins.type >= FADD)
-    entry->cycles += m_fp_delay;
-  entry->pc = ins.pc;
-  entry->reg_id = ins.regs & reg_mask;
-  entry->isfp = (ins.type >= FADD);
 
-  m_tail = tail_incr (m_tail);
 
-  ReadinROB ((ins.regs >> 4) & reg_mask);
-  ReadinROB ((ins.regs >> 8) & reg_mask);
-}
 
-//TODO CHANGE add stuff
+
+
[…]

### 2011-06-07 src/rob_latch.h

@@ -27,7 +27,7 @@ class ROB_Latch : public ROB_Circ {
     lentry* lbuf;
     lentry* lbuf_prev;
 
-    void ReadinROB(uint16_t reg);
+    bool ReadinROB(uint16_t reg);
     bool ReadinLatch(uint16_t reg);
     uint32_t fwdcnt;
 

### 2011-06-07 src/main.cpp
rob latch prints

@@ -115,7 +115,7 @@ int main (int argc, char *argv[])
 
     case MODE_LATCH:
       cout << "Retention latches" << endl;
-      rob = new ROB_Latch (rob_size, 3, 1, 4, print_flags);
+      rob = new ROB_Latch (rob_size, 3, 1, 16, print_flags);
       break;
 
     default: break;

### 2011-06-07 paper/paper.tex
Buggy.  Henry helping.
Finished the instruction generation section.

@@
 \subsection{Instruction Generation}
+
 We model the input to our simulator as a stream of integer and floatpoint instructions.
@@
 \end{table*}
+
 Registers are generated in bit representation as 4 bits each totalling 12 bits for the s1, s2 and d registers in each instruction.
@@
 This history buffer is implemented as a rotating bit register from which prior destination registers can be indexed.
+
 Each instruction also features a program counter associated with the instruction.
@@
 Simply starting the PC of the first instruction in the stream at 0 and incrementing by 4 each instruction guarantees uniqueness among PCs and serves the purposes of an instruction tag.
+
 The aforementioned instruction fields have multiple modes of generation.
@@
 The instruction profile features: 1) a percent rate of occurrence for each instruction type, 2) a rate corresponding to the interdependence of registers between instructions, and 3) an overall instruction count.
+
 The instruction generator is currently capable of generating four distinct instruction streams corresponding to different instruction workloads.
@@
 This register dependence rate was selected arbitrarily to act as a starting point for modeling data dependencies between instructions.
+
 The remaining three instruction streams are modeled after SPEC-FP benchmarks.
@@
 Each of the SPEC-FP benchmarks feature a 6% register forwarding frequency as is reported by [XX paper ref].
+
 For the purposes of our power comparisons, each of the above instruction workloads are normalized to generate 10000 instructions.
@@
 The specific values used for each workload are listed in Table XX.
-
+\begin{table*}
+\centering
+\begin{tabular}{|c|c|c|c|c|} \hline
+ins{\char'137}profile&default{\char'137}profile&173{\char'056}applu&191{\char'056}fma3d&183{\char'056}equake\\ \hline
+loadProc&0.10&0.32&0.12&0.37\\ \hline
+storeProc&0.10&0.12&0.24&0.06\\ \hline
+addProc&0.10&0.01&0.15&0.08\\ \hline
+subProc&0.10&0.01&0.15&0.08\\ \hline
+multProc&0.10&0.01&0.15&0.08\\ \hline
+divProc&0.10&0.01&0.15&0.08\\ \hline
+faddProc&0.10&0.13&0.01&0.07\\ \hline
+fsubProc&0.10&0.13&0.01&0.06\\ \hline
+fmultProc&0.10&0.13&0.01&0.06\\ \hline
+fdivProc&0.10&0.13&0.01&0.06\\ \hline
+forwardingRate&0.10&0.06&0.06&0.06\\ \hline
+instCount&10000&10000&10000&10000\\ \hline\end{tabular}
+\caption{Benchmark Instruction Profiles showing percent rate of occurrence of each instruction type.}
+\end{table*}
 \subsection{Model Implementation}

### 2011-06-07 paper/paper.tex
Added more paper sections.

@@
 
-We model the input to our simulator as a stream of integer and floatpoint instructions.
-The instructions are represented at the bit-level and model MIPS instructions as closely as possible.
-Each instruction consists of 1) Instruction type, 2) Source and destination registers 3) a PC.
-Table XX indicates the types of instructions modeled.
-Each instruction type also has a predetermined cycle execution time designating the number of cycles that must elapse between instruction issue and the completion of the instruction.
-As modeling functional units goes beyond the scope of this simulator, instructions are flagged as complete after the listed number of cycles has elapsed.
-The actual cycle counts are the same as those modeled in MIT's Graphite many-core processor simulator[REFXX], with the exception of LOAD and STORE.
-The cycle counts corresponding to LOAD and STORE instructions are estimated as a predetermined 20 cycles each.
-This is simply an estimation as the actual cycle count is heavily dependent on cache architecture and benchmark specific cache behavior.
-Since both of these topics are beyond the scope of the simulator, we estimate the completion time of LOAD and STORE instructions as a set value substantially longer than the other more accurate instruction types.
 \begin{table*}
@@
 \caption{Modeled instruction types and their corresponding completion cycle counts.}
+\label{tab:instypetable}
 \end{table*}
+
+We model the input to our simulator as a stream of integer and floatpoint instructions.
+The instructions are represented at the bit-level and model MIPS instructions as closely as possible.
+Each instruction consists of 1) Instruction type, 2) Source and destination registers 3) a PC.
+Table ~\ref{tab:instypetable} indicates the types of instructions modeled.
+Each instruction type also has a predetermined cycle execution time designating the number of cycles that must elapse between instruction issue and the completion of the instruction.
+As modeling functional units goes beyond the scope of this simulator, instructions are flagged as complete after the listed number of cycles has elapsed.
+The actual cycle counts are the same as those modeled in MIT's Graphite many-core processor simulator\cite{graphite}, with the exception of LOAD and STORE.
+The cycle counts corresponding to LOAD and STORE instructions are estimated as a predetermined 20 cycles each.
+This is simply an estimation as the actual cycle count is heavily dependent on cache architecture and benchmark specific cache behavior.
+Since both of these topics are beyond the scope of the simulator, we estimate the completion time of LOAD and STORE instructions as a set value substantially longer than the other more accurate instruction types.
 
@@
 
-The remaining three instruction streams are modeled after SPEC-FP benchmarks.
+The remaining three instruction streams are modeled after SPEC-FP benchmarks \cite{rupnow}.
 The benchmarks were chosen to represent three major modes of operation:  floating point heavy execution, integer heavy execution, and mixed type execution.
@@
 As such it acts as a more representative general workload than the original default workload.
-Each of the SPEC-FP benchmarks feature a 6% register forwarding frequency as is reported by [XX paper ref].
+Each of the SPEC-FP benchmarks feature a 6{\char'045} register forwarding frequency as is reported by \cite{kucuk}.
 
-For the purposes of our power comparisons, each of the above instruction workloads are normalized to generate 10000 instructions.
-Such a large instruction window is ideal to allow buffers to fill and behavioral characteristics to develop.
-The specific values used for each workload are listed in Table XX.
 \begin{table*}
@@
 \caption{Benchmark Instruction Profiles showing percent rate of occurrence of each instruction type.}
+\label{tab:insprofile}
 \end{table*}
+
+For the purposes of our power comparisons, each of the above instruction workloads are normalized to generate 10000 instructions.
+Such a large instruction window is ideal to allow buffers to fill and behavioral characteristics to develop.
+The specific values used for each workload are listed in Table ~\ref{tab:insprofile}.
 \subsection{Model Implementation}
+The following section details the model implementatations in our simulator.
+The basic re-order buffer model we use as a base-line for comparison is a standard circular ROB model.
+Section ~\ref{sec:circ} discusses its implementation.
+We further model three optimizations for re-order buffers in an attempt to improve overall power usage: 1) Distributed (Cluster based) ROB, 2) Dynamically Sized ROB, 3) Reduced Output Port ROB (Retention Latch).
+Each of these optimized models are built on top of the underlying circular ROB model to provide as close a behavioral and power comparison as possible.
+Each of the optimizations, as well as the standard circular reorder buffer, are described in detail below.
 \subsubsection{Circular ROB}
-\subsubsection{Distributed ROB}
-\subsubsection{Dynamic ROB}
+\label{sec:circ}
+\subsubsection{Distributed (Cluster based) ROB}
+\subsubsection{Dynamically Sized ROB}
 \subsubsection{Reduced Output Port ROB (Retention Latch)}
@@
 %TODO add in bib 
-
+\begin{thebibliography}{------}
+\bibitem{graphite}
+	MIT Graphite Simulator
+	https://github.com/mit-carbon/Graphite/wiki
+\bibitem{kucuk}
+Gurhan Kucuk, Dmitry Ponomarev, and Kanad Ghose. 2002. Low-complexity reorder buffer architecture. In Proceedings
+of the 16th international conference on Supercomputing (ICS '02). ACM, New York, NY, USA, 57-66. DOI=10.
+1145/514191.514202 http://doi.acm.org/10.1145/514191.514202
+\bibitem{rupnow}
+Kyle Rupnow, Arun Rodrigues, Keith Underwood, and Katherine Compton.
+2006.
+Scientific applications vs.
+SPEC-FP: a comparison of program behavior.
+In Proceedings of the 20th annual international conference on Supercomputing (ICS '06).
+ACM, New York, NY, USA, 66-74.
+DOI=10.1145/1183401.1183413 http://doi.acm.org/10.1145/1183401.1183413
+\end{thebibliography}
 \end{document}

### 2011-06-07 paper/paper.tex
paper

--- 2011-06-07 paper/paper.tex
@@
 a ReOrder Buffer (ROB), to commit the registers in program order.  This 
-structure has grown to be monolithic numerous entries and read/write ports.
+structure has grown to be monolithic with numerous entries and read/write ports.
 There is no current methods to efficiently and equally compare various
-ROB advancements to improve power efficiency.  
+ROB advancements to improve power efficiency. We propose a loosely clycle
+accurate model that keeps track of the power usages of the ROB.  Using the 
+simulation tool, we are able to compare and contrast the affects of improvements
+made on the ROB.  This also allows for a first line test to validate proof of 
+concepts for future power efficient ROB designs. 
 \end{abstract}
 
-\terms{Theory}
+\terms{Power, Energy, Architecture}
 
-\keywords{ACM proceedings, \LaTeX, text tagging} % NOT required for Proceedings
+\keywords{Power efficiency, ROB, Simulation} % NOT required for Proceedings
 
@@
 BLAH BLAH BLAH INTRO WEEEE
+\section{Background}
+    An out of order execution system has two additional stages 
+to the lifetime of the instruction on top of the traditional pipeline 
+stages.  Each of which a small memory array to keep track of instuction flow.
+The first addtion, the instruction queue(IQ) which holds instructions that are 
+waiting to be executed.  Each entry will hold the instruction -- including 
+type, operand registers, and destination registers.  The second structure 
+holds the instructions that are in flight but not yet commited to the register
+file. Each entry will hold both PC of the instruction, as well as the destination
+register and the result of the excution unit.  There is also a flag that will
+mark the result of the entry as valid.
+
+Replacing the traditional instruction dispatch stage, An placeholder will be made for the 
+instruction in ReOrder Buffer (ROB), and the instructions will be placed
+in the instruction queue(IQ). In this stage instructions will first wait for their operands if not already
+available.  When all operands are available, the instruction is ready to 
+be executed so wait until its appropriate execution unit is free.  At this 
+point the instruction exits the IQ and begins execution.
+
+Upon execution completion, the output value is stored in the ROB with its 
+corresponding entry and that entry is marked as valid or completed with a 
+1 bit flag.  An entry is commited from the ROB to the architectrual register
+file (ARF) when that entry is the oldest (lowest PC) and its result has been 
+marked as valid. 
+\subsection{Circular ROB}
+The circular ROB is 
+\subsection{Latch ROB}
+\subsection{Dynamic ROB}
+\subsection{Distributed ROB}
+
 \section{Project Implementation}
@@
 The specific values used for each workload are listed in Table XX.
-
+\begin{table*}
+\centering
+\begin{tabular}{|c|c|c|c|c|} \hline
+ins\_profile&default\_profile&173.applu&191.fma3d&183.equake\\ \hline
+loadProc&0.10&0.32&0.12&0.37\\ \hline
+storeProc&0.10&0.12&0.24&0.06\\ \hline
+addProc&0.10&0.01&0.15&0.08\\ \hline
+subProc&0.10&0.01&0.15&0.08\\ \hline
+multProc&0.10&0.01&0.15&0.08\\ \hline
+divProc&0.10&0.01&0.15&0.08\\ \hline
+faddProc&0.10&0.13&0.01&0.07\\ \hline
+fsubProc&0.10&0.13&0.01&0.06\\ \hline
+fmultProc&0.10&0.13&0.01&0.06\\ \hline
+fdivProc&0.10&0.13&0.01&0.06\\ \hline
+forwardingRate&0.10&0.06&0.06&0.06\\ \hline
+instCount&10000&10000&10000&10000\\ \hline\end{tabular}
+\caption{Benchmark Instruction Profiles}
+\end{table*}
 \subsection{Model Implementation}

### 2011-06-07 src/main.cpp
ROB_Dyn: Add optmized version. Implement first optimization: aggressive downsizing.
* main.cpp: (main) Parse dynamic-opt option, for optimized dynamic
ROB.
ROB is now 16 partitions, rather than 8 before.
(generate_instruction_stream) 100k instructions generated rather
than 10k.
Insturction generation messages now require -d option to display.
* rob_dyn.cpp: (constructor) Initially m_opt member to denote
optimized version.
(rob_process) Resize by more than 1 partiion if m_opt is true.

@@ -16,12 +16,13 @@ enum {
   MODE_BASE,
   MODE_CIRC,
   MODE_DYN,
+  MODE_DYNOPT,
   MODE_DIST,
   MODE_LATCH
 };
 
 void help ();
-ins_t *generate_instruction_stream(int benchmark);
+ins_t *generate_instruction_stream(int benchmark, int print_flags);
 int generate_type(ins_profile input, int gen);
 uint32_t generate_register(uint32_t history, float forwardRate);
 
@@ -35,7 +36,7 @@ int main (int argc, char *argv[])
   ROB *rob;
   ins_t *instructions;
   int rob_size = 128;
-  int print_flags = 1;
+  int print_flags = 0;
 
   while ((opt = getopt (argc, argv, "dhm:b:")) != -1)
     {
@@ -57,6 +58,8 @@ int main (int argc, char *argv[])
             mode = MODE_CIRC;
           else if (strcmp (optarg, "dynamic") == 0)
             mode = MODE_DYN;
+          else if (strcmp (optarg, "dynamic-opt") == 0)
+            mode = MODE_DYNOPT;
           else if (strcmp (optarg, "dist") == 0)
             mode = MODE_DIST;
           else if (strcmp (optarg, "latch") == 0)
@@ -89,7 +92,7 @@ int main (int argc, char *argv[])
   srand (10);
 
   // For now, generate arbitrary mix of instructions.
-  instructions = generate_instruction_stream(bench);
+  instructions = generate_instruction_stream(bench, print_flags);
 
   switch (mode)
     {
@@ -105,7 +108,13 @@ int main (int argc, char *argv[])
 
     case MODE_DYN:
       cout << "Dynamic" << endl;
-      rob = new ROB_Dyn (rob_size, 3, 1, 256, 8, print_flags);
+      // Overflow = 256, partitions = 8
+      rob = new ROB_Dyn (rob_size, 3, 1, 256, 16, print_flags, false);
+      break;
+
+    case MODE_DYNOPT:
+      cout << "Dynamic Optimized" << endl;
+      rob = new ROB_Dyn (rob_size, 3, 1, 256, 16, print_flags, true);
       break;
 
     case MODE_DIST:
@@ -150,7 +159,7 @@ void help ()
   cout << "  3          183.equake SPEC-FP Seismic Wave Propagation benchmark.  Mix of FP and INT ops." << endl;
 }
 
-ins_t *generate_instruction_stream(int benchmark) {
[…]

### 2011-06-07 src/rob_dyn.cpp

@@ -5,10 +5,10 @@ using namespace std;
 
 extern int ins_cyc[];
 
-ROB_Dyn::ROB_Dyn (int s, int in, int fn, int overflow, int parts, int pflags)
+ROB_Dyn::ROB_Dyn (int s, int in, int fn, int overflow, int parts, int pflags, bool opt)
   : ROB_Circ (s, in, fn, pflags)
 {
-  m_update_period = 512;
+  m_update_period = 2048;
   m_sample_period = 64;
   m_samples_taken = 0;
   m_current_size = s;
@@ -36,6 +36,8 @@ ROB_Dyn::ROB_Dyn (int s, int in, int fn, int overflow, int parts, int pflags)
           m_prev_parts[i].buf[j] = empty_entry;
         }
     }
+
+  m_opt = opt;
 
   /*
   int tmp = 0;
@@ -128,15 +130,20 @@ void ROB_Dyn::dyn_process (int cycles)
       //cout << "  Active: " << m_active_size;
       //cout << "  overflow " << m_overflow_cnt << endl;
 
-      if (m_current_size > m_active_size && m_current_size - m_active_size > m_part_size)
-        {
-          // Shrink buffer.
-          dyn_shrink (1);
-        }
       if (m_overflow_cnt > m_overflow_thresh)
         {
           // Grow buffer.
           dyn_grow (1);
+        }
+      else if (m_current_size > m_active_size && m_current_size - m_active_size > m_part_size)
+        {
+
+          int n_shrink = 1;
+          if (m_opt)
+            n_shrink = (m_current_size - m_active_size) / m_part_size;
+
+          // Shrink buffer.
+          dyn_shrink (n_shrink);
         }
 
       m_overflow_cnt = 0;

### 2011-06-07 src/rob_dyn.h

@@ -21,7 +21,7 @@ typedef struct partition {
 
 class ROB_Dyn : public ROB_Circ {
  public:
-  ROB_Dyn  (int s, int in, int fn, int overflow, int parts, int pflags);
+  ROB_Dyn  (int s, int in, int fn, int overflow, int parts, int pflags, bool opt);
   virtual ~ROB_Dyn ();
   virtual void run (ins_t instructions[]);
 
@@ -54,6 +54,7 @@ class ROB_Dyn : public ROB_Circ {
   uint32_t m_overflow_cnt;
   uint32_t m_overflow_thresh;
   uint32_t m_partition_cycles;
+  bool m_opt;
 };
 
 #endif // ROB_DYN_H

### 2011-06-08 src/main.cpp
Delayed partition enabling implemented.
* src/main.cpp: Overflow threshold upped to 512.
* src/rob_dyn.cpp: (run) m_overflow_cnt increments by the number
of instructions stalled, rather than 1 each time.
(dyn_process) Separated buffer enable/disable for regular and
optimized dynamic ROB.
Comparator usage marked for checking pointer location conditions.
(tail_incr_update) New function to support delayed partition
enabling. In incrementing tail pointer, enable partition if marked
TO_ENABLE.

@@ -109,12 +109,12 @@ int main (int argc, char *argv[])
     case MODE_DYN:
       cout << "Dynamic" << endl;
       // Overflow = 256, partitions = 8
-      rob = new ROB_Dyn (rob_size, 3, 1, 256, 16, print_flags, false);
+      rob = new ROB_Dyn (rob_size, 3, 1, 512, 16, print_flags, false);
       break;
 
     case MODE_DYNOPT:
       cout << "Dynamic Optimized" << endl;
-      rob = new ROB_Dyn (rob_size, 3, 1, 256, 16, print_flags, true);
+      rob = new ROB_Dyn (rob_size, 3, 1, 512, 16, print_flags, true);
       break;
 
     case MODE_DIST:

### 2011-06-08 src/rob_dyn.cpp

@@ -84,7 +84,7 @@ void ROB_Dyn::run (ins_t ins[])
       // 1) If less than m_n instructions were processed
       // 2) Buffer is full
       if (old_ins_num + (int) m_n > ins_num && (m_head == m_tail && !m_empty))
-        m_overflow_cnt++;
+        m_overflow_cnt += (uint32_t) (old_ins_num + (int) m_n - ins_num);
 
       cycles++;
       dyn_process (cycles);
@@ -130,12 +130,7 @@ void ROB_Dyn::dyn_process (int cycles)
       //cout << "  Active: " << m_active_size;
       //cout << "  overflow " << m_overflow_cnt << endl;
 
-      if (m_overflow_cnt > m_overflow_thresh)
-        {
-          // Grow buffer.
-          dyn_grow (1);
-        }
-      else if (m_current_size > m_active_size && m_current_size - m_active_size > m_part_size)
+      if (m_current_size > m_active_size && m_current_size - m_active_size > m_part_size)
         {
 
           int n_shrink = 1;
@@ -144,6 +139,13 @@ void ROB_Dyn::dyn_process (int cycles)
 
           // Shrink buffer.
           dyn_shrink (n_shrink);
+          //cerr << "Shrink " << n_shrink << endl;
+        }
+      if (m_overflow_cnt > m_overflow_thresh)
+        {
+          // Grow buffer.
+          dyn_grow (1);
+          //cerr << "Grow" << endl;
         }
 
       m_overflow_cnt = 0;
@@ -153,29 +155,51 @@ void ROB_Dyn::dyn_process (int cycles)
   // If possible to turn buffer marked TO_DISABLE off, or TO_ENABLE on, do so.
   for (int i = 0; i < m_nparts; i++)
     {
-      if (m_parts[i].state == TO_DISABLE || m_parts[i].state == TO_ENABLE)
+      // If optmized, we do just-in-time allocation, so we only handle disabling here.
+      if (m_opt && m_parts[i].state == TO_DISABLE)
         {
           uint32_t part_start = i * m_part_size;
-          uint32_t part_end = part_start + m_part_size;
+          //uint32_t part_end = part_start + m_part_size;
 
           // State change if section is not in middle of used region.
           // 1) Used region is before partition.
           // 2) Used region is after partition.
           // 3) Used region starts after partition and ends before partition
           //    after wrap around.
[…]

### 2011-06-08 src/rob_dyn.h

@@ -36,6 +36,7 @@ class ROB_Dyn : public ROB_Circ {
   virtual entry_t *get_entry (uint32_t ptr);
   virtual entry_t *get_prev_buf_entry (uint32_t ptr);
   virtual void error_diag ();
+  void write_entry (entry_t *entry, ins_t ins);
 
   void pre_cycle_power_snapshot ();
   void post_cycle_power_tabulation ();
@@ -45,6 +46,7 @@ class ROB_Dyn : public ROB_Circ {
   void dyn_process (int cycles);
   void dyn_shrink (int parts);
   void dyn_grow (int parts);
+  uint32_t tail_incr_update (uint32_t ptr);
 
   uint32_t m_update_period;
   uint32_t m_sample_period;

### 2011-06-08 src/main.cpp
Dynamic ROB optimizations complete.
Note: circular to dynamic power savings is minimal, investigate.

@@ -28,8 +28,6 @@ uint32_t generate_register(uint32_t history, float forwardRate);
 
 int main (int argc, char *argv[])
 {
-  //srand ( time(NULL) );
-  srand(2);
   char opt;
   int mode = MODE_BASE;
   int bench = 0;
@@ -109,12 +107,12 @@ int main (int argc, char *argv[])
     case MODE_DYN:
       cout << "Dynamic" << endl;
       // Overflow = 256, partitions = 8
-      rob = new ROB_Dyn (rob_size, 3, 1, 512, 16, print_flags, false);
+      rob = new ROB_Dyn (rob_size, 3, 1, 256, 16, print_flags, false);
       break;
 
     case MODE_DYNOPT:
       cout << "Dynamic Optimized" << endl;
-      rob = new ROB_Dyn (rob_size, 3, 1, 512, 16, print_flags, true);
+      rob = new ROB_Dyn (rob_size, 3, 1, 256, 16, print_flags, true);
       break;
 
     case MODE_DIST:

### 2011-06-08 src/rob_circ.cpp

@@ -46,7 +46,7 @@ void ROB_Circ::run (ins_t ins[])
 
       // TODO: Count the read ports being driven for operands.
       // Get percentage this happens from Henry.
-
+ 
       post_cycle_power_tabulation ();
       update_power_totals ();
 
@@ -202,9 +202,9 @@ uint32_t ROB_Circ::tail_incr (uint32_t tail_ptr)
 uint32_t ROB_Circ::getCyclesToCompletion(uint32_t reg) {
   uint32_t i = m_head;
   if (!m_empty)
-    do { 
+    do {
       entry_t *entry = get_entry(i);
-      if(entry->reg_id == reg) 
+      if(entry->reg_id == reg)
         return entry->cycles;
       i = ((i+1)%m_size);
     } while(i != m_tail);

### 2011-06-08 src/rob_dyn.cpp

@@ -9,7 +9,7 @@ ROB_Dyn::ROB_Dyn (int s, int in, int fn, int overflow, int parts, int pflags, bo
   : ROB_Circ (s, in, fn, pflags)
 {
   m_update_period = 2048;
-  m_sample_period = 64;
+  m_sample_period = 128;
   m_samples_taken = 0;
   m_current_size = s;
   m_active_size = 0;
@@ -71,6 +71,33 @@ void ROB_Dyn::run (ins_t ins[])
     {
       pre_cycle_power_snapshot ();
 
+      // Check for sanity: head and tail pointers must be in an enabled partition.
+      int head_part = m_head / m_part_size;
+      int tail_part = m_tail / m_part_size;
+
+      if (m_parts[head_part].state == DISABLED
+          || m_parts[head_part].state == TO_ENABLE
+          || m_parts[tail_part].state == DISABLED
+          || m_parts[tail_part].state == TO_ENABLE)
+        {
+          cerr << "CORRUPT BUFFER h:" << m_head << "  t:" << m_tail;
+
+          for (int i = 0; i < m_nparts; i++)
+            {
+              if (m_parts[i].state == DISABLED)
+                cerr << "  [ D]";
+              else if (m_parts[i].state == TO_DISABLE)
+                cerr << "  [TD]";
+              else if (m_parts[i].state == TO_ENABLE)
+                cerr << "  [TE]";
+              else if (m_parts[i].state == ENABLED)
+                cerr << "  [ E]";
+            }
+          cerr << endl;
+        }
+
+
+
       bool old_empty = m_empty;
       int old_ins_num = ins_num;
       update_entries ();
@@ -84,7 +111,7 @@ void ROB_Dyn::run (ins_t ins[])
       // 1) If less than m_n instructions were processed
       // 2) Buffer is full
       if (old_ins_num + (int) m_n > ins_num && (m_head == m_tail && !m_empty))
-        m_overflow_cnt += (uint32_t) (old_ins_num + (int) m_n - ins_num);
+        m_overflow_cnt++;
 
       cycles++;
       dyn_process (cycles);
@@ -126,9 +153,9 @@ void ROB_Dyn::dyn_process (int cycles)
   // Update buffer size.
   if (cycles % m_update_period == 0)
     {
-      //cout << "Cur: " << m_current_size;
-      //cout << "  Active: " << m_active_size;
-      //cout << "  overflow " << m_overflow_cnt << endl;
[…]

### 2011-06-08 paper/paper.tex
paper bg

@@
 There is no current methods to efficiently and equally compare various
-ROB advancements to improve power efficiency. We propose a loosely clycle
+ROB advancements to improve power efficiency. We propose a loosely cycle
 accurate model that keeps track of the power usages of the ROB.  Using the 
@@
 to the lifetime of the instruction on top of the traditional pipeline 
-stages.  Each of which a small memory array to keep track of instuction flow.
-The first addtion, the instruction queue(IQ) which holds instructions that are 
+stages.  Each of which a small memory array to keep track of instruction flow.
+The first addition, the instruction queue(IQ) which holds instructions that are 
 waiting to be executed.  Each entry will hold the instruction -- including 
 type, operand registers, and destination registers.  The second structure 
-holds the instructions that are in flight but not yet commited to the register
+holds the instructions that are in flight but not yet committed to the register
 file. Each entry will hold both PC of the instruction, as well as the destination
-register and the result of the excution unit.  There is also a flag that will
+register and the result of the execution unit.  There is also a flag that will
 mark the result of the entry as valid.
@@
 corresponding entry and that entry is marked as valid or completed with a 
-1 bit flag.  An entry is commited from the ROB to the architectrual register
+1 bit flag.  An entry is committed from the ROB to the architectural register
 file (ARF) when that entry is the oldest (lowest PC) and its result has been 
@@
 \subsection{Circular ROB}
-The circular ROB is 
+The circular ROB is one of the most basic implementations of the ROB.  Simply using
+a circular buffer as the structure to hold each entry.  Since the ROB is basically
+just memory, there are two special slots for to keep track of the head and tail
+pointer to create the circular buffer.  As entries are written into the ROB, the
+entry is written at the tail pointer and then the tail pointer is incremented.
+Likewise when an entries is written to the ARF, the head pointer is incremented.
+The circular buffer method is widely used since it inherently keeps track of the 
+oldest instruction (which is at the head pointer), as well as which entry slots
+are used or empty.   However it may not be very power efficient if there is a large
+number of empty slots at any given point of the execution.    
+\subsection{Dynamic ROB}
+
+\subsection{Distributed ROB}
+
 \subsection{Latch ROB}
-\subsection{Dynamic ROB}
-\subsection{Distributed ROB}
+In addition to large power requirement to maintain the entries of the ROB, another 
+large power consumer is the read ports of the ROB.  The read port of the ROB is used
+to forward the results of an entry to the dispatch unit (DU) to use as an operand.
+All though forwarding from the ROB occurs about 6\%\cite{kucuk} of all instructions 
+in an N way super scalar processor there are 2N ports (to accommodate FP
+ instructions) which add unnecessary complexity and power consumption for a small 
+percentage of instructions.  It has been proposed by eliminating the read ports 
+in the ROB all together, we can reduce energy consumption\cite{kucuk}.  However
+without read ports, a small window appears where the data cannot be accessed,
+starting from when the result is placed in to the ROB until the ROB commits that 
+entry to the ARF.  To help mitigate stalls that might happen due to this hole, 
+we introduce retention latches.
 
@@
 
-Registers are generated in bit representation as 4 bits each totalling 12 bits for the s1, s2 and d registers in each instruction.
+Registers are generated in bit representation as 4 bits each totaling 12 bits for the s1, s2 and d registers in each instruction.
 Currently, our representation of instructions make no distinction between zero, 1, and 2 operand instructions.
@@
 DOI=10.1145/1183401.1183413 http://doi.acm.org/10.1145/1183401.1183413
+
 \end{thebibliography}

### 2011-06-08 paper/paper.tex
paper bg

@@
 entry to the ARF.  To help mitigate stalls that might happen due to this hole, 
-we introduce retention latches.
+we introduce retention latches.   Retention latches are small memory elements and 
+we use them to cache the most recent results that have been written into the ROB.
+This way instructions are still able to access the most recently produced results
+and not stall.  However if the operand needed is not in the ROB, the stream will 
+stall until a commit is made to the ARF, and that result value is forwarded to 
+the dispatch unit immediately.
 

### 2011-06-08 paper/paper.tex
Paper sections for circular ROB and dynamic + dynamic optimized ROBs done.

--- 2011-06-08 paper/paper.tex
@@
 marked as valid.
-\subsection{Circular ROB}
-The circular ROB is 
-\subsection{Latch ROB}
-\subsection{Dynamic ROB}
-\subsection{Distributed ROB}
 
@@
 The first instruction stream is marked as the default instruction stream.
-It features equal percentages of each type of instruction as well as a 10% register dependence rate.
+It features equal percentages of each type of instruction as well as a 10\% register dependence rate.
 This register dependence rate was selected arbitrarily to act as a starting point for modeling data dependencies between instructions.
@@
 forwardingRate&0.10&0.06&0.06&0.06\\ \hline
-instCount&10000&10000&10000&10000\\ \hline\end{tabular}
+instCount&100000&100000&100000&100000\\ \hline\end{tabular}
 \caption{Benchmark Instruction Profiles showing percent rate of occurrence of each instruction type.}
@@
 \label{sec:circ}
+The circular ROB is the basic current implementation of the reorder buffer.
+It is implemented as a circular FIFO buffer, and the beginning and end of the used region are tracked by the head and tail pointers, respectively.
+When an instruction is issued, space is reserved for the instruction's result at the tail pointer, and the tail pointer is incremented.
+When the instruction completes, the result is inserted into the reorder buffer at the reserved space.
+The result is finally committed to the architecture register file when the head pointer passes through it; each cycle, results are committed from the head pointer and the head pointer is incremented repeatedly, until either an unfinished instruction result slot is reached or all write ports are used up.
+
+The circular ROB is considered full when the tail pointer reaches the head pointer, as the head pointer denotes the start of the in-use region and no more space can be reserved from the tail pointer.
+Conversely, then buffer is empty when the head pointer catches up to the tail pointer, as every instruction in the buffer has now been committed to the ARF.
+
 \subsubsection{Distributed (Cluster based) ROB}
+
 \subsubsection{Dynamically Sized ROB}
+The dynamically sized ROB is an optimization on the circular ROB where the buffer is divided into partitions, and those partitions deemed unnecessary are powered off to conserve energy.
+Our implementation of the basic dynamic ROB is behaviorally the same as that detailed in \cite{kucuk2}.
+
+Downsizing the ROB is based on a periodic sampling of the buffer state.
+A number of cycles, {\it update\_period}, is defined, and every {\it update\_period} cycles the size of the ROB is updated.
+Every {\it sample\_period} cycles, the number of active entries is recorded, and the quantity {\it active\_size} is recorded as the average of these samples between each {\it update\_period}.
+The current maximum capacity of the reorder buffer is defined as the {\it current\_size}, and the size of a single ROB partition is defined as {\it partition\_size}.
+After each {\it update\_period} cycles, the ROB checks how full the buffer is, and if {\it current\_size} - {\it active\_size} > {\it partition\_size}, then a single partition is turned off.
+
+Increasing the size of the ROB is dependent on how often the buffer fills up.
+A value, defined as the {\it overflow\_count} is incremented each cycle that instruction issue is stalled because the ROB is full.
+An {\it overflow\_threshold} is defined, and if during an {\it update\_period}, {\it overflow\_count} > {\it overflow\_threshold}, then a partition is turned on.
+
+A conservative approach of deallocating strictly from one end of the buffer, and allocating strictly towards that end is followed.
+For example, if the ROB consisted of four partitions, downsizing the buffer would turn off the fourth, and then the third.
+At this point, if we were to increase the size of the buffer, the third partition would be turned on first, and the fourth partition next.
+
+When the ROB's size is updated, partitions are not immediately turned off, since the partition could coincide with the active region of the buffer.
+If deallocating, this would disable a part of the active region, making instructions in the region impossible to complete.
+If allocating, this would insert an empty segment right into the middle of the FIFO queue, logically breaking the buffer.
+It is only safe to turn a partition off/on when the active region is not wrapped around (that is, the head pointer is less than the tail pointer), and the tail pointer comes before the first index into the partition to turn on/off.
+\cite{kucuk3} details a worst case scenario of deallocating, where the active region is wrapped (head is greater than tail), and both pointers lie in the region to be turned off.
+In this case, the head pointer must first clear the entries already in the buffer.
+Since the tail pointer is already in the partition, it will continue to write into the partition, and the head pointer must come around a second time to clear out these new entries.
+Only then can the partition be safely deallocated, after the head pointer has wrapped around to zero twice.
+
+\subsubsection{Optimized Dynamic ROB}
+We implement an optimized version of the dynamic ROB detailed in \cite{kucuk2} and \cite{kucuk3}, opting to be more aggressive with turning off partitions in order to save power, and to improve the partition allocation/deallocation scheme.
+
+When downsizing the regular dynamic ROB, a single partition is turned off if {\it current\_size} - {\it active\_size} > {\it partition\_size}.
+In an effort to minimize the buffer size and thus the power consumed, we choose to downsize by
+
+\begin{center}
+$\lfloor \frac{current\_size - active\_size}{partition\_size} \rfloor$
+\end{center}
+
+Thus if much less than the capacity of the buffer is active, several partitions may be turned off at once.
+
+To more efficiently resize the buffer, we found that choosing the first valid partition immediately after the one housing the tail pointer is optimal.
+For example, if we are allocating a partition in a four partition buffer, and the tail pointer resides in partition 3, we first check if partition 4 can be turned on, then partition 1, then partition 2.
+
+In allocating the buffer, the reasoning is as follows.
+We aim to enable a disabled partition.
+It is impossible for the head pointer to lie in the disabled partition, thus there is no need to consider the head pointer.
+To maintain the logical consistency of the buffer, we cannot enable a partition that would divide the active region of the buffer.
+Thus, we must allocate a partition that comes after the tail pointer, and if we wish the buffer size to grow as soon as possible, the partition to be allocated must be the first disabled partition after the tail pointer.
+
+In deallocating the buffer, turning off the first partition after the tail pointer is also optimal.
+Being after the tail pointer, once we mark the partition for disabling we can prevent the tail pointer from writing into it.
+This prevents the worst case scenario where the head pointer must wrap around 0 twice to disable a partition.
+The next enabled partition immediately after the tail pointer is also the one the active region will leave first (if the region is in the partition; if not the partition is already empty and ready to deactivate), thus it will empty out first.
+If we are to maximize the power savings, we must disable the first partition that is available to be disabled, thus the first active partition after the tail is the optimal choice.
+
+One last optmization we implemented was just-in-time allocation of partitions.
+Normally, when the tail pointer increments, if a partition is not active the tail pointer simply skips it.
+In our optimization, before skipping the tail pointer will check if the disabled partition is marked to be enabled.
+If so, instead of skipping over it, the partition will be turned on and the tail pointer will write into it.
+This ensures that each partition to be turned on is enabled at the last possible moment, saving the energy that would otherwise be consumed by an empty partition waiting to be written into.
+
 \subsubsection{Reduced Output Port ROB (Retention Latch)}
+
 \subsection{Power Tabulation Implementation}
@@
 1145/514191.514202 http://doi.acm.org/10.1145/514191.514202
+\bibitem{kucuk2}
+Gurhan Kucuk, Dmitry Ponomarev, and Kanad Ghose.
+2002.
+Energy-Efficient Design of the Reorder Buffer.
+In Proceedings of the International Workshop on Power and Timing Modeling, Optimization, and Simulation, September 2002.
+\bibitem{kucuk3}
+Gorhan Kucuk, Dmitry Ponomarev, and Kanad Ghose.
+2001.
+Reducing Power Requirements of Instruction Scheduling Through Dynamic Allocation of Multiple Datapath Resources.
+In Proceedings of the International Symposium on Microarchitecture, December 2001.
 \bibitem{rupnow}

### 2011-06-08 src/main.cpp

@@ -33,7 +33,7 @@ int main (int argc, char *argv[])
   int bench = 0;
   ROB *rob;
   ins_t *instructions;
-  int rob_size = 128;
+  int rob_size = 256;
   int print_flags = 0;
 
   while ((opt = getopt (argc, argv, "dhm:b:")) != -1)
@@ -106,13 +106,13 @@ int main (int argc, char *argv[])
 
     case MODE_DYN:
       cout << "Dynamic" << endl;
-      // Overflow = 256, partitions = 8
-      rob = new ROB_Dyn (rob_size, 3, 1, 256, 16, print_flags, false);
+      // Overflow = 512, partitions = 16
+      rob = new ROB_Dyn (rob_size, 3, 1, 512, 16, print_flags, false);
       break;
 
     case MODE_DYNOPT:
       cout << "Dynamic Optimized" << endl;
-      rob = new ROB_Dyn (rob_size, 3, 1, 256, 16, print_flags, true);
+      rob = new ROB_Dyn (rob_size, 3, 1, 512, 16, print_flags, true);
       break;
 
     case MODE_DIST:

### 2011-06-08 paper/paper.tex

@@
 to the lifetime of the instruction on top of the traditional pipeline 
-stages.  Each of which a small memory array to keep track of instuction flow.
-The first addtion, the instruction queue(IQ) which holds instructions that are
+stages.  Each of which a small memory array to keep track of instruction flow.
+The first addition, the instruction queue(IQ) which holds instructions that are 
 waiting to be executed.  Each entry will hold the instruction -- including 
 type, operand registers, and destination registers.  The second structure 
-holds the instructions that are in flight but not yet commited to the register
+holds the instructions that are in flight but not yet committed to the register
 file. Each entry will hold both PC of the instruction, as well as the destination
-register and the result of the excution unit.  There is also a flag that will
+register and the result of the execution unit.  There is also a flag that will
 mark the result of the entry as valid.
@@
 corresponding entry and that entry is marked as valid or completed with a 
-1 bit flag.  An entry is commited from the ROB to the architectrual register
+1 bit flag.  An entry is committed from the ROB to the architectural register
 file (ARF) when that entry is the oldest (lowest PC) and its result has been 
 marked as valid. 
+\subsection{Circular ROB}
+The circular ROB is one of the most basic implementations of the ROB.  Simply using
+a circular buffer as the structure to hold each entry.  Since the ROB is basically
+just memory, there are two special slots for to keep track of the head and tail
+pointer to create the circular buffer.  As entries are written into the ROB, the
+entry is written at the tail pointer and then the tail pointer is incremented.
+Likewise when an entries is written to the ARF, the head pointer is incremented.
+The circular buffer method is widely used since it inherently keeps track of the 
+oldest instruction (which is at the head pointer), as well as which entry slots
+are used or empty.   However it may not be very power efficient if there is a large
+number of empty slots at any given point of the execution.    
+\subsection{Dynamic ROB}
+
+\subsection{Distributed ROB}
+
+\subsection{Latch ROB}
+In addition to large power requirement to maintain the entries of the ROB, another 
+large power consumer is the read ports of the ROB.  The read port of the ROB is used
+to forward the results of an entry to the dispatch unit (DU) to use as an operand.
+All though forwarding from the ROB occurs about 6\%\cite{kucuk} of all instructions 
+in an N way super scalar processor there are 2N ports (to accommodate FP
+ instructions) which add unnecessary complexity and power consumption for a small 
+percentage of instructions.  It has been proposed by eliminating the read ports 
+in the ROB all together, we can reduce energy consumption\cite{kucuk}.  However
+without read ports, a small window appears where the data cannot be accessed,
+starting from when the result is placed in to the ROB until the ROB commits that 
+entry to the ARF.  To help mitigate stalls that might happen due to this hole, 
+we introduce retention latches.   Retention latches are small memory elements and 
+we use them to cache the most recent results that have been written into the ROB.
+This way instructions are still able to access the most recently produced results
+and not stall.  However if the operand needed is not in the ROB, the stream will 
+stall until a commit is made to the ARF, and that result value is forwarded to 
+the dispatch unit immediately.
 

### 2011-06-08 paper/paper.tex
Background info on dynamic buffer.

@@
 \subsection{Dynamic ROB}
+The dynamic ROB aims to reduce power by turning off portions of the buffer when they
+are not needed, and turning them back on when they become necessary. By partitioning
+the buffer into what is essentially smaller buffers, the complexity of each partition
+is reduced. Further, the ROB will then be custom sized to need, thus reducing energy
+wasted powering large stretches of empty entries. Due to the large percentage of energy
+consumption stemming from leakage power, turning off partitions contribute significant
+power savings over the base circular ROB.
 

### 2011-06-08 paper/paper.tex
Completed my sections.  I just need to insert an image when I ask henry how, and find a reference paper I used.

--- 2011-06-08 paper/paper.tex
@@
 marked as valid. 
+\subsection{Circular ROB}
+The circular ROB is 
+\subsection{Distributed ROB}
+An accepted optimization for improving the power consumption of an ROB unit 
+is to break up the ROB buffer into clusters REF[XXX].  In this organization, the in-flight 
+instruction information is distributed across all of the units and then re-assembled 
+back into program order after the fact.  The primary advantage of this approach 
+is that it dramatically lessens the size of the individual ROB units.  This reduces overall
+power by reducing the bit line width as well as unit capacitance.
+
+...IMAGE FROM OUR SLIDES...
+
+The overhead associated with this scheme is that the addition of a separate FIFO unit
+is required to maintain a global state of instruction order between the separate ROB units.
+This central FIFO unit is then referenced upon instruction re-order to commit the instructions
+in program order.  The contents of the FIFO unit are fairly simple and incur reasonable hardware
+overhead.  Each entry needs only contain the PC of the instruction (or some other unique identifier)
+and the corresponding ROB unit that houses the placeholder of the instruction as it is in-flight. 
+The total ROB bit storage count remains the same from the non-distributed model to the distributed model
+as the overall size of the ROB remains the same.  However, the inclusion of the central FIFO queue 
+induces some overhead bits that must be considered when analyzing the overall effectiveness of the 
+model.
+
+The distribution or clustering of the ROB unit also inherently induces a minimal performance overhead 
+in particular cases.  By breaking the ROB into smaller pieces and distributing the instruction 
+data across the unit, the placement of instructions becomes a performance factor.  Typically, a 
+distributed ROB model features only a single write port per ROB cluster.  As a result only one operand 
+can be forwarded at a time from the ROB.  In the rare case that an instruction requires two operands be 
+forwarded from the same ROB cluster a delay of one cycle is imposed while the operands are forwarded 
+sequentially.  It is reported, however, that this happens very rarely in practice and induces less than 1\% 
+performance degradation REF[XXX].
+\subsection{Dynamic ROB}
+\subsection{Latch ROB}
 
@@
 The first instruction stream is marked as the default instruction stream.
-It features equal percentages of each type of instruction as well as a 10\% register dependence rate.
+It features equal percentages of each type of instruction as well as a 10% register dependence rate.
 This register dependence rate was selected arbitrarily to act as a starting point for modeling data dependencies between instructions.
@@
 forwardingRate&0.10&0.06&0.06&0.06\\ \hline
-instCount&100000&100000&100000&100000\\ \hline\end{tabular}
+instCount&10000&10000&10000&10000\\ \hline\end{tabular}
 \caption{Benchmark Instruction Profiles showing percent rate of occurrence of each instruction type.}
@@
 \label{sec:circ}
-The circular ROB is the basic current implementation of the reorder buffer.
-It is implemented as a circular FIFO buffer, and the beginning and end of the used region are tracked by the head and tail pointers, respectively.
-When an instruction is issued, space is reserved for the instruction's result at the tail pointer, and the tail pointer is incremented.
-When the instruction completes, the result is inserted into the reorder buffer at the reserved space.
-The result is finally committed to the architecture register file when the head pointer passes through it; each cycle, results are committed from the head pointer and the head pointer is incremented repeatedly, until either an unfinished instruction result slot is reached or all write ports are used up.
+\subsubsection{Distributed (Cluster based) ROB}
+Our distributed (cluster based) ROB is modeled similarly to the circular ROB.  Instead of having a single ROB buffer, 
+numerous ROB buffers are instantiated.  In a real system, there would be one ROB buffer for each functional 
+unit cluster.  However, since we do not model functional units in our simulator, we simply divide the ROB 
+into multiple smaller ROBs.  Each of these ROBs implement the default circular ROB model.  Each ROB segment 
+features a head pointer and a tail pointer and contains an equal portion of the total ROB space.  No further 
+enhancements need to be made to the default circular ROB model as the inclusion of the central FIFO allows the 
+ROB clusters to operate entirely independently of one another.  Each ROB might as well be the only ROB in the system.
 
-The circular ROB is considered full when the tail pointer reaches the head pointer, as the head pointer denotes the start of the in-use region and no more space can be reserved from the tail pointer.
-Conversely, then buffer is empty when the head pointer catches up to the tail pointer, as every instruction in the buffer has now been committed to the ARF.
+The largest modeling change required by the distributed ROB is the inclusion of the central FIFO queue.  This 
+FIFO queue is responsible for maintaining a global state of order between instructions in the different ROB buffers. 
+The FIFO queue is modeled as a modified circular buffer and each entry contains: 1) PC of instruction, 
+2) ROB_ID where the instruction placeholder resides, and 3) a valid bit corresponding to the completion state of 
+the instruction.  The total number of entries in the FIFO queue is equal to the total entry size of the ROB.  
+Each active entry in the FIFO queue correlates with an active entry in one of the ROB clusters.
 
-\subsubsection{Distributed (Cluster based) ROB}
+The distributed ROB requires only minor changes to the basic circular buffer to realize its effectiveness.  Instructions are 
+read from the Instruction Queue as usual and inserted into their corresponding ROB entry while updating the corresponding tail 
+pointer.  Since there are now multiple ROBs into which instructions can be inserted, a quick modulus is performed on the 
+destination register of the instruction to divide the instructions equally between each ROB.  This presents the possibility 
+for a worst-case situation where the entire instruction stream is blocked due to improper balancing between ROB clusters.  
+This situation would occur very rarely in practice and would induce no form of deadlock or livelock in the system -- 
+it would merely affect performance.  Possible avoidance techniques can be researched as future work.
+
+At the same time as inserting the instruction from the IQ into the ROB cluster, an entry must be recorded in the central FIFO 
+queue to maintain global program order.  At this point, both the instruction PC and the recently determined ROB_ID are 
+noted in the central FIFO queue.  The tail pointer of the central FIFO queue is increased accordingly to match.
+
+Upon completion of an instruction, an instruction must be flagged as valid in both the ROB cluster in which it resides as well 
+as its corresponding central FIFO queue entry.  The central FIFO queue then directs the appropriate ROB to write the instruction 
+output to its write port when all prior instructions have been written so that program order is maintained.  This is the second 
+major behavioral difference between the circular ROB and the distributed ROB, as writes and write-order is driven and determined by 
+the central FIFO queue instead of the ROB cluster itself.
 
 \subsubsection{Dynamically Sized ROB}
-The dynamically sized ROB is an optimization on the circular ROB where the buffer is divided into partitions, and those partitions deemed unnecessary are powered off to conserve energy.
-Our implementation of the basic dynamic ROB is behaviorally the same as that detailed in \cite{kucuk2}.
-
-Downsizing the ROB is based on a periodic sampling of the buffer state.
-A number of cycles, {\it update\_period}, is defined, and every {\it update\_period} cycles the size of the ROB is updated.
-Every {\it sample\_period} cycles, the number of active entries is recorded, and the quantity {\it active\_size} is recorded as the average of these samples between each {\it update\_period}.
-The current maximum capacity of the reorder buffer is defined as the {\it current\_size}, and the size of a single ROB partition is defined as {\it partition\_size}.
-After each {\it update\_period} cycles, the ROB checks how full the buffer is, and if {\it current\_size} - {\it active\_size} > {\it partition\_size}, then a single partition is turned off.
-
-Increasing the size of the ROB is dependent on how often the buffer fills up.
-A value, defined as the {\it overflow\_count} is incremented each cycle that instruction issue is stalled because the ROB is full.
-An {\it overflow\_threshold} is defined, and if during an {\it update\_period}, {\it overflow\_count} > {\it overflow\_threshold}, then a partition is turned on.
-
-A conservative approach of deallocating strictly from one end of the buffer, and allocating strictly towards that end is followed.
-For example, if the ROB consisted of four partitions, downsizing the buffer would turn off the fourth, and then the third.
-At this point, if we were to increase the size of the buffer, the third partition would be turned on first, and the fourth partition next.
-
-When the ROB's size is updated, partitions are not immediately turned off, since the partition could coincide with the active region of the buffer.
-If deallocating, this would disable a part of the active region, making instructions in the region impossible to complete.
-If allocating, this would insert an empty segment right into the middle of the FIFO queue, logically breaking the buffer.
-It is only safe to turn a partition off/on when the active region is not wrapped around (that is, the head pointer is less than the tail pointer), and the tail pointer comes before the first index into the partition to turn on/off.
-\cite{kucuk3} details a worst case scenario of deallocating, where the active region is wrapped (head is greater than tail), and both pointers lie in the region to be turned off.
-In this case, the head pointer must first clear the entries already in the buffer.
-Since the tail pointer is already in the partition, it will continue to write into the partition, and the head pointer must come around a second time to clear out these new entries.
-Only then can the partition be safely deallocated, after the head pointer has wrapped around to zero twice.
-
-\subsubsection{Optimized Dynamic ROB}
-We implement an optimized version of the dynamic ROB detailed in \cite{kucuk2} and \cite{kucuk3}, opting to be more aggressive with turning off partitions in order to save power, and to improve the partition allocation/deallocation scheme.
-
-When downsizing the regular dynamic ROB, a single partition is turned off if {\it current\_size} - {\it active\_size} > {\it partition\_size}.
-In an effort to minimize the buffer size and thus the power consumed, we choose to downsize by
-
-\begin{center}
-$\lfloor \frac{current\_size - active\_size}{partition\_size} \rfloor$
-\end{center}
-
-Thus if much less than the capacity of the buffer is active, several partitions may be turned off at once.
-
-To more efficiently resize the buffer, we found that choosing the first valid partition immediately after the one housing the tail pointer is optimal.
-For example, if we are allocating a partition in a four partition buffer, and the tail pointer resides in partition 3, we first check if partition 4 can be turned on, then partition 1, then partition 2.
-
-In allocating the buffer, the reasoning is as follows.
-We aim to enable a disabled partition.
-It is impossible for the head pointer to lie in the disabled partition, thus there is no need to consider the head pointer.
-To maintain the logical consistency of the buffer, we cannot enable a partition that would divide the active region of the buffer.
-Thus, we must allocate a partition that comes after the tail pointer, and if we wish the buffer size to grow as soon as possible, the partition to be allocated must be the first disabled partition after the tail pointer.
-
-In deallocating the buffer, turning off the first partition after the tail pointer is also optimal.
-Being after the tail pointer, once we mark the partition for disabling we can prevent the tail pointer from writing into it.
-This prevents the worst case scenario where the head pointer must wrap around 0 twice to disable a partition.
-The next enabled partition immediately after the tail pointer is also the one the active region will leave first (if the region is in the partition; if not the partition is already empty and ready to deactivate), thus it will empty out first.
-If we are to maximize the power savings, we must disable the first partition that is available to be disabled, thus the first active partition after the tail is the optimal choice.
-
-One last optmization we implemented was just-in-time allocation of partitions.
-Normally, when the tail pointer increments, if a partition is not active the tail pointer simply skips it.
-In our optimization, before skipping the tail pointer will check if the disabled partition is marked to be enabled.
-If so, instead of skipping over it, the partition will be turned on and the tail pointer will write into it.
-This ensures that each partition to be turned on is enabled at the last possible moment, saving the energy that would otherwise be consumed by an empty partition waiting to be written into.
-
 \subsubsection{Reduced Output Port ROB (Retention Latch)}
-
 \subsection{Power Tabulation Implementation}
@@
 1145/514191.514202 http://doi.acm.org/10.1145/514191.514202
-\bibitem{kucuk2}
-Gurhan Kucuk, Dmitry Ponomarev, and Kanad Ghose.
-2002.
-Energy-Efficient Design of the Reorder Buffer.
-In Proceedings of the International Workshop on Power and Timing Modeling, Optimization, and Simulation, September 2002.
-\bibitem{kucuk3}
-Gorhan Kucuk, Dmitry Ponomarev, and Kanad Ghose.
-2001.
-Reducing Power Requirements of Instruction Scheduling Through Dynamic Allocation of Multiple Datapath Resources.
-In Proceedings of the International Symposium on Microarchitecture, December 2001.
 \bibitem{rupnow}

### 2011-06-08 paper/paper.tex
Changes

--- 2011-06-08 paper/paper.tex
@@
 \subsection{Distributed ROB}
+An accepted optimization for improving the power consumption of an ROB unit 
+is to break up the ROB buffer into clusters REF[XXX].  In this organization, the in-flight 
+instruction information is distributed across all of the units and then re-assembled 
+back into program order after the fact.  The primary advantage of this approach 
+is that it dramatically lessens the size of the individual ROB units.  This reduces overall
+power by reducing the bit line width as well as unit capacitance.
+
+...IMAGE FROM OUR SLIDES...
+
+The overhead associated with this scheme is that the addition of a separate FIFO unit
+is required to maintain a global state of instruction order between the separate ROB units.
+This central FIFO unit is then referenced upon instruction re-order to commit the instructions
+in program order.  The contents of the FIFO unit are fairly simple and incur reasonable hardware
+overhead.  Each entry needs only contain the PC of the instruction (or some other unique identifier)
+and the corresponding ROB unit that houses the placeholder of the instruction as it is in-flight. 
+The total ROB bit storage count remains the same from the non-distributed model to the distributed model
+as the overall size of the ROB remains the same.  However, the inclusion of the central FIFO queue 
+induces some overhead bits that must be considered when analyzing the overall effectiveness of the 
+model.
+
+The distribution or clustering of the ROB unit also inherently induces a minimal performance overhead 
+in particular cases.  By breaking the ROB into smaller pieces and distributing the instruction 
+data across the unit, the placement of instructions becomes a performance factor.  Typically, a 
+distributed ROB model features only a single write port per ROB cluster.  As a result only one operand 
+can be forwarded at a time from the ROB.  In the rare case that an instruction requires two operands be 
+forwarded from the same ROB cluster a delay of one cycle is imposed while the operands are forwarded 
+sequentially.  It is reported, however, that this happens very rarely in practice and induces less than 1\% 
+performance degradation REF[XXX].
+\subsection{Dynamic ROB}
+\subsection{Latch ROB}
 
@@
 \subsubsection{Distributed (Cluster based) ROB}
+Our distributed (cluster based) ROB is modeled similarly to the circular ROB.  Instead of having a single ROB buffer, 
+numerous ROB buffers are instantiated.  In a real system, there would be one ROB buffer for each functional 
+unit cluster.  However, since we do not model functional units in our simulator, we simply divide the ROB 
+into multiple smaller ROBs.  Each of these ROBs implement the default circular ROB model.  Each ROB segment 
+features a head pointer and a tail pointer and contains an equal portion of the total ROB space.  No further 
+enhancements need to be made to the default circular ROB model as the inclusion of the central FIFO allows the 
+ROB clusters to operate entirely independently of one another.  Each ROB might as well be the only ROB in the system.
+
+The largest modeling change required by the distributed ROB is the inclusion of the central FIFO queue.  This 
+FIFO queue is responsible for maintaining a global state of order between instructions in the different ROB buffers. 
+The FIFO queue is modeled as a modified circular buffer and each entry contains: 1) PC of instruction, 
+2) ROB{\char'137}ID where the instruction placeholder resides, and 3) a valid bit corresponding to the completion state of 
+the instruction.  The total number of entries in the FIFO queue is equal to the total entry size of the ROB.  
+Each active entry in the FIFO queue correlates with an active entry in one of the ROB clusters.
+
+The distributed ROB requires only minor changes to the basic circular buffer to realize its effectiveness.  Instructions are 
+read from the Instruction Queue as usual and inserted into their corresponding ROB entry while updating the corresponding tail 
+pointer.  Since there are now multiple ROBs into which instructions can be inserted, a quick modulus is performed on the 
+destination register of the instruction to divide the instructions equally between each ROB.  This presents the possibility 
+for a worst-case situation where the entire instruction stream is blocked due to improper balancing between ROB clusters.  
+This situation would occur very rarely in practice and would induce no form of deadlock or livelock in the system -- 
+it would merely affect performance.  Possible avoidance techniques can be researched as future work.
+
+At the same time as inserting the instruction from the IQ into the ROB cluster, an entry must be recorded in the central FIFO 
+queue to maintain global program order.  At this point, both the instruction PC and the recently determined ROB{\char'137}ID are 
+noted in the central FIFO queue.  The tail pointer of the central FIFO queue is increased accordingly to match.
+
+Upon completion of an instruction, an instruction must be flagged as valid in both the ROB cluster in which it resides as well 
+as its corresponding central FIFO queue entry.  The central FIFO queue then directs the appropriate ROB to write the instruction 
+output to its write port when all prior instructions have been written so that program order is maintained.  This is the second 
+major behavioral difference between the circular ROB and the distributed ROB, as writes and write-order is driven and determined by 
+the central FIFO queue instead of the ROB cluster itself.
 

### 2011-06-08 paper/bg.tex
multi file paper

--- 2011-06-08 paper/paper.tex
@@
-% THIS IS SIGPROC-SP.TEX - VERSION 3.1
-% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
-% APRIL 2009
-%
-% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
-% LaTeX2e document class file for Conference Proceedings submissions.
-% ----------------------------------------------------------------------------------------------------------------
-% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
-%       1) The Permission Statement
-%       2) The Conference (location) Info information
-%       3) The Copyright Line with ACM data
-%       4) Page numbering
-% ---------------------------------------------------------------------------------------------------------------
-% It is an example which *does* use the .bib file (from which the .bbl file
-% is produced).
-% REMEMBER HOWEVER: After having produced the .bbl file,
-% and prior to final submission,
-% you need to 'insert'  your .bbl file into your source .tex file so as to provide
-% ONE 'self-contained' source file.
-%
-% Questions regarding SIGS should be sent to
-% Adrienne Griscti ---> griscti@acm.org
-%
-% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
-% Gerald Murray ---> murray@hq.acm.org
-%
-% For tracking purposes - this is V3.1SP - APRIL 2009
-
-\documentclass{acm_proc_article-sp}
-
-\begin{document}
-%TODO
-\title{Comparing Power Optimizations in the Reorder Buffer}
-%
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
-
-\numberofauthors{3} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
-\author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
-\alignauthor
-Chen Guo\\
-       \affaddr{UCLA Computer Science Department}\\
-       \affaddr{4732 Boelter Hall}\\
-       \affaddr{Los Angeles, CA 90095}\\
-% 2nd. author
-\alignauthor
-Alexander Edwards\\
-       \affaddr{UCLA Computer Science Department}\\
-       \affaddr{4732 Boelter Hall}\\
-       \affaddr{Los Angeles, CA 90095}\\
-% 3rd. author
-\alignauthor
-Henry Hsieh\\
-       \affaddr{UCLA Computer Science Department}\\
-       \affaddr{4732 Boelter Hall}\\
-       \affaddr{Los Angeles, CA 90095}\\
-}
-% There's nothing stopping you putting the seventh, eighth, etc.
-% author on the opening page (as the 'third row') but we ask,
-% for aesthetic reasons that you place these 'additional authors'
-% in the \additional authors block, viz.
-% Just remember to make sure that the TOTAL number of authors
-% is the number that will appear on the first page PLUS the
-% number that will appear in the \additionalauthors section.
-
-\maketitle
-\begin{abstract}
-	In an effort to increase instruction level parallelism techniques such 
-as pipelining, out of order execution and operand forwarding.  However 
-these efforts come at a large power cost.  Out of order execution requires
-a ReOrder Buffer (ROB), to commit the registers in program order.  This 
-structure has grown to be monolithic with numerous entries and read/write ports.
-There is no current methods to efficiently and equally compare various
-ROB advancements to improve power efficiency. We propose a loosely clycle
-accurate model that keeps track of the power usages of the ROB.  Using the 
-simulation tool, we are able to compare and contrast the affects of improvements
-made on the ROB.  This also allows for a first line test to validate proof of 
-concepts for future power efficient ROB designs. 
-\end{abstract}
-
-\terms{Power, Energy, Architecture}
-
-\keywords{Power efficiency, ROB, Simulation} % NOT required for Proceedings
-
-\section{Introduction}
-BLAH BLAH BLAH INTRO WEEEE
 \section{Background}
@@
 \subsection{Circular ROB}
-The circular ROB is 
+The circular ROB is one of the most basic implementations of the ROB.  Simply using
+a circular buffer as the structure to hold each entry.  Since the ROB is basically
+just memory, there are two special slots for to keep track of the head and tail
+pointer to create the circular buffer.  As entries are written into the ROB, the
+entry is written at the tail pointer and then the tail pointer is incremented.
+Likewise when an entries is written to the ARF, the head pointer is incremented.
+The circular buffer method is widely used since it inherently keeps track of the 
+oldest instruction (which is at the head pointer), as well as which entry slots
+are used or empty.   However it may not be very power efficient if there is a large
+number of empty slots at any given point of the execution.    
+\subsection{Dynamic ROB}
+The dynamic ROB aims to reduce power by turning off portions of the buffer when they
+are not needed, and turning them back on when they become necessary. By partitioning
+the buffer into what is essentially smaller buffers, the complexity of each partition
+is reduced. Further, the ROB will then be custom sized to need, thus reducing energy
+wasted powering large stretches of empty entries. Due to the large percentage of energy
+consumption stemming from leakage power, turning off partitions contribute significant
+power savings over the base circular ROB.
+
 \subsection{Distributed ROB}
@@
 performance degradation REF[XXX].
-\subsection{Dynamic ROB}
 \subsection{Latch ROB}
-
-\section{Project Implementation}
-Our simulator provides a cycle accurate behavioral model of a standard circular re-order buffer.
-At the outset of the project, the standard re-order buffer we intended to implement was a non-circular RAM-based re-order buffer.
-However, after feedback from the course instructor and students it was decided that a circular buffer is more widely held as being the 'standard' buffer type and would serve as an ideal underlying model for power optimizations.
-Thus, we implement and compare the standard circular re-order buffer to circular re-order buffers with the following optimizations: 1) Distributed (Cluster based) ROB, 2) Dynamically Sized ROB, 3) Reduced Output Port ROB (Retention Latch).
-The following sections outline each of the above model implementations as well as input instruction generation and power results tabulation.
-\subsection{Instruction Generation}
-
-\begin{table*}
-\centering
-\begin{tabular}{|c|c|c|c|} \hline
-Instruction Type&Cycle Count&Instruction Type&Cycle Count\\ \hline
-\texttt{LOAD}&20&\texttt{STORE}&20\\ \hline
-\texttt{ADD}&1&\texttt{FADD}&3\\ \hline
-\texttt{SUB}&1&\texttt{FSUB}&3\\ \hline
-\texttt{MULT}&3&\texttt{FMULT}&5\\ \hline
-\texttt{DIV}&18&\texttt{FDIV}&6\\ \hline\end{tabular}
-\caption{Modeled instruction types and their corresponding completion cycle counts.}
-\label{tab:instypetable}
-\end{table*}
-
-We model the input to our simulator as a stream of integer and floatpoint instructions.
-The instructions are represented at the bit-level and model MIPS instructions as closely as possible.
-Each instruction consists of 1) Instruction type, 2) Source and destination registers 3) a PC.
-Table ~\ref{tab:instypetable} indicates the types of instructions modeled.
-Each instruction type also has a predetermined cycle execution time designating the number of cycles that must elapse between instruction issue and the completion of the instruction.
-As modeling functional units goes beyond the scope of this simulator, instructions are flagged as complete after the listed number of cycles has elapsed.
-The actual cycle counts are the same as those modeled in MIT's Graphite many-core processor simulator\cite{graphite}, with the exception of LOAD and STORE.
-The cycle counts corresponding to LOAD and STORE instructions are estimated as a predetermined 20 cycles each.
-This is simply an estimation as the actual cycle count is heavily dependent on cache architecture and benchmark specific cache behavior.
-Since both of these topics are beyond the scope of the simulator, we estimate the completion time of LOAD and STORE instructions as a set value substantially longer than the other more accurate instruction types.
-
-Registers are generated in bit representation as 4 bits each totalling 12 bits for the s1, s2 and d registers in each instruction.
-Currently, our representation of instructions make no distinction between zero, 1, and 2 operand instructions.
-All instructions regardless of type feature two operands.
-This feature is supported by the fact that a memory hierarchy and functional units are not simulated by our system and as such immediate values have no direct use.
-Further, every instruction features a d register regardless of type.
-Because our simulator features operand forwarding from the ROB, it is necessary for the generation of registers to keep track of past destination registers to use as dependent operands in future instructions.
-This history buffer is implemented as a rotating bit register from which prior destination registers can be indexed.
-
-Each instruction also features a program counter associated with the instruction.
-While this is clearly a simplification of a real system, it serves the purposes of our simulation as a uniquely identifiable tag by which to differentiate instructions.
-In both a real system and our simulator, a statically set 32 bit value is used to represent this PC/tag.
-As such, the specific values contained in the PC/tag will yield accurate transition results (nonexistent) and is thus are irrelevant.
-Simply starting the PC of the first instruction in the stream at 0 and incrementing by 4 each instruction guarantees uniqueness among PCs and serves the purposes of an instruction tag.
-
-The aforementioned instruction fields have multiple modes of generation.
-Our simulator synthetically generates an instruction input trace based on a prescribed instruction profile.
-The instruction profile features: 1) a percent rate of occurrence for each instruction type, 2) a rate corresponding to the interdependence of registers between instructions, and 3) an overall instruction count.
-
-The instruction generator is currently capable of generating four distinct instruction streams corresponding to different instruction workloads.
-The first instruction stream is marked as the default instruction stream.
-It features equal percentages of each type of instruction as well as a 10% register dependence rate.
-This register dependence rate was selected arbitrarily to act as a starting point for modeling data dependencies between instructions.
-
-The remaining three instruction streams are modeled after SPEC-FP benchmarks \cite{rupnow}.
-The benchmarks were chosen to represent three major modes of operation:  floating point heavy execution, integer heavy execution, and mixed type execution.
-For a floating point heavy execution, 173.applu was selected and modeled.
-173.applu models parabolic/elliptical partial differential equations and features almost exclusively floating point operations as well as a large number of memory operations making it useful as both a floating point benchmark as well as a long running memory intensive benchmark.
-191.fma3d is used to model integer heavy execution.
-It models finite-element crash simulation and features very few floating point operations and substantially fewer memory operations than 173.applu making it an ideal light-weight integer workload.
-Finally, 183.equake is used to represent the average workload.
-It models seismic wave propagation and features both integer and floating point operations as well as moderate memory usage.
-As such it acts as a more representative general workload than the original default workload.
-Each of the SPEC-FP benchmarks feature a 6{\char'045} register forwarding frequency as is reported by \cite{kucuk}.
-
-\begin{table*}
-\centering
-\begin{tabular}{|c|c|c|c|c|} \hline
-ins{\char'137}profile&default{\char'137}profile&173{\char'056}applu&191{\char'056}fma3d&183{\char'056}equake\\ \hline
-loadProc&0.10&0.32&0.12&0.37\\ \hline
-storeProc&0.10&0.12&0.24&0.06\\ \hline
-addProc&0.10&0.01&0.15&0.08\\ \hline
-subProc&0.10&0.01&0.15&0.08\\ \hline
-multProc&0.10&0.01&0.15&0.08\\ \hline
-divProc&0.10&0.01&0.15&0.08\\ \hline
-faddProc&0.10&0.13&0.01&0.07\\ \hline
-fsubProc&0.10&0.13&0.01&0.06\\ \hline
-fmultProc&0.10&0.13&0.01&0.06\\ \hline
-fdivProc&0.10&0.13&0.01&0.06\\ \hline
-forwardingRate&0.10&0.06&0.06&0.06\\ \hline
-instCount&10000&10000&10000&10000\\ \hline\end{tabular}
-\caption{Benchmark Instruction Profiles showing percent rate of occurrence of each instruction type.}
-\label{tab:insprofile}
-\end{table*}
-
-For the purposes of our power comparisons, each of the above instruction workloads are normalized to generate 10000 instructions.
-Such a large instruction window is ideal to allow buffers to fill and behavioral characteristics to develop.
-The specific values used for each workload are listed in Table ~\ref{tab:insprofile}.
-\subsection{Model Implementation}
-The following section details the model implementatations in our simulator.
-The basic re-order buffer model we use as a base-line for comparison is a standard circular ROB model.
-Section ~\ref{sec:circ} discusses its implementation.
-We further model three optimizations for re-order buffers in an attempt to improve overall power usage: 1) Distributed (Cluster based) ROB, 2) Dynamically Sized ROB, 3) Reduced Output Port ROB (Retention Latch).
-Each of these optimized models are built on top of the underlying circular ROB model to provide as close a behavioral and power comparison as possible.
-Each of the optimizations, as well as the standard circular reorder buffer, are described in detail below.
-\subsubsection{Circular ROB}
-\label{sec:circ}
-\subsubsection{Distributed (Cluster based) ROB}
-Our distributed (cluster based) ROB is modeled similarly to the circular ROB.  Instead of having a single ROB buffer, 
-numerous ROB buffers are instantiated.  In a real system, there would be one ROB buffer for each functional 
-unit cluster.  However, since we do not model functional units in our simulator, we simply divide the ROB 
-into multiple smaller ROBs.  Each of these ROBs implement the default circular ROB model.  Each ROB segment 
-features a head pointer and a tail pointer and contains an equal portion of the total ROB space.  No further 
-enhancements need to be made to the default circular ROB model as the inclusion of the central FIFO allows the 
-ROB clusters to operate entirely independently of one another.  Each ROB might as well be the only ROB in the system.
-
-The largest modeling change required by the distributed ROB is the inclusion of the central FIFO queue.  This 
-FIFO queue is responsible for maintaining a global state of order between instructions in the different ROB buffers. 
-The FIFO queue is modeled as a modified circular buffer and each entry contains: 1) PC of instruction, 
-2) ROB_ID where the instruction placeholder resides, and 3) a valid bit corresponding to the completion state of 
-the instruction.  The total number of entries in the FIFO queue is equal to the total entry size of the ROB.  
-Each active entry in the FIFO queue correlates with an active entry in one of the ROB clusters.
-
-The distributed ROB requires only minor changes to the basic circular buffer to realize its effectiveness.  Instructions are 
-read from the Instruction Queue as usual and inserted into their corresponding ROB entry while updating the corresponding tail 
-pointer.  Since there are now multiple ROBs into which instructions can be inserted, a quick modulus is performed on the 
-destination register of the instruction to divide the instructions equally between each ROB.  This presents the possibility 
-for a worst-case situation where the entire instruction stream is blocked due to improper balancing between ROB clusters.  
-This situation would occur very rarely in practice and would induce no form of deadlock or livelock in the system -- 
-it would merely affect performance.  Possible avoidance techniques can be researched as future work.
-
-At the same time as inserting the instruction from the IQ into the ROB cluster, an entry must be recorded in the central FIFO 
-queue to maintain global program order.  At this point, both the instruction PC and the recently determined ROB_ID are 
-noted in the central FIFO queue.  The tail pointer of the central FIFO queue is increased accordingly to match.
-
-Upon completion of an instruction, an instruction must be flagged as valid in both the ROB cluster in which it resides as well 
-as its corresponding central FIFO queue entry.  The central FIFO queue then directs the appropriate ROB to write the instruction 
-output to its write port when all prior instructions have been written so that program order is maintained.  This is the second 
-major behavioral difference between the circular ROB and the distributed ROB, as writes and write-order is driven and determined by 
-the central FIFO queue instead of the ROB cluster itself.
-
-\subsubsection{Dynamically Sized ROB}
-\subsubsection{Reduced Output Port ROB (Retention Latch)}
-\subsection{Power Tabulation Implementation}
-\section{Results}
-hi
-\subsection{subsection titl}
-sub section.. with table!
-\begin{table*}
-\centering
-\caption{Some Typical Commands}
-\begin{tabular}{|c|c|l|} \hline
-Command&A Number&Comments\\ \hline
-\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
-\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
-\texttt{{\char'134}table}& 300 & For tables\\ \hline
-\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
-\end{table*}
-% end the environment with {table*}, NOTE not {table}!
-
-\section{Conclusions}
-ending stuff
-%\end{document}  % This is where a 'short' article might terminate
-
-%ACKNOWLEDGMENTS are optional
-\section{Acknowledgments}
-if we really need this....
-
-%TODO add in bib 
-\begin{thebibliography}{------}
-\bibitem{graphite}
-	MIT Graphite Simulator
-	https://github.com/mit-carbon/Graphite/wiki
-\bibitem{kucuk}
-Gurhan Kucuk, Dmitry Ponomarev, and Kanad Ghose. 2002. Low-complexity reorder buffer architecture. In Proceedings
-of the 16th international conference on Supercomputing (ICS '02). ACM, New York, NY, USA, 57-66. DOI=10.
-1145/514191.514202 http://doi.acm.org/10.1145/514191.514202
-\bibitem{rupnow}
-Kyle Rupnow, Arun Rodrigues, Keith Underwood, and Katherine Compton.
-2006.
-Scientific applications vs.
-SPEC-FP: a comparison of program behavior.
-In Proceedings of the 20th annual international conference on Supercomputing (ICS '06).
-ACM, New York, NY, USA, 66-74.
-DOI=10.1145/1183401.1183413 http://doi.acm.org/10.1145/1183401.1183413
-\end{thebibliography}
-\end{document}
+In addition to large power requirement to maintain the entries of the ROB, another 
+large power consumer is the read ports of the ROB.  The read port of the ROB is used
+to forward the results of an entry to the dispatch unit (DU) to use as an operand.
+All though forwarding from the ROB occurs about 6\%\cite{kucuk} of all instructions 
+in an N way super scalar processor there are 2N ports (to accommodate FP
+ instructions) which add unnecessary complexity and power consumption for a small 
+percentage of instructions.  It has been proposed by eliminating the read ports 
+in the ROB all together, we can reduce energy consumption\cite{kucuk}.  However
+without read ports, a small window appears where the data cannot be accessed,
+starting from when the result is placed in to the ROB until the ROB commits that 
+entry to the ARF.  To help mitigate stalls that might happen due to this hole, 
+we introduce retention latches.   Retention latches are small memory elements and 
+we use them to cache the most recent results that have been written into the ROB.
+This way instructions are still able to access the most recently produced results
+and not stall.  However if the operand needed is not in the ROB, the stream will 
+stall until a commit is made to the ARF, and that result value is forwarded to 
+the dispatch unit immediately.

### 2011-06-08 paper/end.tex
multi file paper

--- 2011-06-08 paper/paper.tex
@@
-% THIS IS SIGPROC-SP.TEX - VERSION 3.1
-% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
-% APRIL 2009
-%
-% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
-% LaTeX2e document class file for Conference Proceedings submissions.
-% ----------------------------------------------------------------------------------------------------------------
-% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
-%       1) The Permission Statement
-%       2) The Conference (location) Info information
-%       3) The Copyright Line with ACM data
-%       4) Page numbering
-% ---------------------------------------------------------------------------------------------------------------
-% It is an example which *does* use the .bib file (from which the .bbl file
-% is produced).
-% REMEMBER HOWEVER: After having produced the .bbl file,
-% and prior to final submission,
-% you need to 'insert'  your .bbl file into your source .tex file so as to provide
-% ONE 'self-contained' source file.
-%
-% Questions regarding SIGS should be sent to
-% Adrienne Griscti ---> griscti@acm.org
-%
-% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
-% Gerald Murray ---> murray@hq.acm.org
-%
-% For tracking purposes - this is V3.1SP - APRIL 2009
-
-\documentclass{acm_proc_article-sp}
-
-\begin{document}
-%TODO
-\title{Comparing Power Optimizations in the Reorder Buffer}
-%
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
-
-\numberofauthors{3} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
-\author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
-\alignauthor
-Chen Guo\\
-       \affaddr{UCLA Computer Science Department}\\
-       \affaddr{4732 Boelter Hall}\\
-       \affaddr{Los Angeles, CA 90095}\\
-% 2nd. author
-\alignauthor
-Alexander Edwards\\
-       \affaddr{UCLA Computer Science Department}\\
-       \affaddr{4732 Boelter Hall}\\
-       \affaddr{Los Angeles, CA 90095}\\
-% 3rd. author
-\alignauthor
-Henry Hsieh\\
-       \affaddr{UCLA Computer Science Department}\\
-       \affaddr{4732 Boelter Hall}\\
-       \affaddr{Los Angeles, CA 90095}\\
-}
-% There's nothing stopping you putting the seventh, eighth, etc.
-% author on the opening page (as the 'third row') but we ask,
-% for aesthetic reasons that you place these 'additional authors'
-% in the \additional authors block, viz.
-% Just remember to make sure that the TOTAL number of authors
-% is the number that will appear on the first page PLUS the
-% number that will appear in the \additionalauthors section.
-
-\maketitle
-\begin{abstract}
-	In an effort to increase instruction level parallelism techniques such 
-as pipelining, out of order execution and operand forwarding.  However 
-these efforts come at a large power cost.  Out of order execution requires
-a ReOrder Buffer (ROB), to commit the registers in program order.  This 
-structure has grown to be monolithic with numerous entries and read/write ports.
-There is no current methods to efficiently and equally compare various
-ROB advancements to improve power efficiency. We propose a loosely clycle
-accurate model that keeps track of the power usages of the ROB.  Using the 
-simulation tool, we are able to compare and contrast the affects of improvements
-made on the ROB.  This also allows for a first line test to validate proof of 
-concepts for future power efficient ROB designs. 
-\end{abstract}
-
-\terms{Power, Energy, Architecture}
-
-\keywords{Power efficiency, ROB, Simulation} % NOT required for Proceedings
-
-\section{Introduction}
-BLAH BLAH BLAH INTRO WEEEE
-\section{Background}
-    An out of order execution system has two additional stages 
-to the lifetime of the instruction on top of the traditional pipeline 
-stages.  Each of which a small memory array to keep track of instuction flow.
-The first addtion, the instruction queue(IQ) which holds instructions that are 
-waiting to be executed.  Each entry will hold the instruction -- including 
-type, operand registers, and destination registers.  The second structure 
-holds the instructions that are in flight but not yet commited to the register
-file. Each entry will hold both PC of the instruction, as well as the destination
-register and the result of the excution unit.  There is also a flag that will
-mark the result of the entry as valid.
-
-Replacing the traditional instruction dispatch stage, An placeholder will be made for the 
-instruction in ReOrder Buffer (ROB), and the instructions will be placed
-in the instruction queue(IQ). In this stage instructions will first wait for their operands if not already
-available.  When all operands are available, the instruction is ready to 
-be executed so wait until its appropriate execution unit is free.  At this 
-point the instruction exits the IQ and begins execution.
-
-Upon execution completion, the output value is stored in the ROB with its 
-corresponding entry and that entry is marked as valid or completed with a 
-1 bit flag.  An entry is commited from the ROB to the architectrual register
-file (ARF) when that entry is the oldest (lowest PC) and its result has been 
-marked as valid. 
-\subsection{Circular ROB}
-The circular ROB is 
-\subsection{Distributed ROB}
-An accepted optimization for improving the power consumption of an ROB unit 
-is to break up the ROB buffer into clusters REF[XXX].  In this organization, the in-flight 
-instruction information is distributed across all of the units and then re-assembled 
-back into program order after the fact.  The primary advantage of this approach 
-is that it dramatically lessens the size of the individual ROB units.  This reduces overall
-power by reducing the bit line width as well as unit capacitance.
-
-...IMAGE FROM OUR SLIDES...
-
-The overhead associated with this scheme is that the addition of a separate FIFO unit
-is required to maintain a global state of instruction order between the separate ROB units.
-This central FIFO unit is then referenced upon instruction re-order to commit the instructions
-in program order.  The contents of the FIFO unit are fairly simple and incur reasonable hardware
-overhead.  Each entry needs only contain the PC of the instruction (or some other unique identifier)
-and the corresponding ROB unit that houses the placeholder of the instruction as it is in-flight. 
-The total ROB bit storage count remains the same from the non-distributed model to the distributed model
-as the overall size of the ROB remains the same.  However, the inclusion of the central FIFO queue 
-induces some overhead bits that must be considered when analyzing the overall effectiveness of the 
-model.
-
-The distribution or clustering of the ROB unit also inherently induces a minimal performance overhead 
-in particular cases.  By breaking the ROB into smaller pieces and distributing the instruction 
-data across the unit, the placement of instructions becomes a performance factor.  Typically, a 
-distributed ROB model features only a single write port per ROB cluster.  As a result only one operand 
-can be forwarded at a time from the ROB.  In the rare case that an instruction requires two operands be 
-forwarded from the same ROB cluster a delay of one cycle is imposed while the operands are forwarded 
-sequentially.  It is reported, however, that this happens very rarely in practice and induces less than 1\% 
-performance degradation REF[XXX].
-\subsection{Dynamic ROB}
-\subsection{Latch ROB}
-
-\section{Project Implementation}
-Our simulator provides a cycle accurate behavioral model of a standard circular re-order buffer.
-At the outset of the project, the standard re-order buffer we intended to implement was a non-circular RAM-based re-order buffer.
-However, after feedback from the course instructor and students it was decided that a circular buffer is more widely held as being the 'standard' buffer type and would serve as an ideal underlying model for power optimizations.
-Thus, we implement and compare the standard circular re-order buffer to circular re-order buffers with the following optimizations: 1) Distributed (Cluster based) ROB, 2) Dynamically Sized ROB, 3) Reduced Output Port ROB (Retention Latch).
-The following sections outline each of the above model implementations as well as input instruction generation and power results tabulation.
-\subsection{Instruction Generation}
-
-\begin{table*}
-\centering
-\begin{tabular}{|c|c|c|c|} \hline
-Instruction Type&Cycle Count&Instruction Type&Cycle Count\\ \hline
-\texttt{LOAD}&20&\texttt{STORE}&20\\ \hline
-\texttt{ADD}&1&\texttt{FADD}&3\\ \hline
-\texttt{SUB}&1&\texttt{FSUB}&3\\ \hline
-\texttt{MULT}&3&\texttt{FMULT}&5\\ \hline
-\texttt{DIV}&18&\texttt{FDIV}&6\\ \hline\end{tabular}
-\caption{Modeled instruction types and their corresponding completion cycle counts.}
-\label{tab:instypetable}
-\end{table*}
-
-We model the input to our simulator as a stream of integer and floatpoint instructions.
-The instructions are represented at the bit-level and model MIPS instructions as closely as possible.
-Each instruction consists of 1) Instruction type, 2) Source and destination registers 3) a PC.
-Table ~\ref{tab:instypetable} indicates the types of instructions modeled.
-Each instruction type also has a predetermined cycle execution time designating the number of cycles that must elapse between instruction issue and the completion of the instruction.
-As modeling functional units goes beyond the scope of this simulator, instructions are flagged as complete after the listed number of cycles has elapsed.
-The actual cycle counts are the same as those modeled in MIT's Graphite many-core processor simulator\cite{graphite}, with the exception of LOAD and STORE.
-The cycle counts corresponding to LOAD and STORE instructions are estimated as a predetermined 20 cycles each.
-This is simply an estimation as the actual cycle count is heavily dependent on cache architecture and benchmark specific cache behavior.
-Since both of these topics are beyond the scope of the simulator, we estimate the completion time of LOAD and STORE instructions as a set value substantially longer than the other more accurate instruction types.
-
-Registers are generated in bit representation as 4 bits each totalling 12 bits for the s1, s2 and d registers in each instruction.
-Currently, our representation of instructions make no distinction between zero, 1, and 2 operand instructions.
-All instructions regardless of type feature two operands.
-This feature is supported by the fact that a memory hierarchy and functional units are not simulated by our system and as such immediate values have no direct use.
-Further, every instruction features a d register regardless of type.
-Because our simulator features operand forwarding from the ROB, it is necessary for the generation of registers to keep track of past destination registers to use as dependent operands in future instructions.
-This history buffer is implemented as a rotating bit register from which prior destination registers can be indexed.
-
-Each instruction also features a program counter associated with the instruction.
-While this is clearly a simplification of a real system, it serves the purposes of our simulation as a uniquely identifiable tag by which to differentiate instructions.
-In both a real system and our simulator, a statically set 32 bit value is used to represent this PC/tag.
-As such, the specific values contained in the PC/tag will yield accurate transition results (nonexistent) and is thus are irrelevant.
-Simply starting the PC of the first instruction in the stream at 0 and incrementing by 4 each instruction guarantees uniqueness among PCs and serves the purposes of an instruction tag.
-
-The aforementioned instruction fields have multiple modes of generation.
-Our simulator synthetically generates an instruction input trace based on a prescribed instruction profile.
-The instruction profile features: 1) a percent rate of occurrence for each instruction type, 2) a rate corresponding to the interdependence of registers between instructions, and 3) an overall instruction count.
-
-The instruction generator is currently capable of generating four distinct instruction streams corresponding to different instruction workloads.
-The first instruction stream is marked as the default instruction stream.
-It features equal percentages of each type of instruction as well as a 10% register dependence rate.
-This register dependence rate was selected arbitrarily to act as a starting point for modeling data dependencies between instructions.
-
-The remaining three instruction streams are modeled after SPEC-FP benchmarks \cite{rupnow}.
-The benchmarks were chosen to represent three major modes of operation:  floating point heavy execution, integer heavy execution, and mixed type execution.
-For a floating point heavy execution, 173.applu was selected and modeled.
-173.applu models parabolic/elliptical partial differential equations and features almost exclusively floating point operations as well as a large number of memory operations making it useful as both a floating point benchmark as well as a long running memory intensive benchmark.
-191.fma3d is used to model integer heavy execution.
-It models finite-element crash simulation and features very few floating point operations and substantially fewer memory operations than 173.applu making it an ideal light-weight integer workload.
-Finally, 183.equake is used to represent the average workload.
-It models seismic wave propagation and features both integer and floating point operations as well as moderate memory usage.
-As such it acts as a more representative general workload than the original default workload.
-Each of the SPEC-FP benchmarks feature a 6{\char'045} register forwarding frequency as is reported by \cite{kucuk}.
-
-\begin{table*}
-\centering
-\begin{tabular}{|c|c|c|c|c|} \hline
-ins{\char'137}profile&default{\char'137}profile&173{\char'056}applu&191{\char'056}fma3d&183{\char'056}equake\\ \hline
-loadProc&0.10&0.32&0.12&0.37\\ \hline
-storeProc&0.10&0.12&0.24&0.06\\ \hline
-addProc&0.10&0.01&0.15&0.08\\ \hline
-subProc&0.10&0.01&0.15&0.08\\ \hline
-multProc&0.10&0.01&0.15&0.08\\ \hline
-divProc&0.10&0.01&0.15&0.08\\ \hline
-faddProc&0.10&0.13&0.01&0.07\\ \hline
-fsubProc&0.10&0.13&0.01&0.06\\ \hline
-fmultProc&0.10&0.13&0.01&0.06\\ \hline
-fdivProc&0.10&0.13&0.01&0.06\\ \hline
-forwardingRate&0.10&0.06&0.06&0.06\\ \hline
-instCount&10000&10000&10000&10000\\ \hline\end{tabular}
-\caption{Benchmark Instruction Profiles showing percent rate of occurrence of each instruction type.}
-\label{tab:insprofile}
-\end{table*}
-
-For the purposes of our power comparisons, each of the above instruction workloads are normalized to generate 10000 instructions.
-Such a large instruction window is ideal to allow buffers to fill and behavioral characteristics to develop.
-The specific values used for each workload are listed in Table ~\ref{tab:insprofile}.
-\subsection{Model Implementation}
-The following section details the model implementatations in our simulator.
-The basic re-order buffer model we use as a base-line for comparison is a standard circular ROB model.
-Section ~\ref{sec:circ} discusses its implementation.
-We further model three optimizations for re-order buffers in an attempt to improve overall power usage: 1) Distributed (Cluster based) ROB, 2) Dynamically Sized ROB, 3) Reduced Output Port ROB (Retention Latch).
-Each of these optimized models are built on top of the underlying circular ROB model to provide as close a behavioral and power comparison as possible.
-Each of the optimizations, as well as the standard circular reorder buffer, are described in detail below.
-\subsubsection{Circular ROB}
-\label{sec:circ}
-\subsubsection{Distributed (Cluster based) ROB}
-Our distributed (cluster based) ROB is modeled similarly to the circular ROB.  Instead of having a single ROB buffer, 
-numerous ROB buffers are instantiated.  In a real system, there would be one ROB buffer for each functional 
-unit cluster.  However, since we do not model functional units in our simulator, we simply divide the ROB 
-into multiple smaller ROBs.  Each of these ROBs implement the default circular ROB model.  Each ROB segment 
-features a head pointer and a tail pointer and contains an equal portion of the total ROB space.  No further 
-enhancements need to be made to the default circular ROB model as the inclusion of the central FIFO allows the 
-ROB clusters to operate entirely independently of one another.  Each ROB might as well be the only ROB in the system.
-
-The largest modeling change required by the distributed ROB is the inclusion of the central FIFO queue.  This 
-FIFO queue is responsible for maintaining a global state of order between instructions in the different ROB buffers. 
-The FIFO queue is modeled as a modified circular buffer and each entry contains: 1) PC of instruction, 
-2) ROB_ID where the instruction placeholder resides, and 3) a valid bit corresponding to the completion state of 
-the instruction.  The total number of entries in the FIFO queue is equal to the total entry size of the ROB.  
-Each active entry in the FIFO queue correlates with an active entry in one of the ROB clusters.
-
-The distributed ROB requires only minor changes to the basic circular buffer to realize its effectiveness.  Instructions are 
-read from the Instruction Queue as usual and inserted into their corresponding ROB entry while updating the corresponding tail 
-pointer.  Since there are now multiple ROBs into which instructions can be inserted, a quick modulus is performed on the 
-destination register of the instruction to divide the instructions equally between each ROB.  This presents the possibility 
-for a worst-case situation where the entire instruction stream is blocked due to improper balancing between ROB clusters.  
-This situation would occur very rarely in practice and would induce no form of deadlock or livelock in the system -- 
-it would merely affect performance.  Possible avoidance techniques can be researched as future work.
-
-At the same time as inserting the instruction from the IQ into the ROB cluster, an entry must be recorded in the central FIFO 
-queue to maintain global program order.  At this point, both the instruction PC and the recently determined ROB_ID are 
-noted in the central FIFO queue.  The tail pointer of the central FIFO queue is increased accordingly to match.
-
-Upon completion of an instruction, an instruction must be flagged as valid in both the ROB cluster in which it resides as well 
-as its corresponding central FIFO queue entry.  The central FIFO queue then directs the appropriate ROB to write the instruction 
-output to its write port when all prior instructions have been written so that program order is maintained.  This is the second 
-major behavioral difference between the circular ROB and the distributed ROB, as writes and write-order is driven and determined by 
-the central FIFO queue instead of the ROB cluster itself.
-
-\subsubsection{Dynamically Sized ROB}
-\subsubsection{Reduced Output Port ROB (Retention Latch)}
-\subsection{Power Tabulation Implementation}
 \section{Results}
@@
 ending stuff
-%\end{document}  % This is where a 'short' article might terminate
-
-%ACKNOWLEDGMENTS are optional
-\section{Acknowledgments}
-if we really need this....
-
-%TODO add in bib 
-\begin{thebibliography}{------}
-\bibitem{graphite}
-	MIT Graphite Simulator
-	https://github.com/mit-carbon/Graphite/wiki
-\bibitem{kucuk}
-Gurhan Kucuk, Dmitry Ponomarev, and Kanad Ghose. 2002. Low-complexity reorder buffer architecture. In Proceedings
-of the 16th international conference on Supercomputing (ICS '02). ACM, New York, NY, USA, 57-66. DOI=10.
-1145/514191.514202 http://doi.acm.org/10.1145/514191.514202
-\bibitem{rupnow}
-Kyle Rupnow, Arun Rodrigues, Keith Underwood, and Katherine Compton.
-2006.
-Scientific applications vs.
-SPEC-FP: a comparison of program behavior.
-In Proceedings of the 20th annual international conference on Supercomputing (ICS '06).
-ACM, New York, NY, USA, 66-74.
-DOI=10.1145/1183401.1183413 http://doi.acm.org/10.1145/1183401.1183413
-\end{thebibliography}
-\end{document}

### 2011-06-08 paper/paper.tex
multi file paper

--- 2011-06-08 paper/paper.tex
@@
 \maketitle
-\begin{abstract}
-	In an effort to increase instruction level parallelism techniques such 
-as pipelining, out of order execution and operand forwarding.  However 
-these efforts come at a large power cost.  Out of order execution requires
-a ReOrder Buffer (ROB), to commit the registers in program order.  This 
-structure has grown to be monolithic with numerous entries and read/write ports.
-There is no current methods to efficiently and equally compare various
-ROB advancements to improve power efficiency. We propose a loosely clycle
-accurate model that keeps track of the power usages of the ROB.  Using the 
-simulation tool, we are able to compare and contrast the affects of improvements
-made on the ROB.  This also allows for a first line test to validate proof of 
-concepts for future power efficient ROB designs. 
-\end{abstract}
+\input{start}
 
-\terms{Power, Energy, Architecture}
+\input{bg}
 
-\keywords{Power efficiency, ROB, Simulation} % NOT required for Proceedings
+\input{proj}
 
-\section{Introduction}
-BLAH BLAH BLAH INTRO WEEEE
-\section{Background}
-    An out of order execution system has two additional stages 
-to the lifetime of the instruction on top of the traditional pipeline 
-stages.  Each of which a small memory array to keep track of instuction flow.
-The first addtion, the instruction queue(IQ) which holds instructions that are 
-waiting to be executed.  Each entry will hold the instruction -- including 
-type, operand registers, and destination registers.  The second structure 
-holds the instructions that are in flight but not yet commited to the register
-file. Each entry will hold both PC of the instruction, as well as the destination
-register and the result of the excution unit.  There is also a flag that will
-mark the result of the entry as valid.
+\input{end}
 
-Replacing the traditional instruction dispatch stage, An placeholder will be made for the 
-instruction in ReOrder Buffer (ROB), and the instructions will be placed
-in the instruction queue(IQ). In this stage instructions will first wait for their operands if not already
-available.  When all operands are available, the instruction is ready to 
-be executed so wait until its appropriate execution unit is free.  At this 
-point the instruction exits the IQ and begins execution.
-
-Upon execution completion, the output value is stored in the ROB with its 
-corresponding entry and that entry is marked as valid or completed with a 
-1 bit flag.  An entry is commited from the ROB to the architectrual register
-file (ARF) when that entry is the oldest (lowest PC) and its result has been 
-marked as valid. 
-\subsection{Circular ROB}
-The circular ROB is 
-\subsection{Distributed ROB}
-An accepted optimization for improving the power consumption of an ROB unit 
-is to break up the ROB buffer into clusters REF[XXX].  In this organization, the in-flight 
-instruction information is distributed across all of the units and then re-assembled 
-back into program order after the fact.  The primary advantage of this approach 
-is that it dramatically lessens the size of the individual ROB units.  This reduces overall
-power by reducing the bit line width as well as unit capacitance.
-
-...IMAGE FROM OUR SLIDES...
-
-The overhead associated with this scheme is that the addition of a separate FIFO unit
-is required to maintain a global state of instruction order between the separate ROB units.
-This central FIFO unit is then referenced upon instruction re-order to commit the instructions
-in program order.  The contents of the FIFO unit are fairly simple and incur reasonable hardware
-overhead.  Each entry needs only contain the PC of the instruction (or some other unique identifier)
-and the corresponding ROB unit that houses the placeholder of the instruction as it is in-flight. 
-The total ROB bit storage count remains the same from the non-distributed model to the distributed model
-as the overall size of the ROB remains the same.  However, the inclusion of the central FIFO queue 
-induces some overhead bits that must be considered when analyzing the overall effectiveness of the 
-model.
-
-The distribution or clustering of the ROB unit also inherently induces a minimal performance overhead 
-in particular cases.  By breaking the ROB into smaller pieces and distributing the instruction 
-data across the unit, the placement of instructions becomes a performance factor.  Typically, a 
-distributed ROB model features only a single write port per ROB cluster.  As a result only one operand 
-can be forwarded at a time from the ROB.  In the rare case that an instruction requires two operands be 
-forwarded from the same ROB cluster a delay of one cycle is imposed while the operands are forwarded 
-sequentially.  It is reported, however, that this happens very rarely in practice and induces less than 1\% 
-performance degradation REF[XXX].
-\subsection{Dynamic ROB}
-\subsection{Latch ROB}
-
-\section{Project Implementation}
-Our simulator provides a cycle accurate behavioral model of a standard circular re-order buffer.
-At the outset of the project, the standard re-order buffer we intended to implement was a non-circular RAM-based re-order buffer.
-However, after feedback from the course instructor and students it was decided that a circular buffer is more widely held as being the 'standard' buffer type and would serve as an ideal underlying model for power optimizations.
-Thus, we implement and compare the standard circular re-order buffer to circular re-order buffers with the following optimizations: 1) Distributed (Cluster based) ROB, 2) Dynamically Sized ROB, 3) Reduced Output Port ROB (Retention Latch).
-The following sections outline each of the above model implementations as well as input instruction generation and power results tabulation.
-\subsection{Instruction Generation}
-
-\begin{table*}
-\centering
-\begin{tabular}{|c|c|c|c|} \hline
-Instruction Type&Cycle Count&Instruction Type&Cycle Count\\ \hline
-\texttt{LOAD}&20&\texttt{STORE}&20\\ \hline
-\texttt{ADD}&1&\texttt{FADD}&3\\ \hline
-\texttt{SUB}&1&\texttt{FSUB}&3\\ \hline
-\texttt{MULT}&3&\texttt{FMULT}&5\\ \hline
-\texttt{DIV}&18&\texttt{FDIV}&6\\ \hline\end{tabular}
-\caption{Modeled instruction types and their corresponding completion cycle counts.}
-\label{tab:instypetable}
-\end{table*}
-
-We model the input to our simulator as a stream of integer and floatpoint instructions.
-The instructions are represented at the bit-level and model MIPS instructions as closely as possible.
-Each instruction consists of 1) Instruction type, 2) Source and destination registers 3) a PC.
-Table ~\ref{tab:instypetable} indicates the types of instructions modeled.
-Each instruction type also has a predetermined cycle execution time designating the number of cycles that must elapse between instruction issue and the completion of the instruction.
-As modeling functional units goes beyond the scope of this simulator, instructions are flagged as complete after the listed number of cycles has elapsed.
-The actual cycle counts are the same as those modeled in MIT's Graphite many-core processor simulator\cite{graphite}, with the exception of LOAD and STORE.
-The cycle counts corresponding to LOAD and STORE instructions are estimated as a predetermined 20 cycles each.
-This is simply an estimation as the actual cycle count is heavily dependent on cache architecture and benchmark specific cache behavior.
-Since both of these topics are beyond the scope of the simulator, we estimate the completion time of LOAD and STORE instructions as a set value substantially longer than the other more accurate instruction types.
-
-Registers are generated in bit representation as 4 bits each totalling 12 bits for the s1, s2 and d registers in each instruction.
-Currently, our representation of instructions make no distinction between zero, 1, and 2 operand instructions.
-All instructions regardless of type feature two operands.
-This feature is supported by the fact that a memory hierarchy and functional units are not simulated by our system and as such immediate values have no direct use.
-Further, every instruction features a d register regardless of type.
-Because our simulator features operand forwarding from the ROB, it is necessary for the generation of registers to keep track of past destination registers to use as dependent operands in future instructions.
-This history buffer is implemented as a rotating bit register from which prior destination registers can be indexed.
-
-Each instruction also features a program counter associated with the instruction.
-While this is clearly a simplification of a real system, it serves the purposes of our simulation as a uniquely identifiable tag by which to differentiate instructions.
-In both a real system and our simulator, a statically set 32 bit value is used to represent this PC/tag.
-As such, the specific values contained in the PC/tag will yield accurate transition results (nonexistent) and is thus are irrelevant.
-Simply starting the PC of the first instruction in the stream at 0 and incrementing by 4 each instruction guarantees uniqueness among PCs and serves the purposes of an instruction tag.
-
-The aforementioned instruction fields have multiple modes of generation.
-Our simulator synthetically generates an instruction input trace based on a prescribed instruction profile.
-The instruction profile features: 1) a percent rate of occurrence for each instruction type, 2) a rate corresponding to the interdependence of registers between instructions, and 3) an overall instruction count.
-
-The instruction generator is currently capable of generating four distinct instruction streams corresponding to different instruction workloads.
-The first instruction stream is marked as the default instruction stream.
-It features equal percentages of each type of instruction as well as a 10% register dependence rate.
-This register dependence rate was selected arbitrarily to act as a starting point for modeling data dependencies between instructions.
-
-The remaining three instruction streams are modeled after SPEC-FP benchmarks \cite{rupnow}.
-The benchmarks were chosen to represent three major modes of operation:  floating point heavy execution, integer heavy execution, and mixed type execution.
-For a floating point heavy execution, 173.applu was selected and modeled.
-173.applu models parabolic/elliptical partial differential equations and features almost exclusively floating point operations as well as a large number of memory operations making it useful as both a floating point benchmark as well as a long running memory intensive benchmark.
-191.fma3d is used to model integer heavy execution.
-It models finite-element crash simulation and features very few floating point operations and substantially fewer memory operations than 173.applu making it an ideal light-weight integer workload.
-Finally, 183.equake is used to represent the average workload.
-It models seismic wave propagation and features both integer and floating point operations as well as moderate memory usage.
-As such it acts as a more representative general workload than the original default workload.
-Each of the SPEC-FP benchmarks feature a 6{\char'045} register forwarding frequency as is reported by \cite{kucuk}.
-
-\begin{table*}
-\centering
-\begin{tabular}{|c|c|c|c|c|} \hline
-ins{\char'137}profile&default{\char'137}profile&173{\char'056}applu&191{\char'056}fma3d&183{\char'056}equake\\ \hline
-loadProc&0.10&0.32&0.12&0.37\\ \hline
-storeProc&0.10&0.12&0.24&0.06\\ \hline
-addProc&0.10&0.01&0.15&0.08\\ \hline
-subProc&0.10&0.01&0.15&0.08\\ \hline
-multProc&0.10&0.01&0.15&0.08\\ \hline
-divProc&0.10&0.01&0.15&0.08\\ \hline
-faddProc&0.10&0.13&0.01&0.07\\ \hline
-fsubProc&0.10&0.13&0.01&0.06\\ \hline
-fmultProc&0.10&0.13&0.01&0.06\\ \hline
-fdivProc&0.10&0.13&0.01&0.06\\ \hline
-forwardingRate&0.10&0.06&0.06&0.06\\ \hline
-instCount&10000&10000&10000&10000\\ \hline\end{tabular}
-\caption{Benchmark Instruction Profiles showing percent rate of occurrence of each instruction type.}
-\label{tab:insprofile}
-\end{table*}
-
-For the purposes of our power comparisons, each of the above instruction workloads are normalized to generate 10000 instructions.
-Such a large instruction window is ideal to allow buffers to fill and behavioral characteristics to develop.
-The specific values used for each workload are listed in Table ~\ref{tab:insprofile}.
-\subsection{Model Implementation}
-The following section details the model implementatations in our simulator.
-The basic re-order buffer model we use as a base-line for comparison is a standard circular ROB model.
-Section ~\ref{sec:circ} discusses its implementation.
-We further model three optimizations for re-order buffers in an attempt to improve overall power usage: 1) Distributed (Cluster based) ROB, 2) Dynamically Sized ROB, 3) Reduced Output Port ROB (Retention Latch).
-Each of these optimized models are built on top of the underlying circular ROB model to provide as close a behavioral and power comparison as possible.
-Each of the optimizations, as well as the standard circular reorder buffer, are described in detail below.
-\subsubsection{Circular ROB}
-\label{sec:circ}
-\subsubsection{Distributed (Cluster based) ROB}
-Our distributed (cluster based) ROB is modeled similarly to the circular ROB.  Instead of having a single ROB buffer, 
-numerous ROB buffers are instantiated.  In a real system, there would be one ROB buffer for each functional 
-unit cluster.  However, since we do not model functional units in our simulator, we simply divide the ROB 
-into multiple smaller ROBs.  Each of these ROBs implement the default circular ROB model.  Each ROB segment 
-features a head pointer and a tail pointer and contains an equal portion of the total ROB space.  No further 
-enhancements need to be made to the default circular ROB model as the inclusion of the central FIFO allows the 
-ROB clusters to operate entirely independently of one another.  Each ROB might as well be the only ROB in the system.
-
-The largest modeling change required by the distributed ROB is the inclusion of the central FIFO queue.  This 
-FIFO queue is responsible for maintaining a global state of order between instructions in the different ROB buffers. 
-The FIFO queue is modeled as a modified circular buffer and each entry contains: 1) PC of instruction, 
-2) ROB_ID where the instruction placeholder resides, and 3) a valid bit corresponding to the completion state of 
-the instruction.  The total number of entries in the FIFO queue is equal to the total entry size of the ROB.  
-Each active entry in the FIFO queue correlates with an active entry in one of the ROB clusters.
-
-The distributed ROB requires only minor changes to the basic circular buffer to realize its effectiveness.  Instructions are 
-read from the Instruction Queue as usual and inserted into their corresponding ROB entry while updating the corresponding tail 
-pointer.  Since there are now multiple ROBs into which instructions can be inserted, a quick modulus is performed on the 
-destination register of the instruction to divide the instructions equally between each ROB.  This presents the possibility 
-for a worst-case situation where the entire instruction stream is blocked due to improper balancing between ROB clusters.  
-This situation would occur very rarely in practice and would induce no form of deadlock or livelock in the system -- 
-it would merely affect performance.  Possible avoidance techniques can be researched as future work.
-
-At the same time as inserting the instruction from the IQ into the ROB cluster, an entry must be recorded in the central FIFO 
-queue to maintain global program order.  At this point, both the instruction PC and the recently determined ROB_ID are 
-noted in the central FIFO queue.  The tail pointer of the central FIFO queue is increased accordingly to match.
-
-Upon completion of an instruction, an instruction must be flagged as valid in both the ROB cluster in which it resides as well 
-as its corresponding central FIFO queue entry.  The central FIFO queue then directs the appropriate ROB to write the instruction 
-output to its write port when all prior instructions have been written so that program order is maintained.  This is the second 
-major behavioral difference between the circular ROB and the distributed ROB, as writes and write-order is driven and determined by 
-the central FIFO queue instead of the ROB cluster itself.
-
-\subsubsection{Dynamically Sized ROB}
-\subsubsection{Reduced Output Port ROB (Retention Latch)}
-\subsection{Power Tabulation Implementation}
-\section{Results}
-hi
-\subsection{subsection titl}
-sub section.. with table!
-\begin{table*}
-\centering
-\caption{Some Typical Commands}
-\begin{tabular}{|c|c|l|} \hline
-Command&A Number&Comments\\ \hline
-\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
-\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
-\texttt{{\char'134}table}& 300 & For tables\\ \hline
-\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
-\end{table*}
-% end the environment with {table*}, NOTE not {table}!
-
-\section{Conclusions}
-ending stuff
 %\end{document}  % This is where a 'short' article might terminate
@@
 %ACKNOWLEDGMENTS are optional
-\section{Acknowledgments}
-if we really need this....
 

### 2011-06-08 paper/proj.tex
multi file paper

\section{Project Implementation}
Our simulator provides a cycle accurate behavioral model of a standard circular re-order buffer.  At the outset of the project, the standard re-order buffer we intended to implement was a non-circular RAM-based re-order buffer.  However, after feedback from the course instructor and students it was decided that a circular buffer is more widely held as being the 'standard' buffer type and would serve as an ideal underlying model for power optimizations.  Thus, we implement and compare the standard circular re-order buffer to circular re-order buffers with the following optimizations: 1) Distributed (Cluster based) ROB, 2) Dynamically Sized ROB, 3) Reduced Output Port ROB (Retention Latch). The following sections outline each of the above model implementations as well as input instruction generation and power results tabulation.
\subsection{Instruction Generation}

\begin{table*}
\centering
\begin{tabular}{|c|c|c|c|} \hline
Instruction Type&Cycle Count&Instruction Type&Cycle Count\\ \hline
\texttt{LOAD}&20&\texttt{STORE}&20\\ \hline
\texttt{ADD}&1&\texttt{FADD}&3\\ \hline
\texttt{SUB}&1&\texttt{FSUB}&3\\ \hline
\texttt{MULT}&3&\texttt{FMULT}&5\\ \hline
\texttt{DIV}&18&\texttt{FDIV}&6\\ \hline\end{tabular}
\caption{Modeled instruction types and their corresponding completion cycle counts.}
\label{tab:instypetable}
\end{table*}

We model the input to our simulator as a stream of integer and floatpoint instructions.  The instructions are represented at the bit-level and model MIPS instructions as closely as possible.  Each instruction consists of 1) Instruction type, 2) Source and destination registers 3) a PC.  Table ~\ref{tab:instypetable} indicates the types of instructions modeled.  Each instruction type also has a predetermined cycle execution time designating the number of cycles that must elapse between instruction issue and the completion of the instruction.  As modeling functional units goes beyond the scope of this simulator, instructions are flagged as complete after the listed number of cycles has elapsed.  The actual cycle counts are the same as those modeled in MIT's Graphite many-core processor simulator\cite{graphite}, with the exception of LOAD and STORE.  The cycle counts corresponding to LOAD and STORE instructions are estimated as a predetermined 20 cycles each.  This is simply an estimation as the actual cycle count is heavily dependent on cache architecture and benchmark specific cache behavior.  Since both of these topics are beyond the scope of the simulator, we estimate the completion time of LOAD and STORE instructions as a set value substantially longer than the other more accurate instruction types.

Registers are generated in bit representation as 4 bits each totaling 12 bits for the s1, s2 and d registers in each instruction. Currently, our representation of instructions make no distinction between zero, 1, and 2 operand instructions.  All instructions regardless of type feature two operands.  This feature is supported by the fact that a memory hierarchy and functional units are not simulated by our system and as such immediate values have no direct use.  Further, every instruction features a d register regardless of type.  Because our simulator features operand forwarding from the ROB, it is necessary for the generation of registers to keep track of past destination registers to use as dependent operands in future instructions.  This history buffer is implemented as a rotating bit register from which prior destination registers can be indexed.

Each instruction also features a program counter associated with the instruction.  While this is clearly a simplification of a real system, it serves the purposes of our simulation as a uniquely identifiable tag by which to differentiate instructions.  In both a real system and our simulator, a statically set 32 bit value is used to represent this PC/tag.  As such, the specific values contained in the PC/tag will yield accurate transition results (nonexistent) and is thus are irrelevant.  Simply starting the PC of the first instruction in the stream at 0 and incrementing by 4 each instruction guarantees uniqueness among PCs and serves the purposes of an instruction tag.

The aforementioned instruction fields have multiple modes of generation.  Our simulator synthetically generates an instruction input trace based on a prescribed instruction profile.  The instruction profile features: 1) a percent rate of occurrence for each instruction type, 2) a rate corresponding to the interdependence of registers between instructions, and 3) an overall instruction count.

The instruction generator is currently capable of generating four distinct instruction streams corresponding to different instruction workloads.  The first instruction stream is marked as the default instruction stream.  It features equal percentages of each type of instruction as well as a 10\% register dependence rate.  This register dependence rate was selected arbitrarily to act as a starting point for modeling data dependencies between instructions.

The remaining three instruction streams are modeled after SPEC-FP benchmarks \cite{rupnow}.  The benchmarks were chosen to represent three major modes of operation:  floating point heavy execution, integer heavy execution, and mixed type execution.  For a floating point heavy execution, 173.applu was selected and modeled.  173.applu models parabolic/elliptical partial differential equations and features almost exclusively floating point operations as well as a large number of memory operations making it useful as both a floating point benchmark as well as a long running memory intensive benchmark.  191.fma3d is used to model integer heavy execution.  It models finite-element crash simulation and features very few floating point operations and substantially fewer memory operations than 173.applu making it an ideal light-weight integer workload.  Finally, 183.equake is used to represent the average workload.  It models seismic wave propagation and features both integer and floating point operations as well as moderate memory usage.  As such it acts as a more representative general workload than the original default workload.  Each of the SPEC-FP benchmarks feature a 6{\char'045} register forwarding frequency as is reported by \cite{kucuk}.

\begin{table*}
\centering
\begin{tabular}{|c|c|c|c|c|} \hline
ins{\char'137}profile&default{\char'137}profile&173{\char'056}applu&191{\char'056}fma3d&183{\char'056}equake\\ \hline
loadProc&0.10&0.32&0.12&0.37\\ \hline
storeProc&0.10&0.12&0.24&0.06\\ \hline
addProc&0.10&0.01&0.15&0.08\\ \hline
subProc&0.10&0.01&0.15&0.08\\ \hline
multProc&0.10&0.01&0.15&0.08\\ \hline
divProc&0.10&0.01&0.15&0.08\\ \hline
faddProc&0.10&0.13&0.01&0.07\\ \hline
fsubProc&0.10&0.13&0.01&0.06\\ \hline
fmultProc&0.10&0.13&0.01&0.06\\ \hline
fdivProc&0.10&0.13&0.01&0.06\\ \hline
forwardingRate&0.10&0.06&0.06&0.06\\ \hline
instCount&100000&100000&100000&100000\\ \hline\end{tabular}
\caption{Benchmark Instruction Profiles showing percent rate of occurrence of each instruction type.}
\label{tab:insprofile}
\end{table*}

For the purposes of our power comparisons, each of the above instruction workloads are normalized to generate 10000 instructions.  Such a large instruction window is ideal to allow buffers to fill and behavioral characteristics to develop.  The specific values used for each workload are listed in Table ~\ref{tab:insprofile}.
\subsection{Model Implementation}
The following section details the model implementatations in our simulator.  The basic re-order buffer model we use as a base-line for comparison is a standard circular ROB model.  Section ~\ref{sec:circ} discusses its implementation.  We further model three optimizations for re-order buffers in an attempt to improve overall power usage: 1) Distributed (Cluster based) ROB, 2) Dynamically Sized ROB, 3) Reduced Output Port ROB (Retention Latch).  Each of these optimized models are built on top of the underlying circular ROB model to provide as close a behavioral and power comparison as possible.  Each of the optimizations, as well as the standard circular reorder buffer, are described in detail below.
\subsubsection{Circular ROB}
\label{sec:circ}
The circular ROB is the basic current implementation of the reorder buffer. It is implemented as a circular FIFO buffer, and the beginning and end of the used region are tracked by the head and tail pointers, respectively. When an instruction is issued, space is reserved for the instruction's result at the tail pointer, and the tail pointer is incremented. When the instruction completes, the result is inserted into the reorder buffer at the reserved space. The result is finally committed to the architecture register file when the head pointer passes through it; each cycle, results are committed from the head pointer and the head pointer is incremented repeatedly, until either an unfinished instruction result slot is reached or all write ports are used up.

The circular ROB is considered full when the tail pointer reaches the head pointer, as the head pointer denotes the start of the in-use region and no more space can be reserved from the tail pointer. Conversely, then buffer is empty when the head pointer catches up to the tail pointer, as every instruction in the buffer has now been committed to the ARF.

\subsubsection{Distributed (Cluster based) ROB}
Our distributed (cluster based) ROB is modeled similarly to the circular ROB.  Instead of having a single ROB buffer, 
numerous ROB buffers are instantiated.  In a real system, there would be one ROB buffer for each functional 
unit cluster.  However, since we do not model functional units in our simulator, we simply divide the ROB 
into multiple smaller ROBs.  Each of these ROBs implement the default circular ROB model.  Each ROB segment 
features a head pointer and a tail pointer and contains an equal portion of the total ROB space.  No further 
enhancements need to be made to the default circular ROB model as the inclusion of the central FIFO allows the 
ROB clusters to operate entirely independently of one another.  Each ROB might as well be the only ROB in the system.

The largest modeling change required by the distributed ROB is the inclusion of the central FIFO queue.  This 
FIFO queue is responsible for maintaining a global state of order between instructions in the different ROB buffers. 
The FIFO queue is modeled as a modified circular buffer and each entry contains: 1) PC of instruction, 
2) ROB{\char'137}ID where the instruction placeholder resides, and 3) a valid bit corresponding to the completion state of 
the instruction.  The total number of entries in the FIFO queue is equal to the total entry size of the ROB.  
Each active entry in the FIFO queue correlates with an active entry in one of the ROB clusters.

The distributed ROB requires only minor changes to the basic circular buffer to realize its effectiveness.  Instructions are 
read from the Instruction Queue as usual and inserted into their corresponding ROB entry while updating the corresponding tail 
pointer.  Since there are now multiple ROBs into which instructions can be inserted, a quick modulus is performed on the 
destination register of the instruction to divide the instructions equally between each ROB.  This presents the possibility 
for a worst-case situation where the entire instruction stream is blocked due to improper balancing between ROB clusters.  
This situation would occur very rarely in practice and would induce no form of deadlock or livelock in the system -- 
it would merely affect performance.  Possible avoidance techniques can be researched as future work.

At the same time as inserting the instruction from the IQ into the ROB cluster, an entry must be recorded in the central FIFO 
queue to maintain global program order.  At this point, both the instruction PC and the recently determined ROB{\char'137}ID are 
noted in the central FIFO queue.  The tail pointer of the central FIFO queue is increased accordingly to match.

Upon completion of an instruction, an instruction must be flagged as valid in both the ROB cluster in which it resides as well 
as its corresponding central FIFO queue entry.  The central FIFO queue then directs the appropriate ROB to write the instruction 
output to its write port when all prior instructions have been written so that program order is maintained.  This is the second 
major behavioral difference between the circular ROB and the distributed ROB, as writes and write-order is driven and determined by 
the central FIFO queue instead of the ROB cluster itself.

\subsubsection{Dynamically Sized ROB}
The dynamically sized ROB is an optimization on the circular ROB where the buffer is divided into partitions, and those partitions deemed unnecessary are powered off to conserve energy. Our implementation of the basic dynamic ROB is behaviorally the same as that detailed in \cite{kucuk2}.

Downsizing the ROB is based on a periodic sampling of the buffer state. A number of cycles, {\it update\_period}, is defined, and every {\it update\_period} cycles the size of the ROB is updated. Every {\it sample\_period} cycles, the number of active entries is recorded, and the quantity {\it active\_size} is recorded as the average of these samples between each {\it update\_period}. The current maximum capacity of the reorder buffer is defined as the {\it current\_size}, and the size of a single ROB partition is defined as {\it partition\_size}. After each {\it update\_period} cycles, the ROB checks how full the buffer is, and if {\it current\_size} - {\it active\_size} > {\it partition\_size}, then a single partition is turned off.

Increasing the size of the ROB is dependent on how often the buffer fills up. A value, defined as the {\it overflow\_count} is incremented each cycle that instruction issue is stalled because the ROB is full. An {\it overflow\_threshold} is defined, and if during an {\it update\_period}, {\it overflow\_count} > {\it overflow\_threshold}, then a partition is turned on.

A conservative approach of deallocating strictly from one end of the buffer, and allocating strictly towards that end is followed. For example, if the ROB consisted of four partitions, downsizing the buffer would turn off the fourth, and then the third. At this point, if we were to increase the size of the buffer, the third partition would be turned on first, and the fourth partition next.

When the ROB's size is updated, partitions are not immediately turned off, since the partition could coincide with the active region of the buffer. If deallocating, this would disable a part of the active region, making instructions in the region impossible to complete. If allocating, this would insert an empty segment right into the middle of the FIFO queue, logically breaking the buffer. It is only safe to turn a partition off/on when the active region is not wrapped around (that is, the head pointer is less than the tail pointer), and the tail pointer comes before the first index into the partition to turn on/off. \cite{kucuk3} details a worst case scenario of deallocating, where the active region is wrapped (head is greater than tail), and both pointers lie in the region to be turned off. In this case, the head pointer must first clear the entries already in the buffer. Since the tail pointer is already in the partition, it will continue to write into the partition, and the head pointer must come around a second time to clear out these new entries. Only then can the partition be safely deallocated, after the head pointer has wrapped around to zero twice.

\subsubsection{Optimized Dynamic ROB}
We implement an optimized version of the dynamic ROB detailed in \cite{kucuk2} and \cite{kucuk3}, opting to be more aggressive with turning off partitions in order to save power, and to improve the partition allocation/deallocation scheme.

When downsizing the regular dynamic ROB, a single partition is turned off if {\it current\_size} - {\it active\_size} > {\it partition\_size}. In an effort to minimize the buffer size and thus the power consumed, we choose to downsize by

\begin{center}
$\lfloor \frac{current\_size - active\_size}{partition\_size} \rfloor$
\end{center}

Thus if much less than the capacity of the buffer is active, several partitions may be turned off at once.

To more efficiently resize the buffer, we found that choosing the first valid partition immediately after the one housing the tail pointer is optimal. For example, if we are allocating a partition in a four partition buffer, and the tail pointer resides in partition 3, we first check if partition 4 can be turned on, then partition 1, then partition 2.

In allocating the buffer, the reasoning is as follows. We aim to enable a disabled partition. It is impossible for the head pointer to lie in the disabled partition, thus there is no need to consider the head pointer. To maintain the logical consistency of the buffer, we cannot enable a partition that would divide the active region of the buffer. Thus, we must allocate a partition that comes after the tail pointer, and if we wish the buffer size to grow as soon as possible, the partition to be allocated must be the first disabled partition after the tail pointer.

In deallocating the buffer, turning off the first partition after the tail pointer is also optimal. Being after the tail pointer, once we mark the partition for disabling we can prevent the tail pointer from writing into it. This prevents the worst case scenario where the head pointer must wrap around 0 twice to disable a partition. The next enabled partition immediately after the tail pointer is also the one the active region will leave first (if the region is in the partition; if not the partition is already empty and ready to deactivate), thus it will empty out first. If we are to maximize the power savings, we must disable the first partition that is available to be disabled, thus the first active partition after the tail is the optimal choice.

One last optmization we implemented was just-in-time allocation of partitions. Normally, when the tail pointer increments, if a partition is not active the tail pointer simply skips it. In our optimization, before skipping the tail pointer will check if the disabled partition is marked to be enabled. If so, instead of skipping over it, the partition will be turned on and the tail pointer will write into it. This ensures that each partition to be turned on is enabled at the last possible moment, saving the energy that would otherwise be consumed by an empty partition waiting to be written into.

\subsubsection{Reduced Output Port ROB (Retention Latch)}

\subsection{Power Tabulation Implementation}

### 2011-06-08 paper/start.tex
multi file paper

\begin{abstract}
        In an effort to increase instruction level parallelism techniques such
as pipelining, out of order execution and operand forwarding.  However
these efforts come at a large power cost.  Out of order execution requires
a ReOrder Buffer (ROB), to commit the registers in program order.  This
structure has grown to be monolithic with numerous entries and read/write ports.
There is no current methods to efficiently and equally compare various
ROB advancements to improve power efficiency. We propose a loosely cycle
accurate model that keeps track of the power usages of the ROB.  Using the 
simulation tool, we are able to compare and contrast the affects of improvements
made on the ROB.  This also allows for a first line test to validate proof of
concepts for future power efficient ROB designs.
\end{abstract}

\terms{Power, Energy, Architecture}

\keywords{Power efficiency, ROB, Simulation} % NOT required for Proceedings

\section{Introduction}
BLAH BLAH BLAH INTRO WEEEE

### 2011-06-08 paper/bg.tex
added 1 ref ans an img

@@
 An accepted optimization for improving the power consumption of an ROB unit 
-is to break up the ROB buffer into clusters REF[XXX].  In this organization, the in-flight 
+is to break up the ROB buffer into clusters \cite{rabaey}.  In this organization shown in Figure ~\ref{fig:distrob}, 
+the in-flight 
 instruction information is distributed across all of the units and then re-assembled 
@@
 
-...IMAGE FROM OUR SLIDES...
+\begin{figure*}
+\centering
+\includegraphics[height=2in]{clusterROB.png}
+\caption{Distributed ROB layout}
+\label{fig:distrob}
+\end{figure*}
 
@@
 sequentially.  It is reported, however, that this happens very rarely in practice and induces less than 1\% 
-performance degradation REF[XXX].
+performance degradation \cite{rabaey}.
 \subsection{Latch ROB}

### 2011-06-08 paper/proj.tex
paper, rob latch update

@@
 
-One last optmization we implemented was just-in-time allocation of partitions.
+One last optimization we implemented was just-in-time allocation of partitions.
 Normally, when the tail pointer increments, if a partition is not active the tail pointer simply skips it.
@@
 
-\subsubsection{Reduced Output Port ROB (Retention Latch)}
+\subsubsection{Omitted Output Port ROB (Retention Latch)}
+The main difference between the Circular ROB and the Omitted Output Port (OOP) ROB is that the OOP adds in a retention latches and is not allowed to forward any results directly to DU.  
+This our implementation of the OOP ROB simply extends the circular ROB, adding in a structure for the retention latches.  The latches is modeled very similar to the ROB entries, an array
+of a latch entry structure.  Each latch entry contains a 32 bit value for the results to be stored as well small tag to the corresponding destination register of the result.  As an artifact of our 
+simulation a flag to denote that the entry is part of a floating point operation is required as well.  Also an additional counter was added to count the additional re-forwarding when a value
+is committed to the ARF but also needs to be forwarded to the DU unit as well.  Though this is not an increasing number of ports driven, it may affect future power calculations relating to 
+capacitance.  Lastly as changing the semantic of the number of forwards to the DU, the count refers to the number time the entry is read from a retention latch.
 

### 2011-06-08 paper/proj.tex
more paper

@@
 
+As a few minor changes to the behavioral function of the OOP ROB, the first of which is the lack of the ability to forward operands.  Thus in the simulated stage of reading from the instruction
+stream, when we would normally check if the operand is in the ROB and forward.  We instead do the following, we still check to see if the operand is in the ROB, and the retention latches. 
+If the entry is not in the ROB, or the latches, the operand would be available in the ARF or memory, in the converse case,  the entry in both, the ROB, and the latches, the operand is forwarded
+from the latches and the counter is incremented .
+In the case that the entry exists in the ROB, but not in the latches, a flag would be set in the ROB to forward the result when committing to the ARF.
+The last case, that is the entry exists in the latches and not in the ROB, means that the result has already been committed and the result could be forwarded from the ARF.  Next in our update entries 
+(when the cycle counter expires -- execution has completed) we would add in the new entry to the retention latch.  Lastly when the ROB entires are valid, and ready to be committed to the ARF,
+if the forwarding flag is set, the forwarding counter is incremented.
 \subsection{Power Tabulation Implementation}

### 2011-06-08 paper/proj.tex
paper stuff.

@@
 \subsection{Power Tabulation Implementation}
+In order for our simulation to determine the power utilization, we do not try and equate each event with an associated cost in joules.  As technology changes, and new techniques are 
+developed to decease circuit level power usages, our simulation will become less and less accurate.  Instead we chose to simply tabulate each event that is associated with 
+power cost, such as bit flips from 0 to 1, register comparator usage, and bit lines driven.  To help measure leakage power and overall states of the ROB, we also keep track of the 
+time (in cycles) each bit stays in the high state, or the low state.
+
+At the start of each cycle

### 2011-06-08 paper/proj.tex
paper

@@
 
-At the start of each cycle
+At the start of each cycle the simulator take a snapshot of the entire ROB, this includes each ROB entry, head/tail pointers (circular ROB), forward flags (OOP ROB), and etc as the type 
+of ROB changes.  Then at the end of each cycle completion, we are able to compare which bits have changed, from low to high or high to low.   This results of the comparisons are tabulated
+and printed out at the of the simulation runs.  The checkpoint comparison allows the simulator to easily modularize power data collection for any type of ROB.
+
+The events that cannot be captured by the checkpoint comparison, is captured during the behavioral simulation.  These items are stateless such as comparators used, bit lines driven, and etc.
+As part of the behavioral simulation, the respective counters would simply be incremented as part of simulating the usage of the item.
+
+These final tabulations allow the user to fairly comparing the type and algorithm of the reorder buffer with our worrying about the circuit level technology underneath.

### 2011-06-08 paper/start.tex
paper

@@
 \section{Introduction}
-BLAH BLAH BLAH INTRO WEEEE
+Out of order execution is a common technique used to aggressively push instruction
+throughput to meet the demands of modern computer users.  As the name suggests
+this allows for instructions to be inflight while previous instructions may still be waiting 
+for an available execution unit or operands.  To accomplish this, two additional structures/stages
+are necessary  on top of the traditional pipelined architecture.  Both of these additional
+structures are essentially memory elements. The first one to keep track of which instructions
+are ready to be executed and the other known as the ROB, allows the processor  to commit the
+results of execution in program order to maintain program correctness.
+
+Through out its evolution the the ROB has grown to a large monolithic structure that consumes
+on average 7\% of the overall processor power budget\cite{rabey}.  Various new implementations
+of the ROB have been proposed in an effort to reduce power consumption.  Some of which 
+dynamically change the the size of the ROB, or a banked buffer structure and other proposals argues in the 
+extreme case to eliminate read ports all together\cite{kucuk}\cite{kucuk2} \cite{kucuk3}.
+
+With the rapid change of manufacturing technologies and other impacts on energy utilization it 
+becomes hard to determine the gain of each new proposed implementation.  We propose a new
+a loosely cycle accurate ROB simulator that provides a fair comparison of  different ROB designs.
+ROB\_pwr is a simulator that tabulates each action that would cost power.  This allows for a fair
+comparison between each ROB design, without the concern for the underlying technology.

### 2011-06-08 src/rob_latch.h
Fix stray "m" in file src/rob_latch.h

@@ -40,8 +40,8 @@ class ROB_Latch : public ROB_Circ {
 
     void pre_cycle_power_snapshot();
     void post_cycle_power_tabulation();
-	void print_power_stats (int cycles);
-	uint32_t getCyclesToCompletion(uint32_t reg);m
+    void print_power_stats (int cycles);
+    uint32_t getCyclesToCompletion(uint32_t reg);
 };
 
 #endif /* #ifndef ROB_LATCH_H */

### 2011-06-08 paper/proj.tex
spell check

@@
 
-We model the input to our simulator as a stream of integer and floatpoint instructions.
+We model the input to our simulator as a stream of integer and floating point instructions.
 The instructions are represented at the bit-level and model MIPS instructions as closely as possible.
@@
 \subsection{Model Implementation}
-The following section details the model implementatations in our simulator.
+The following section details the model implementations in our simulator.
 The basic re-order buffer model we use as a base-line for comparison is a standard circular ROB model.

### 2011-06-08 src/main.cpp
In main, document usage of optimized dynamic buffer.

@@ -148,6 +148,7 @@ void help ()
   cout << "Modes:" << endl;
   cout << "  circular   Circular ROB." << endl;
   cout << "  dynamic    Dynamically resized ROB." << endl;
+  cout << "  dyn-opt  Optimized dynamic ROB." << endl;
   cout << "  dist       Distributed ROB." << endl;
   cout << "  latch      ROB with retention latches." << endl;
   cout << "Benchmarks:" << endl;

### 2011-06-08 paper/bg.tex
paper, rob latch update
Paper stuff. Put printouts behind m_print flag (enable with -d)

@@
 \subsection{Dynamic ROB}
-The dynamic ROB aims to reduce power by turning off portions of the buffer when they
-are not needed, and turning them back on when they become necessary. By partitioning
-the buffer into what is essentially smaller buffers, the complexity of each partition
-is reduced. Further, the ROB will then be custom sized to need, thus reducing energy
-wasted powering large stretches of empty entries. Due to the large percentage of energy
-consumption stemming from leakage power, turning off partitions contribute significant
-power savings over the base circular ROB.
+
+The dynamic ROB aims to reduce power usage by turning off portions of the buffer when they are not needed, and turning them back on when they become necessary.
+By partitioning the buffer into what is essentially smaller buffers, the complexity of each partition is reduced.
+Further, the ROB will then be custom sized to need, thus reducing energy wasted powering large stretches of empty entries.
+Due to the large percentage of energy consumption stemming from leakage power, turning off partitions contribute significant power savings over the base circular ROB.
+
+Downsizing the ROB is based on a periodic sampling of the buffer state.
+A number of cycles, {\it update\_period}, is defined, and every {\it update\_period} cycles the size of the ROB is updated.
+Every {\it sample\_period} cycles, the number of active entries is recorded, and the quantity {\it active\_size} is recorded as the average of these samples between each {\it update\_period}.
+The current maximum capacity of the reorder buffer is defined as the {\it current\_size}, and the size of a single ROB partition is defined as {\it partition\_size}.
+After each {\it update\_period} cycles, the ROB checks how full the buffer is, and if {\it current\_size} - {\it active\_size} > {\it partition\_size}, then a single partition is turned off.
+
+Increasing the size of the ROB is dependent on how often the buffer fills up.
+A value, defined as the {\it overflow\_count} is incremented each cycle that instruction issue is stalled because the ROB is full.
+An {\it overflow\_threshold} is defined, and if during an {\it update\_period}, {\it overflow\_count} > {\it overflow\_threshold}, then a partition is turned on.
 
@@
 percentage of instructions.  It has been proposed by eliminating the read ports 
-in the ROB all together, we can reduce energy consumption\cite{kucuk}.  However
-without read ports, a small window appears where the data cannot be accessed,
+in the ROB all together, we can reduce energy consumption\cite{kucuk}.
+
+However without read ports, a small window appears where the data cannot be accessed,
 starting from when the result is placed in to the ROB until the ROB commits that 

### 2011-06-08 paper/proj.tex
Paper stuff. Put printouts behind m_print flag (enable with -d)

@@
 \subsubsection{Distributed (Cluster based) ROB}
-Our distributed (cluster based) ROB is modeled similarly to the circular ROB.  Instead of having a single ROB buffer, 
-numerous ROB buffers are instantiated.  In a real system, there would be one ROB buffer for each functional 
-unit cluster.  However, since we do not model functional units in our simulator, we simply divide the ROB 
-into multiple smaller ROBs.  Each of these ROBs implement the default circular ROB model.  Each ROB segment 
-features a head pointer and a tail pointer and contains an equal portion of the total ROB space.  No further 
-enhancements need to be made to the default circular ROB model as the inclusion of the central FIFO allows the 
-ROB clusters to operate entirely independently of one another.  Each ROB might as well be the only ROB in the system.
+Our distributed (cluster based) ROB is modeled similarly to the circular ROB.
+Instead of having a single ROB buffer, numerous ROB buffers are instantiated.
+In a real system, there would be one ROB buffer for each functional unit cluster.
+However, since we do not model functional units in our simulator, we simply divide the ROB into multiple smaller ROBs.
+Each of these ROBs implement the default circular ROB model.
+Each ROB segment features a head pointer and a tail pointer and contains an equal portion of the total ROB space.
+No further enhancements need to be made to the default circular ROB model as the inclusion of the central FIFO allows the ROB clusters to operate entirely independently of one another.
+Each ROB might as well be the only ROB in the system.
 
-The largest modeling change required by the distributed ROB is the inclusion of the central FIFO queue.  This 
-FIFO queue is responsible for maintaining a global state of order between instructions in the different ROB buffers. 
-The FIFO queue is modeled as a modified circular buffer and each entry contains: 1) PC of instruction, 
-2) ROB{\char'137}ID where the instruction placeholder resides, and 3) a valid bit corresponding to the completion state of 
-the instruction.  The total number of entries in the FIFO queue is equal to the total entry size of the ROB.  
+The largest modeling change required by the distributed ROB is the inclusion of the central FIFO queue.
+This FIFO queue is responsible for maintaining a global state of order between instructions in the different ROB buffers.
+The FIFO queue is modeled as a modified circular buffer and each entry contains: 1) PC of instruction, 2) ROB{\char'137}ID where the instruction placeholder resides, and 3) a valid bit corresponding to the completion state of the instruction.
+The total number of entries in the FIFO queue is equal to the total entry size of the ROB.
 Each active entry in the FIFO queue correlates with an active entry in one of the ROB clusters.
 
-The distributed ROB requires only minor changes to the basic circular buffer to realize its effectiveness.  Instructions are 
-read from the Instruction Queue as usual and inserted into their corresponding ROB entry while updating the corresponding tail 
-pointer.  Since there are now multiple ROBs into which instructions can be inserted, a quick modulus is performed on the 
-destination register of the instruction to divide the instructions equally between each ROB.  This presents the possibility 
-for a worst-case situation where the entire instruction stream is blocked due to improper balancing between ROB clusters.  
-This situation would occur very rarely in practice and would induce no form of deadlock or livelock in the system -- 
-it would merely affect performance.  Possible avoidance techniques can be researched as future work.
+The distributed ROB requires only minor changes to the basic circular buffer to realize its effectiveness.
+Instructions are read from the Instruction Queue as usual and inserted into their corresponding ROB entry while updating the corresponding tail pointer.
+Since there are now multiple ROBs into which instructions can be inserted, a quick modulus is performed on the destination register of the instruction to divide the instructions equally between each ROB.
+This presents the possibility for a worst-case situation where the entire instruction stream is blocked due to improper balancing between ROB clusters.
+This situation would occur very rarely in practice and would induce no form of deadlock or livelock in the system -- it would merely affect performance.
+Possible avoidance techniques can be researched as future work.
 
-At the same time as inserting the instruction from the IQ into the ROB cluster, an entry must be recorded in the central FIFO 
-queue to maintain global program order.  At this point, both the instruction PC and the recently determined ROB{\char'137}ID are 
-noted in the central FIFO queue.  The tail pointer of the central FIFO queue is increased accordingly to match.
+At the same time as inserting the instruction from the IQ into the ROB cluster, an entry must be recorded in the central FIFO queue to maintain global program order.
+At this point, both the instruction PC and the recently determined ROB{\char'137}ID are noted in the central FIFO queue.
+The tail pointer of the central FIFO queue is increased accordingly to match.
 
-Upon completion of an instruction, an instruction must be flagged as valid in both the ROB cluster in which it resides as well 
-as its corresponding central FIFO queue entry.  The central FIFO queue then directs the appropriate ROB to write the instruction 
-output to its write port when all prior instructions have been written so that program order is maintained.  This is the second 
-major behavioral difference between the circular ROB and the distributed ROB, as writes and write-order is driven and determined by 
-the central FIFO queue instead of the ROB cluster itself.
+Upon completion of an instruction, an instruction must be flagged as valid in both the ROB cluster in which it resides as well as its corresponding central FIFO queue entry.
+The central FIFO queue then directs the appropriate ROB to write the instruction output to its write port when all prior instructions have been written so that program order is maintained.
+This is the second major behavioral difference between the circular ROB and the distributed ROB, as writes and write-order is driven and determined by the central FIFO queue instead of the ROB cluster itself.
 
@@
 Our implementation of the basic dynamic ROB is behaviorally the same as that detailed in \cite{kucuk2}.
-
-Downsizing the ROB is based on a periodic sampling of the buffer state.
-A number of cycles, {\it update\_period}, is defined, and every {\it update\_period} cycles the size of the ROB is updated.
-Every {\it sample\_period} cycles, the number of active entries is recorded, and the quantity {\it active\_size} is recorded as the average of these samples between each {\it update\_period}.
-The current maximum capacity of the reorder buffer is defined as the {\it current\_size}, and the size of a single ROB partition is defined as {\it partition\_size}.
-After each {\it update\_period} cycles, the ROB checks how full the buffer is, and if {\it current\_size} - {\it active\_size} > {\it partition\_size}, then a single partition is turned off.
-
-Increasing the size of the ROB is dependent on how often the buffer fills up.
-A value, defined as the {\it overflow\_count} is incremented each cycle that instruction issue is stalled because the ROB is full.
-An {\it overflow\_threshold} is defined, and if during an {\it update\_period}, {\it overflow\_count} > {\it overflow\_threshold}, then a partition is turned on.
 
@@
 The main difference between the Circular ROB and the Omitted Output Port (OOP) ROB is that the OOP adds in a retention latches and is not allowed to forward any results directly to DU.
-This our implementation of the OOP ROB simply extends the circular ROB, adding in a structure for the retention latches.  The latches is modeled very similar to the ROB entries, an array
-of a latch entry structure.  Each latch entry contains a 32 bit value for the results to be stored as well small tag to the corresponding destination register of the result.  As an artifact of our 
-simulation a flag to denote that the entry is part of a floating point operation is required as well.  Also an additional counter was added to count the additional re-forwarding when a value
-is committed to the ARF but also needs to be forwarded to the DU unit as well.  Though this is not an increasing number of ports driven, it may affect future power calculations relating to 
-capacitance.  Lastly as changing the semantic of the number of forwards to the DU, the count refers to the number time the entry is read from a retention latch.
+This our implementation of the OOP ROB simply extends the circular ROB, adding in a structure for the retention latches.
+The latches is modeled very similar to the ROB entries, an array
+of a latch entry structure.
+Each latch entry contains a 32 bit value for the results to be stored as well small tag to the corresponding destination register of the result.
+As an artifact of our simulation a flag to denote that the entry is part of a floating point operation is required as well.
+Also an additional counter was added to count the additional re-forwarding when a value is committed to the ARF but also needs to be forwarded to the DU unit as well.
+Though this is not an increasing number of ports driven, it may affect future power calculations relating to capacitance.
+Lastly as changing the semantic of the number of forwards to the DU, the count refers to the number time the entry is read from a retention latch.
 
-As a few minor changes to the behavioral function of the OOP ROB, the first of which is the lack of the ability to forward operands.  Thus in the simulated stage of reading from the instruction
-stream, when we would normally check if the operand is in the ROB and forward.  We instead do the following, we still check to see if the operand is in the ROB, and the retention latches. 
-If the entry is not in the ROB, or the latches, the operand would be available in the ARF or memory, in the converse case,  the entry in both, the ROB, and the latches, the operand is forwarded
-from the latches and the counter is incremented .
+As a few minor changes to the behavioral function of the OOP ROB, the first of which is the lack of the ability to forward operands.
+Thus in the simulated stage of reading from the instruction stream, when we would normally check if the operand is in the ROB and forward.
+We instead do the following, we still check to see if the operand is in the ROB, and the retention latches.
+If the entry is not in the ROB, or the latches, the operand would be available in the ARF or memory, in the converse case,  the entry in both, the ROB, and the latches, the operand is forwarded from the latches and the counter is incremented .
 In the case that the entry exists in the ROB, but not in the latches, a flag would be set in the ROB to forward the result when committing to the ARF.
-The last case, that is the entry exists in the latches and not in the ROB, means that the result has already been committed and the result could be forwarded from the ARF.  Next in our update entries 
-(when the cycle counter expires -- execution has completed) we would add in the new entry to the retention latch.  Lastly when the ROB entires are valid, and ready to be committed to the ARF,
+The last case, that is the entry exists in the latches and not in the ROB, means that the result has already been committed and the result could be forwarded from the ARF.
+Next in our update entries (when the cycle counter expires -- execution has completed) we would add in the new entry to the retention latch.
+Lastly when the ROB entires are valid, and ready to be committed to the ARF,
 if the forwarding flag is set, the forwarding counter is incremented.
+
 \subsection{Power Tabulation Implementation}
-In order for our simulation to determine the power utilization, we do not try and equate each event with an associated cost in joules.  As technology changes, and new techniques are 
-developed to decease circuit level power usages, our simulation will become less and less accurate.  Instead we chose to simply tabulate each event that is associated with 
-power cost, such as bit flips from 0 to 1, register comparator usage, and bit lines driven.  To help measure leakage power and overall states of the ROB, we also keep track of the 
-time (in cycles) each bit stays in the high state, or the low state.
+In order for our simulation to determine the power utilization, we do not try and equate each event with an associated cost in joules.
+As technology changes, and new techniques are developed to decease circuit level power usages, our simulation will become less and less accurate.
+Instead we chose to simply tabulate each event that is associated with power cost, such as bit flips from 0 to 1, register comparator usage, and bit lines driven.
+To help measure leakage power and overall states of the ROB, we also keep track of the time (in cycles) each bit stays in the high state, or the low state.
 
-At the start of each cycle the simulator take a snapshot of the entire ROB, this includes each ROB entry, head/tail pointers (circular ROB), forward flags (OOP ROB), and etc as the type 
-of ROB changes.  Then at the end of each cycle completion, we are able to compare which bits have changed, from low to high or high to low.   This results of the comparisons are tabulated
+At the start of each cycle the simulator take a snapshot of the entire ROB, this includes each ROB entry, head/tail pointers (circular ROB), forward flags (OOP ROB), and etc as the type of ROB changes.
+Then at the end of each cycle completion, we are able to compare which bits have changed, from low to high or high to low.
+This results of the comparisons are tabulated
 and printed out at the of the simulation runs.  The checkpoint comparison allows the simulator to easily modularize power data collection for any type of ROB.

### 2011-06-08 src/rob_circ.cpp

@@ -16,7 +16,6 @@ ROB_Circ::ROB_Circ (int s, int in, int fn, int pflags)
   m_head_size = (uint32_t)(ceil(log(m_size) / log(2)));
   p_bit_count += (2*m_head_size);
 
-  cout << "New ROB_Circ:" << endl;
   cout << "# Entries: " << m_size << endl;
   cout << "Entry size: 70 bits" << endl;
   cout << "Head and tail pointer bits: " << m_head_size << endl;
@@ -208,7 +207,7 @@ uint32_t ROB_Circ::getCyclesToCompletion(uint32_t reg) {
         return entry->cycles;
       i = ((i+1)%m_size);
     } while(i != m_tail);
-  cout << "Couldnt find cycles to completion when expected!" << endl;
+  cerr << "Couldnt find cycles to completion when expected!" << endl;
   exit(1);
 }
 

### 2011-06-08 src/rob_dist.cpp

@@ -105,7 +105,8 @@ void ROB_Dist::run (ins_t ins[])
       update_power_totals();
 
       cycles++;
-      print_msgs (cycles);
+      if (m_print & DBG_FLAG)
+        print_msgs (cycles);
 
       // Break when we've written all the instructions.
       if (fifo_empty && ins[ins_num].type == -1)
@@ -129,7 +130,8 @@ void ROB_Dist::update_entries ()
     fifo_entry_t *fifo_entry = return_fifo_entry (i);
     entry_t *entry = get_entry(fifo_entry->rob_id, fifo_entry->rob_index);
     if (entry->cycles && --entry->cycles == 0) {
-      cout << "Got result for central_fifo[" << i << "] -> rob[" << fifo_entry->rob_id << "][" << fifo_entry->rob_index << "]" << endl;
+      if (m_print & DBG_FLAG)
+        cout << "Got result for central_fifo[" << i << "] -> rob[" << fifo_entry->rob_id << "][" << fifo_entry->rob_index << "]" << endl;
       entry->valid = true;
       fifo_entry->valid = true;
       m_nrex++;
@@ -165,7 +167,8 @@ void ROB_Dist::write_to_arf ()
         isFirstFp = true;
       }
     }
-    cout << "Writing central_fifo[" << fifo_head << "] -> rob[" << fifo_entry->rob_id << "][" << fifo_entry->rob_index << "] to ARF" << endl;
+    if (m_print & DBG_FLAG)
+      cout << "Writing central_fifo[" << fifo_head << "] -> rob[" << fifo_entry->rob_id << "][" << fifo_entry->rob_index << "] to ARF" << endl;
     m_head[fifo_entry->rob_id] = ptr_incr (m_head[fifo_entry->rob_id]);
     if(m_head[fifo_entry->rob_id] == m_tail[fifo_entry->rob_id])
       m_empty[fifo_entry->rob_id] = true;
@@ -195,18 +198,20 @@ int ROB_Dist::read_from_iq (int ins_num, ins_t ins[])
   int clusterId = ((ins[ins_num].regs)&mask)%m_cluster_count;
   while ((m_empty[clusterId] || (m_head[clusterId] != m_tail[clusterId]))
          && (ins[ins_num].type != -1) && (nentries < m_n)) {
[…]

### 2011-06-08 paper/proj.tex
circ imple

--- 2011-06-08 paper/proj.tex
@@
 \label{sec:circ}
-The circular ROB is the basic current implementation of the reorder buffer.
-It is implemented as a circular FIFO buffer, and the beginning and end of the used region are tracked by the head and tail pointers, respectively.
-When an instruction is issued, space is reserved for the instruction's result at the tail pointer, and the tail pointer is incremented.
-When the instruction completes, the result is inserted into the reorder buffer at the reserved space.
-The result is finally committed to the architecture register file when the head pointer passes through it; each cycle, results are committed from the head pointer and the head pointer is incremented repeatedly, until either an unfinished instruction result slot is reached or all write ports are used up.
+The circular ROB is the basic current implementation of the reorder buffer, so far each of the other ROB designs will use the 
+circular array for each instruction entry.  Each instruction entry will have a flag to determine if its result is valid, a 
+result field (currently 32 bits), the corresponding register id tag and the PC of the instructions.  There are two additional
+fields in each entry that is an artifact of our simulation, the first of which is a flag to denote a floating point entry, and 
+mark that the result of the next entry is a continuation of the current one.  The second element that is an artifact of the
+simulator is the cycles counter, which acts as a timing simulator and will determine when the result will finally be produced
+by the execution unit.  The ROB is simply an array of these entry structures, and two pointers (32 bits) that will point
+to the head or tail of the circular FIFO buffer.  Also the variables used for power tabulation are a part of the structure as 
+another artifact of our simulator.
 
-The circular ROB is considered full when the tail pointer reaches the head pointer, as the head pointer denotes the start of the in-use region and no more space can be reserved from the tail pointer.
-Conversely, then buffer is empty when the head pointer catches up to the tail pointer, as every instruction in the buffer has now been committed to the ARF.
+The behavioral simulation is broken down into three main sections.  First is the update stage, which will cycle 
+through each entry of the ROB, and decrement the cycles counter.  When it reaches zero, the valid flag is marked and the
+result is stored into the result element of the entry.  Next is the ARF stage, where the top N entries are checked
+for its valid flag (where N is the ways the processor is super scalar).  The entries are committed to the ARF if it is at 
+the head of the buffer and its valid flag is set.
+Lastly we simulate the interaction with the instruction queues.  Place holder spots are written at the tail of the buffer
+at most N times. The place holder entry will unmark the valid flag (if it was previously set), and insert the pipeline cycle
+time until execution completion. Also in this stage we check if operand forwarding is needed for any instructions that is 
+requested by the IQ.  These three stages will repeat until the end of the instruction stream.
 

### 2011-06-08 src/main.cpp
Fix parse error in main for optimized dynamic ROB.

@@ -56,7 +56,7 @@ int main (int argc, char *argv[])
             mode = MODE_CIRC;
           else if (strcmp (optarg, "dynamic") == 0)
             mode = MODE_DYN;
-          else if (strcmp (optarg, "dynamic-opt") == 0)
+          else if (strcmp (optarg, "dyn-opt") == 0)
             mode = MODE_DYNOPT;
           else if (strcmp (optarg, "dist") == 0)
             mode = MODE_DIST;

### 2011-06-08 paper/proj.tex
Implementation section on dynamic ROB

@@
 \subsubsection{Dynamically Sized ROB}
-The dynamically sized ROB is an optimization on the circular ROB where the buffer is divided into partitions, and those partitions deemed unnecessary are powered off to conserve energy.
-Our implementation of the basic dynamic ROB is behaviorally the same as that detailed in \cite{kucuk2}.
+Our implementation of the basic dynamic ROB is behaviorally the same as that detailed in \cite{kucuk2} and the background section.
 
-A conservative approach of deallocating strictly from one end of the buffer, and allocating strictly towards that end is followed.
+For {\it P} partitions, instead of allocating one region of memory as in the circular ROB we allocate {\it P} regions, and associate with each region two status bits, denote the four states a ROB partition can have: {\tt ENABLED}, {\tt DISABLED}, {\tt TO\_ENABLE}, and {\tt TO\_DISABLE}.
+Also, where pointer incrementation in the circular buffer was done as:
+
+\begin{center}
+$ new\_ptr = (old\_ptr + 1) \: \% \: ROB\_size $
+\end{center}
+
+Which is no long accurate in the dynamic ROB.
+The pointer incrementation module for the dynamic ROB was updated to check if the incremented pointer is stepping into a new partition, and to step over that partition if necessary.
+When incrementing the head pointer, only partitions labeled {\tt ENABLED} and {\tt TO\_DISABLE} are valid.
+When incrementing the tail pointer, only partitions labeled {\tt ENABLED} are valid.
+
+The entry fetch module also requires updating.
+Whereas in the circular buffer, an entry is located simply by taking the head or tail pointer as the offset into the buffer, the dynamic buffer must now map the pointer to a particular partition and an offset into that partition.
+
+At cycle completion, a module was added to perform sampling and resizing.
+Every cycle, the module checks for a stall due to the buffer being full, and if so the {\it overflow\_count} is incremented.
+Every {\it sample\_period} cycles, the buffer's {\it active\_size} is sampled.
+Every {\it update\_period}, the buffer's size is updated by marking the appropriate partitions {\tt TO\_ENABLE} or {\tt TO\_DISABLE}.
+Each cycle, the head and tail pointers are checked against the marked partitions to see if their state can be updated to {\tt ENABLED} or {\tt DISABLED}.
+
+\cite{kucuk3} takes a conservative approach of deallocating strictly from one end of the buffer, and also allocating strictly towards that end.
 For example, if the ROB consisted of four partitions, downsizing the buffer would turn off the fourth, and then the third.

### 2011-06-08 paper/end.tex
table

@@
 \section{Results}
-hi
-\subsection{subsection titl}
-sub section.. with table!
+
 \begin{table*}
 \centering
-\caption{Some Typical Commands}
-\begin{tabular}{|c|c|l|} \hline
-Command&A Number&Comments\\ \hline
-\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
-\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
-\texttt{{\char'134}table}& 300 & For tables\\ \hline
-\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
+\caption{FP Heavy Instruction Profile Results(B1)}
+\begin{tabular}{|c|c|c|c|c|c|} \hline
+    & Circular  & Distrubuted & Dynamic & Dynamic (optimized) & OOP\\ \hline
+Cycles&82,722	&50,541&    	84,211	&84,078&        	82,153\\ \hline
+Bit Cycles&1,483,701,792&	1,399,581,372&	1,453,144,453	&1,470,555,825	&1,522,130,784\\ \hline
+Trans. to High&1,967,483&	1,719,957	&1,942,948	&1,930,476	&3,235,293\\ \hline
+Reg Comp&6,725,088&	6,631,913&	6,710,777&	6,718,043	&11,160,291\\ \hline
+\end{tabular}
 \end{table*}
-% end the environment with {table*}, NOTE not {table}!
+
+\begin{table*}
+\centering
+\caption{int Heavy Instruction Profile Results(B2)}
+\begin{tabular}{|c|c|c|c|c|c|} \hline
+    & Circular  & Distrubuted & Dynamic & Dynamic (optimized) & OOP\\ \hline
+Cycles&25,781&	25,794&	26,079	&25,783	&25,780\\ \hline
+Bit Cycles&462,408,016	&714,287,448	&310,408,263	&317,109,601	&477,651,840\\ \hline
+Trans. to High&1,347,679&	1,134,479	&1,363,418&	1,369,396	&2,262,036\\ \hline
+Reg Comp&3,220,687	&3,221,904	&3,219,866	&3,221,335&	5,155,838\\ \hline
+\end{tabular}
+\end{table*}
+
+\begin{table*}
+\centering
+\caption{Mix Instruction Profile Results(B3)}
+\begin{tabular}{|c|c|c|c|c|c|} \hline
+    & Circular  & Distrubuted & Dynamic & Dynamic (optimized) & OOP\\ \hline
+Cycles&29,858	&29,914	&31,733	&30,509	&29,845\\ \hline
+
+Bit Cycles&535,533,088&	828,378,488&	352,433,432	&342,715,483	&552,968,160\\ \hline
+
+Trans. to High&4,591,862	&4,585,663	&4,588,186&	4,603,780&	6,758,124\\ \hline
+
+Reg Comp&1,688,252&	1,436,544	&1,655,666	&1,657,563	&2,802,135\\ \hline
+\end{tabular}
+\end{table*}
 

### 2011-06-08 paper/proj.tex
Results table now in implementation tex, but correctly shows up in results section.

@@
 
+\begin{table*}[!tb]
+\centering
+\caption{FP Heavy Instruction Profile Results(B1)}
+\begin{tabular}{|c|c|c|c|c|c|} \hline
+    & Circular  & Distrubuted & Dynamic & Dynamic (optimized) & OOP\\ \hline
+Cycles&82,722	&50,541&    	84,211	&84,078&        	82,153\\ \hline
+Bit Cycles&1,483,701,792&	1,399,581,372&	1,453,144,453	&1,470,555,825	&1,522,130,784\\ \hline
+Trans. to High&1,967,483&	1,719,957	&1,942,948	&1,930,476	&3,235,293\\ \hline
+Reg Comp&6,725,088&	6,631,913&	6,710,777&	6,718,043	&11,160,291\\ \hline
+\end{tabular}
+\end{table*}
+
+\begin{table*}[!tb]
+\centering
+\caption{int Heavy Instruction Profile Results(B2)}
+\begin{tabular}{|c|c|c|c|c|c|} \hline
+    & Circular  & Distrubuted & Dynamic & Dynamic (optimized) & OOP\\ \hline
+Cycles&25,781&	25,794&	26,079	&25,783	&25,780\\ \hline
+Bit Cycles&462,408,016	&714,287,448	&310,408,263	&317,109,601	&477,651,840\\ \hline
+Trans. to High&1,347,679&	1,134,479	&1,363,418&	1,369,396	&2,262,036\\ \hline
+Reg Comp&3,220,687	&3,221,904	&3,219,866	&3,221,335&	5,155,838\\ \hline
+\end{tabular}
+\end{table*}
+
+\begin{table*}[!tb]
+\centering
+\caption{Mix Instruction Profile Results(B3)}
+\begin{tabular}{|c|c|c|c|c|c|} \hline
+    & Circular  & Distrubuted & Dynamic & Dynamic (optimized) & OOP\\ \hline
+Cycles&29,858	&29,914	&31,733	&30,509	&29,845\\ \hline
+
+Bit Cycles&535,533,088&	828,378,488&	352,433,432	&342,715,483	&552,968,160\\ \hline
+
+Trans. to High&4,591,862	&4,585,663	&4,588,186&	4,603,780&	6,758,124\\ \hline
+
+Reg Comp&1,688,252&	1,436,544	&1,655,666	&1,657,563	&2,802,135\\ \hline
+\end{tabular}
+\end{table*}
+
 \subsubsection{Omitted Output Port ROB (Retention Latch)}

### 2011-06-08 src/main.cpp

@@ -185,7 +185,7 @@ ins_t *generate_instruction_stream(int benchmark, int print_flags) {
   applu173_fp.fmultProc = 0.13;
   applu173_fp.fdivProc = 0.13;
   applu173_fp.forwardingRate = 0.06;
-  applu173_fp.instCount = 100000;
+  applu173_fp.instCount = 20;
 
   ins_profile fma3d191_int;
   fma3d191_int.loadProc = 0.12;

### 2011-06-09 paper/end.tex
Results table now in implementation tex, but correctly shows up in results section.
Dynamic buffer results write up.

--- 2011-06-08 paper/end.tex
@@
+\def\topfraction{.9}
+\def\floatpagefraction{.8}
+
 \section{Results}
-hi
-\subsection{subsection titl}
-sub section.. with table!
-\begin{table*}
-\centering
-\caption{Some Typical Commands}
-\begin{tabular}{|c|c|l|} \hline
-Command&A Number&Comments\\ \hline
-\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
-\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
-\texttt{{\char'134}table}& 300 & For tables\\ \hline
-\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
-\end{table*}
-% end the environment with {table*}, NOTE not {table}!
+For each of the three modeled SPEC benchmarks, 100,000 instruction were generated, and processed with each of the reorder buffer types.
+Instruction generation is seeded, so for each particular benchmark, each type of reorder buffer processes the same exact instructions.
+We model at 256 entry reorder buffer, running with a 4 way superscalar processor having three integer functional units and one FP functional unit.
+The results are tabulated in tables \ref{tab:bench1}, \ref{tab:bench2}, and \ref{tab:bench3}.
+
+\subsection{Dynamic and Optimized Dynamic ROB}
+In the FP heavy 173.applu benchmark, neither of the dynamic ROB's fared well.
+The IPC degradation was minimal, as the basic dynamic buffer took 1.8\% extra cycles to complete the instruction streem, and the optimized version did little better, at 1.6\% more cycles.
+However, looking at the bit-cycles measurement, we see that not much bits were saved per cycle.
+Dividing by the number of cycles, we see that while the circular buffer powered an average of 17,936 bits per cycle, the dynamic buffer powered 17,256 bits per cycle, and the optimized dynamic buffer powered 17,490 bits per cycle.
+All this is to indicate that the buffer very rarely emptied to a point where a parititon could be turned off.
+This is likely because we have only 1 FP unit, thus the buffer was never able to commit more than one instruction per cycle.
+
+For the integer heavy instruction stream, the dynamic ROB's performed much better.
+The IPC degradation was again minimal, at 1.1\% for the dynamic ROB and amazingly only two cycles for the optimized version.
+This could be because the circular buffer itself rarely filled up, therefore even if the optmized dynamic buffer was running with a smaller capacity stalls due to the buffer being full was extremely rare.
+In terms of power saved, we see both the dynamic buffers were able to turn off partitions with regularity.
+Where the ciruclar buffer averaged 17,936 bits per cycle, in this instruction stream the dynamic ROB averaged 11,903 bits, and the optimized versino averaged 12,299.
+For the total bit-cycles, the dynamic buffer saved 33\%, while the optimized version saved 31\%.
+We see here that the optimized buffer opted to have more partitions enabled in exchange for better IPC performance.
+
+Lastly, the mixed 183.equake instruction stream saw the buffers perform similarly to the integer set.
+The dynamic buffer suffered 6.3\% IPC degradation, while the optimized dynamic ROB had an IPC drop of 2.2\%.
+Bit cycles wise, the dynamic buffer saved 34\%, while the optimized dynamic buffer was able to save 36\%.
+In terms of bits enabled per cycle, the dynamic ROB 11,106 bits per cycle, and the optimized buffer enabled 11,233 bits per cycle.
+
+For the FP heavy stream, we saw essentially no advantages using the dynamic approach, due to our modeled processor having only one FP functional unit clogging the buffer.
+With the other two streams, we saw dramatic improvement in power savings from the dynamic buffer, in the form of saved bit cycles.
 

### 2011-06-09 paper/end.tex
latch results

@@
 
+\subsection{Omited Output Port ROB}
+The Latch ROB results shown in the tables are produced from the parameters 
+speficed above with 16 FIFO retention latches. Looking at the result tables,
+ we see that the OOP ROB has comperable performance (in cycle) when compared 
+to basline Circular ROB.  Showing that the additional slowdown caused by the
+ non-availibilty of the operand when it enters the ROB is negligible, which
+matches with the designers argument \cite{kucuk}.
+
+Next we see that the rest of the power usage parameters are have all been 
+increased.  Both the bit cycles and transitions to high are slightly higher.
+This is to be expected since there is the additional retention latches 
+strucure, which makes the ROB larger and thus more bits to have a transition
+to the high state.  The same logic applies to the large increase of the
+register comparators used.  Each check for an operand has to check the
+register tag of all the elements in the ROB, and the rentention latches 
+as well.
+
+From the results tables, it appears that there is no reasonable reason to use
+the OOP ROB.  This is untrue, as the the read port usages for the OOP drops
+significantly.  Since we approximate about 6\% forwarding from the ROB, of 
+100,000 instructions, we would need to foward about 6,000 operands (as a
+lower bound, assuming all the instructions were INT ops).  This equates to
+192,000, bit lines that needs to be driven for operand forwading.  Also with
+the elimination of the forwarding ports, the design of the ROB becomes much 
+simplier as well. With increasing forward rates required, the energy savings
+will increase.
+
 \section{Conclusions}

### 2011-06-10 paper/end.tex
Results for distributed buffers.

--- 2011-06-09 paper/end.tex
@@
 With the other two streams, we saw dramatic improvement in power savings from the dynamic buffer, in the form of saved bit cycles.
+It is observed that the optimized ROB is a bit more performance minded than the regular dynamic ROB, averaging more bits per cycle and less cycles to complete.
+In the integer heavy set, this combination yielded a 2.2\% increase in bit-cycles for the optimized dynamic buffer, while in the mixed stream the difference in cycles pushed the optimized dynamic ROB's bit-cycles to be 2.8\% better than the non-optimized case.
 

### 2011-06-10 paper/end.tex
Dist ROB section in paper done.

@@
 The results are tabulated in tables \ref{tab:bench1}, \ref{tab:bench2}, and \ref{tab:bench3}.
+
+\subsection{Distributed ROB}
+
+Our results for the distributed ROB are largely invalid for two reasons.
+One, much of the savings in the distributed ROB comes from shorter bit-lines and less copmlex entry selection from the size reduction of each individual buffer, however this is not currently modeled in our reorder-buffer simulator.
+Next, our implementation is actually buggy, as is obvious from the results.
+
+From the first instruction stream, heavy on floating point instructions, we see the distributed buffer measures some thiry thousand less cycles to complete than the circular buffer.
+In \cite{kucuk4}, the improvement for FP instructions over the base buffer did improve very slightly (we note that they were also modelling a processor with many more FP functional units than the ones we used), but to the degree we see is impossible.
+We monitored the buffer and indeed found corruption in the contents within the first 30 instructions, but ultimately were unable to find the source of the error.
+
+For all instruction streams, we note that while the circular ROB powered 17,936 bits pe cycle (bit-cycles / cycles), the distributed ROB powered 27,692 bits per cycle.
+This is largely due to the presence of the central FIFO queue, which must keep a set of pointers to the location of actual entries in the ROB components of each functional unit.
+The extra cost of this is offset by shorter bit-lines and word-select lines, which as previously mentioned are not currently tabulated.
 

### 2011-06-10 paper/end.tex
Describe items in tables to start results section.

@@
 The results are tabulated in tables \ref{tab:bench1}, \ref{tab:bench2}, and \ref{tab:bench3}.
+
+Cycles measure the number of cycles a particular ROB implementation takes from getting the first instruction from issue to committing the last instruction in the stream.
+Bit-cycles denote total number of bits in the ROB that are powered for one cycle.
+E.g., a 20-bit buffer running for 8 cycles yields a measurement of 160 bit-cycles.
+We implement buffers that trade off execution time and extra bits for power savings in various ways, and bit-cycles, once plugged into a energy model system, provides a good measure of how much leakage power will be consumed.
+Transition to high is self-descriptive, and is included in the table as a measure of dynamic power dissipation.
+Register comparators estimate the number of times comparators are used to on register addresses and head/tail pointers.
 
@@
 For all instruction streams, we note that while the circular ROB powered 17,936 bits pe cycle (bit-cycles / cycles), the distributed ROB powered 27,692 bits per cycle.
-This is largely due to the presence of the central FIFO queue, which must keep a set of pointers to the location of actual entries in the ROB components of each functional unit.
+This is due to the presence of the central FIFO queue, which must keep a set of pointers to the location of actual entries in the ROB components of each functional unit.
+The basic circular buffer requires 70 bits per entry.
+For the FIFO queue, 37 extra bits are required for each entry in each ROB component, an increase of 53\%.
 The extra cost of this is offset by shorter bit-lines and word-select lines, which as previously mentioned are not currently tabulated.

### 2011-06-12 paper/end.tex
more future works

@@
 
+\section{Future Works}
+We have proposed a novel loosley cycle accurate power simulator for a processor's
+ROB.  With loose timing, the simulator is able to quickly simulate a stream of
+instructions.  Furthermore by producing only the power tabulation, the simulator
+remains flexible for future emerging technolgies in the underlying circut level.
+This novel approach also allows for fair power effiency comparison between
+existing and new designs.
+
+Currently we have implemented four proposed designs.  This could be expanded covering 
+more ROB designs. Since many of the techniques proposed for energy savings
+are not mutally exclusion, a model could start combineing differnt modes.
+Before preceeding forwards with more models, addtional tools may  be provided to
+catch errornous results early on.  This could prove to be quite useful as currently
+our output results do not make sense.  This could be a small bug in our source or
+a misunderstanding of the mechinism for the ROB.
+
+Additionally, we designed our simultor to be as modular as possible.  This allows
+for future expansions on the simulator. This could include additional stages of 
+the instruction lifetime such as the issue queue, or execution units.  With other
+other component simulators a entire cycle accurate simulator could be created, with 
+the flexibility to tune its accuracy and simulation run time. Another type of 
+add-on module could be created and and translate our power tabulations into actual
+power usages in joules.
+
+Lastly further would could be expaned on our instruction generation.  As we 
+currently generate synthetic instructions based on various profiles.  It would
+be ideal to load in an actual instruction trace produced by alternative cycle 
+accurate simulator.  This would
+
 \section{Conclusions}

### 2011-06-12 paper/end.tex
done?

@@
 be ideal to load in an actual instruction trace produced by alternative cycle 
-accurate simulator.  This would
+accurate simulator.  This would provide the most accurate power modeling of for
+the ROB model.
 
-\section{Conclusions}
-ending stuff
+\section{Conclusion}
+The ROB has become an intergral role in current superscalar procesors.
+As the push for higher instruction throughput the ROB has grown in size
+and complexitly.  Many new design ideas have been proposed to lower the
+power usage of the ROB.  However there still lacks a scientific, fair and
+accurate ways to compare and test the new designs.  We have created a 
+power simulator for the ROB.  This provides insight into current designs
+of the ROB.  Also the results can be used to help design the next generation
+of ROB.

### 2011-06-14 paper/bg.tex
Dist ROB section in paper done.
Editted.

@@
 to the lifetime of the instruction on top of the traditional pipeline 
-stages.  Each of which a small memory array to keep track of instruction flow.
-The first addition, the instruction queue(IQ) which holds instructions that are 
+stages.  Each of which is a small memory array to keep track of instruction flow.
+The first addition, the instruction queue(IQ), holds instructions that are 
 waiting to be executed.  Each entry will hold the instruction -- including 
 type, operand registers, and destination registers.  The second structure 
-holds the instructions that are in flight but not yet committed to the register
+(ROB) holds the instructions that are in flight but not yet committed to the register
 file. Each entry will hold both PC of the instruction, as well as the destination
@@
 
-Replacing the traditional instruction dispatch stage, An placeholder will be made for the
-instruction in ReOrder Buffer (ROB), and the instructions will be placed
+Replacing the traditional instruction dispatch stage, a placeholder will be made for the
+instruction in reorder buffer and the instructions will be placed
 in the instruction queue(IQ). In this stage instructions will first wait for their operands if not already
 available.  When all operands are available, the instruction is ready to
-be executed so wait until its appropriate execution unit is free.  At this
+be executed and waits until its appropriate execution unit is free.  At this
 point the instruction exits the IQ and begins execution.
@@
 \subsection{Circular ROB}
-The circular ROB is one of the most basic implementations of the ROB.  Simply using
-a circular buffer as the structure to hold each entry.  Since the ROB is basically
-just memory, there are two special slots for to keep track of the head and tail
-pointer to create the circular buffer.  As entries are written into the ROB, the
+The circular ROB is one of the most basic implementations of the ROB.  It is implemented simply using
+a circular buffer as the structure to hold each ROB entry.  Since the ROB is basically
+just memory, there are two special slots to keep track of the head and tail
+pointers of the circular buffer.  As entries are written into the ROB, the
 entry is written at the tail pointer and then the tail pointer is incremented.
-Likewise when an entries is written to the ARF, the head pointer is incremented.
+Likewise when a entry is written to the ARF, the head pointer is incremented.
 The circular buffer method is widely used since it inherently keeps track of the 
@@
 number of empty slots at any given point of the execution.
+
 \subsection{Dynamic ROB}
@@
 
+\begin{figure*}
+\centering
+\includegraphics[height=2in]{clusterROB.png}
+\caption{Distributed ROB layout}
+\label{fig:distrob}
+\end{figure*}
+
 \subsection{Distributed ROB}
@@
 power by reducing the bit line width as well as unit capacitance.
-
-\begin{figure*}
-\centering
-\includegraphics[height=2in]{clusterROB.png}
-\caption{Distributed ROB layout}
-\label{fig:distrob}
-\end{figure*}
 

### 2011-06-14 paper/proj.tex
Dynamic buffer results write up.
Editted.

@@
 \label{sec:circ}
-The circular ROB is the basic current implementation of the reorder buffer, so far each of the other ROB designs will use the 
+The circular ROB is the basic current implementation of the reorder buffer -- so far each of the other ROB designs will use the 
 circular array for each instruction entry.  Each instruction entry will have a flag to determine if its result is valid, a 
@@
 \end{tabular}
+\label{tab:bench1}
 \end{table*}
@@
 \end{tabular}
+\label{tab:bench2}
 \end{table*}
@@
 \end{tabular}
+\label{tab:bench3}
 \end{table*}
@@
 \subsubsection{Omitted Output Port ROB (Retention Latch)}
-The main difference between the Circular ROB and the Omitted Output Port (OOP) ROB is that the OOP adds in a retention latches and is not allowed to forward any results directly to DU.
-This our implementation of the OOP ROB simply extends the circular ROB, adding in a structure for the retention latches.
-The latches is modeled very similar to the ROB entries, an array
+The main difference between the Circular ROB and the Omitted Output Port (OOP) ROB is that the OOP adds in retention latches and is not allowed to forward any results directly to DU.
+Our implementation of the OOP ROB simply extends the circular ROB, adding in a structure for the retention latches.
+The latches are modeled very similarly to the ROB entries as simple arrays
 of a latch entry structure.
-Each latch entry contains a 32 bit value for the results to be stored as well small tag to the corresponding destination register of the result.
-As an artifact of our simulation a flag to denote that the entry is part of a floating point operation is required as well.
-Also an additional counter was added to count the additional re-forwarding when a value is committed to the ARF but also needs to be forwarded to the DU unit as well.
-Though this is not an increasing number of ports driven, it may affect future power calculations relating to capacitance.
-Lastly as changing the semantic of the number of forwards to the DU, the count refers to the number time the entry is read from a retention latch.
+Each latch entry contains a 32 bit value for the results to be stored as well as a small tag for the corresponding destination register of the result.
+As an artifact of our simulation a flag is stored to denote that the entry is part of a floating point operation.
+Also an additional counter was added to count the occurrences of re-forwarding when a value is committed to the ARF but also needs to be forwarded to the DU unit as well.
+Though this is not increasing number of ports driven, it may affect future power calculations relating to capacitance.
+Lastly as changing the semantics of the number of forwards to the DU, the count refers to the number of times the entry is read from a retention latch.
 
-As a few minor changes to the behavioral function of the OOP ROB, the first of which is the lack of the ability to forward operands.
-Thus in the simulated stage of reading from the instruction stream, when we would normally check if the operand is in the ROB and forward.
-We instead do the following, we still check to see if the operand is in the ROB, and the retention latches.
-If the entry is not in the ROB, or the latches, the operand would be available in the ARF or memory, in the converse case,  the entry in both, the ROB, and the latches, the operand is forwarded from the latches and the counter is incremented .
+There are a few minor changes to the behavioral function of the OOP ROB, the first of which is the lack of the ability to forward operands.
+Thus in the simulated stage of reading from the instruction stream, we would normally check if the operand is in the ROB and forward.
+We instead do the following: we still check to see if the operand is in the ROB but also check the retention latches.
+If the entry is not in the ROB, or the latches, the operand would be available in the ARF or memory.
+In the converse case,  the entry is in both the ROB and the latches, the operand is forwarded from the latches and the counter is incremented .
 In the case that the entry exists in the ROB, but not in the latches, a flag would be set in the ROB to forward the result when committing to the ARF.
@@
 Next in our update entries (when the cycle counter expires -- execution has completed) we would add in the new entry to the retention latch.
-Lastly when the ROB entires are valid, and ready to be committed to the ARF,
+Lastly when the ROB entires are valid, they are ready to be committed to the ARF,
 if the forwarding flag is set, the forwarding counter is incremented.
@@
 As technology changes, and new techniques are developed to decease circuit level power usages, our simulation will become less and less accurate.
-Instead we chose to simply tabulate each event that is associated with power cost, such as bit flips from 0 to 1, register comparator usage, and bit lines driven.
+Instead we chose to simply tabulate each behavioral event that is associated with power cost, such as bit flips from 0 to 1, register comparator usage, and bit lines driven.
 To help measure leakage power and overall states of the ROB, we also keep track of the time (in cycles) each bit stays in the high state, or the low state.

### 2011-06-14 paper/start.tex
Editted.

@@
 \begin{abstract}
-        In an effort to increase instruction level parallelism techniques such
+        In an effort to increase instruction level parallelism, techniques have been developed such
 as pipelining, out of order execution and operand forwarding.  However
@@
 structure has grown to be monolithic with numerous entries and read/write ports.
-There is no current methods to efficiently and equally compare various
+There are no current methods to efficiently and equally compare various
 ROB advancements to improve power efficiency. We propose a loosely cycle
@@
 are necessary  on top of the traditional pipelined architecture.  Both of these additional
-structures are essentially memory elements. The first one to keep track of which instructions
-are ready to be executed and the other known as the ROB, allows the processor  to commit the
-results of execution in program order to maintain program correctness.
+structures are essentially memory elements. The first structure is to keep track of which instructions
+are ready to be executed and the other, known as the reorder buffer (ROB), allows the processor to 
+execute instructions out of order but commit the results of execution in program order to maintain 
+program correctness.
 
-Through out its evolution the the ROB has grown to a large monolithic structure that consumes
-on average 7\% of the overall processor power budget\cite{rabey}.  Various new implementations
+Throughout its evolution the the ROB has grown to a large monolithic structure that consumes
+on average 7\% of the overall processor power budget\cite{rabaey}.  Various new implementations
 of the ROB have been proposed in an effort to reduce power consumption.  Some of which 
-dynamically change the the size of the ROB, or a banked buffer structure and other proposals argues in the 
+dynamically change the the size of the ROB or propose a banked buffer structure and other proposals argues in the 
 extreme case to eliminate read ports all together\cite{kucuk}\cite{kucuk2} \cite{kucuk3}.
@@
 becomes hard to determine the gain of each new proposed implementation.  We propose a new
-a loosely cycle accurate ROB simulator that provides a fair comparison of  different ROB designs.
-ROB\_pwr is a simulator that tabulates each action that would cost power.  This allows for a fair
+and loosely cycle accurate ROB simulator that provides a fair comparison of  different ROB designs.
+ROB\_pwr is a simulator that tabulates each hardware action that would cost power.  This allows for a fair
 comparison between each ROB design, without the concern for the underlying technology.

### 2011-06-14 src/rob_circ.cpp

@@ -218,11 +218,6 @@ void ROB_Circ::write_entry (entry *entry, ins_t ins)
   entry->cycles = ins_cyc[ins.type];
   if (ins.type >= FADD)
     entry->cycles += m_fp_delay;
-  entry->pc = ins.pc;
-  entry->reg_id = ins.regs & reg_mask;
-  entry->isfp = (ins.type >= FADD);
-
-  m_tail = tail_incr (m_tail);
   uint32_t operandRegA = ((ins.regs >> 4) & reg_mask);
   uint32_t operandRegB = ((ins.regs >> 8) & reg_mask);
   uint32_t extraCyclesA = 0;
@@ -241,6 +236,12 @@ void ROB_Circ::write_entry (entry *entry, ins_t ins)
   if(maxExtraCycles)
     //cout << "Increasing ins.pc=" << ins.pc << " cycle count by " << maxExtraCycles << " to allow forwarding" << endl;
   entry->cycles += maxExtraCycles;
+
+  entry->pc = ins.pc;
+  entry->reg_id = ins.regs & reg_mask;
+  entry->isfp = (ins.type >= FADD);
+
+  m_tail = tail_incr (m_tail);
 }
 
 void ROB_Circ::pre_cycle_power_snapshot () {

### 2011-06-14 src/rob_dist.cpp

@@ -129,7 +129,9 @@ void ROB_Dist::update_entries ()
   do {
     fifo_entry_t *fifo_entry = return_fifo_entry (i);
     entry_t *entry = get_entry(fifo_entry->rob_id, fifo_entry->rob_index);
-    if (entry->cycles && --entry->cycles == 0) {
+    if (m_print & DBG_FLAG)
+      cout << "Decrementing central_fifo[" << i << "]: " << entry->cycles << "->" << entry->cycles-1 << endl;
+    if (entry->cycles > 0 && --entry->cycles == 0) {
       if (m_print & DBG_FLAG)
         cout << "Got result for central_fifo[" << i << "] -> rob[" << fifo_entry->rob_id << "][" << fifo_entry->rob_index << "]" << endl;
       entry->valid = true;
@@ -272,8 +274,10 @@ uint32_t ROB_Dist::getCyclesToCompletion(uint32_t reg) {
     do {
       fifo_entry_t *fifo_entry = return_fifo_entry(i);
       entry_t *entry = get_entry(fifo_entry->rob_id, fifo_entry->rob_index);
-      if(entry->reg_id == reg)
+      if(entry->reg_id == reg) {
+        cout << "Found reg: " << reg << "in entry.pc:" << entry->pc << endl;
         return entry->cycles;
+      }
       i = fifo_ptr_incr(i);
     } while(i != fifo_tail);
   if (m_print & DBG_FLAG)
@@ -292,6 +296,26 @@ void ROB_Dist::write_entry(uint32_t clusterId, ins_t ins)
   entry->cycles = ins_cyc[ins.type];
   if (ins.type >= FADD)
     entry->cycles += m_fp_delay;
+  uint32_t operandRegA = ((ins.regs >> 4) & reg_mask);
+  uint32_t operandRegB = ((ins.regs >> 8) & reg_mask);
+  uint32_t extraCyclesA = 0;
+  uint32_t extraCyclesB = 0;
+
+  if(isinROB (operandRegA)) {
+    extraCyclesA = getCyclesToCompletion(operandRegA);
+    cout << "Forwarding ins.pc=" << ins.pc << " operandA in " << extraCyclesA << " cycles" << endl;
+    m_nwdu++;
+  }
+  if(isinROB (operandRegB)) {
+    extraCyclesB = getCyclesToCompletion(operandRegB);
[…]

### 2011-06-14 src/main.cpp
Output cosmetic changes

@@ -140,7 +140,9 @@ void help ()
 {
   cout << "ROB Power Simulator usage information:" << endl;
   cout << "  $ rob_pwr [OPTIONS]..." << endl;
-  cout << "Options" << endl;
+  cout << "Example use:" << endl;
+  cout << "./rob_pwr -m circular -b 1" << endl;
+  cout << "Options:" << endl;
   cout << "  -d         Enable debug messages." << endl;
   cout << "  -h         Display this help message." << endl;
   cout << "  -m         Set operating mode. See \"Modes\"" << endl;

### 2011-06-14 src/rob_dist.cpp

@@ -275,7 +275,7 @@ uint32_t ROB_Dist::getCyclesToCompletion(uint32_t reg) {
       fifo_entry_t *fifo_entry = return_fifo_entry(i);
       entry_t *entry = get_entry(fifo_entry->rob_id, fifo_entry->rob_index);
       if(entry->reg_id == reg) {
-        cout << "Found reg: " << reg << "in entry.pc:" << entry->pc << endl;
+        //cout << "Found reg: " << reg << "in entry.pc:" << entry->pc << endl;
         return entry->cycles;
       }
       i = fifo_ptr_incr(i);
@@ -303,17 +303,17 @@ void ROB_Dist::write_entry(uint32_t clusterId, ins_t ins)
 
   if(isinROB (operandRegA)) {
     extraCyclesA = getCyclesToCompletion(operandRegA);
-    cout << "Forwarding ins.pc=" << ins.pc << " operandA in " << extraCyclesA << " cycles" << endl;
+    //cout << "Forwarding ins.pc=" << ins.pc << " operandA in " << extraCyclesA << " cycles" << endl;
     m_nwdu++;
   }
   if(isinROB (operandRegB)) {
     extraCyclesB = getCyclesToCompletion(operandRegB);
-    cout << "Forwarding ins.pc=" << ins.pc << " operandB in " << extraCyclesB << " cycles" << endl;
+    //cout << "Forwarding ins.pc=" << ins.pc << " operandB in " << extraCyclesB << " cycles" << endl;
     m_nwdu++;
   }
   uint32_t maxExtraCycles = max(extraCyclesA, extraCyclesB);
   if(maxExtraCycles)
-    cout << "Increasing ins.pc=" << ins.pc << " cycles cout by " << maxExtraCycles << " to allow forwarding" << endl;
+    //cout << "Increasing ins.pc=" << ins.pc << " cycles cout by " << maxExtraCycles << " to allow forwarding" << endl;
   entry->cycles += maxExtraCycles;
 
   entry->pc = ins.pc;
